### 2016-03-27 tests/HJM/manuscript.tex
struggle to make latex work

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

%line suggested by my compiler (Fabian)
\pgfplotsset{compat=1.9}

% put figures at the end of the document, while I am writing is very annoying to have the figures in the middle of the text! 
% \usepackage[nomarkers,figuresonly]{endfloat}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\input{macros.tex}

\title[MIMC and MLMC HJM]{Multi-Index and Multi-Level Monte Carlo Evaluation of HJM Models}
\author{Juho H\"app\"ol\"a}
\address{\kaustaddress}
\email{juho.happola@iki.fi}

\date{\today}    
                                    % Activate to display a given date or no date
\begin{document}

\begin{abstract}
Abstract here!!!
\end{abstract}

\maketitle
%\tableofcontents

\section{Introduction}
\label{section:Introduction}

%\subsection{Levy processes and finances}
\levy processes form a rich field within mathematical
finance. They allow modelling of asset prices with possibly discontinuous dynamics. An early and probably the best known model involving
a \levy process is
the \cite{Merton1976125} model, which generalises the \cite{black1973pricing} model. More recently, we have seen more complex models allowing
for more general dynamics of the asset price. Examples of such models include
the \cite{kou2002jump} model (see also \cite{Dotsis20073584}), the Normal Inverse Gaussian
model (\cite{barndorff1997normal, rydberg1997normal}),
the Variance Gamma model (\cite{madan1990variance, madan1998variance}),
and the Carr-Geman-Madan-Yor (CGMY) model
(\cite{cgmy_fine,carr2003stochastic}). For a good exposition on jump processes in finance we refer to \cite{conttankov}
(also see \cite{raible2000levy} and \cite{eberlein2001application}).

%\subsection{European option PIDE}
Prices of European options whose underlying asset is driven
by the \levy process are solutions to partial integro-differential
Equations (PIDEs) (\cite{nualart2001backward,briani_chioma_natalini,almendral2005numerical,kiessling2011diffusion})
that generalise
the Black-Scholes equation by incorporating a
non-local integral term to account for the discontinuities in
the asset price. This approach has also been extended to
cases where the option price features path dependence,
for instance in \cite{boyarchenko2002barrier} \cite{d2004penalty} and \cite{lord2008fast}.

%\subsection{Fourier methods}
The \levy-Khintchine formula provides an explicit 
representation of the characteristic function of a 
\levy process (cf, \cite{tankov2004financial}).
As a consequence, one can derive an exact expression for the 
Fourier transform of the solution of the relevant PIDE.
Using the inverse fast Fourier transform (iFFT) method, one
may efficiently compute the option price for a range of asset prices
simultaneously.
Furthermore, in the case of European call options, one may
use the duality property presented by \cite{dupire1997pricing}
and iFFT to efficiently compute option prices for a wide range of strike prices.

%\subsection{Need for an error bound and state of the art (Lee)}
Despite the popularity of Fourier methods for option pricing, few works
can be found on the error analysis and related parameter selection for these methods. 
A bound for the error not only provides an interval for the precise
value of the option, but also suggests a method to select the parameters of the numerical method.
An important
work in this direction is the one by \cite{lee2004option} in which several
payoff functions are considered for a rather general set of models, whose
characteristic function is assumed to be known.
\cite{fenglinetsky} presents the framework and theoretical approach
for the error analysis, and establishes polynomial convergence rates
for approximations of the option prices.
For a more
contemporary review on the error committed in various
FT-related methods we refer the reader to \cite{boyarchenko2011new},
that extends the classical \emph{flat} Fourier methods by deforming the integration countours
on the complex plane, studying discretely monitored barrier options
studied in \cite{de2014pricing}.

%\subsection{Our contribution and comparison with Lee}
In this work, we present a methodology for studying
and bounding  the error committed
when using FT methods to compute option prices.
We also provide a systematic way of choosing 
the parameters of the numerical method, in a way that minimises
the strict error bound, thus guaranteeing adherence to a pre-described
error tolerance. 
We focus on exponential \levy processes that may be of either
diffusive or pure jump. Our contribution is to derive a strict error bound for a Fourier
transform method when pricing options under risk-neutral 
\levy dynamics. We derive a simplified bound that separates the contributions of
the payoff and of the process in an easily processed and extensible
product form that is independent of the asymptotic behaviour of the
option price at extreme prices and at strike parameters.
We also provide a proof for the existence of optimal
parameters of the numerical computation that minimise the presented
error bound.
When comparing our work with Lee's work we find that Lee's work is more general than ours in that he studies a wider range
of processes, on the other hand, our results apply to a larger class of payoffs.
On test examples of practical relevance, we also find that the bound presented produces comparable or better results than the ones previously presented in the literature, with acceptable computational cost.

%\subsection{Organization of the paper}
The paper is organised in the following sections:
In Section \ref{section:Method}
we introduce the PIDE setting in the context of
risk-neutral asset pricing; we show the Fourier representation of the relevant
PIDE for asset pricing with \levy processes and use that
representation for derivative pricing.
In Section \ref{section:Bound}
we derive a representation
for the numerical error and divide it into
quadrature and cutoff contributions. We also describe the
methodology for choosing numerical
parameters to obtain minimal error bounds
for the FT method.
The derivation is supported by numerical examples
using relevant test cases with both diffusive and
pure-jump \levy processes in Section \ref{section:Numerics}.
Numerics are followed by conclusions in Section \ref{section:Conclusion}.

\section{Fourier method for option pricing}

\label{section:Method}
Consider an asset whose price at time $t$ is modelled 
by the stochastic process $S=(S_t)$ defined by $S_t=S_0 \expfs{X_t}$,
where $X=(X_t) \in \R$ is assumed to be
a \levy process whose jump measure $\nu$ satisfies
\begin{align} \label{eq:nuhyp}
\int_{\rz} \min\{y^2,1\} \nu \parent{dy} < \infty
\end{align}

Assuming the risk-neutral dynamic for $S_t$, the price at time $t=T-\tau$ of a European option
with payoff $G$ and maturity time $T$ is given by
\begin{align*}
\Pi(\tau,s)=\expfs{-r\tau} \expp{G\parent{S_T}|S_{T-\tau}=s}
\end{align*}
where $r$ is the short rate that we assume to be constant and $\tau\colon 0\leq\tau\leq T$ is the time to maturity. 
Extensions to non-constant deterministic short rates are straightforward.

The infinitesimal generator of a \levy
process $X$ is given by \citep[see][]{applebaum2004levy}
\begin{align}
\lx f(x)&\equiv 
\lim_{h\to 0} \frac{\expp{f(X_{t+h})|X_t=x}-f(x)}h \notag \\
&= \gamma f'(x)+\frac12\sigma^2f''(x)+\int_{\rz} \left(f(x+y)-f(x)-y1_{|y|\leq 1} f'(x)\right)\nu(\d y)
\end{align}
where $(\gamma,\sigma^2,\nu)$ is the characteristic triple of the \levy process.
The risk-neutral assumption on $(S_t)$ implies
\begin{equation} \label{eq:riskneutralexpint}
\int_{\absval{y}>1}\expfs{y} \nu(\d y)<\infty
\end{equation}
and fixes the drift term (see \cite{kiessling2011diffusion}) $\gamma$ of the \levy process to
% TODO: find more references to this
\begin{align} 
\label{drift_equation}
\gamma = r-\frac12 \sigma^2 - \int_{\rz} \parent{\expfs{y}-1-y1_{|y|\leq 1} }\nu(\d y)
\end{align}
Thus, the infinitesimal generator of $X$ may be written under the risk-neutral assumption as
\begin{align}
\lx f(x)&= \left(r-\frac{\sigma^2}2\right) f'(x)+\frac{\sigma^2}2 f''(x)+\int_{\rz} \left(f(x+y)-f(x)-(\expfs{y}-1)f'(x)\right)\nu(\d y)
\end{align}
Consider $g$ as the reward function in log prices (ie, defined by 
$g(x)=G(S_0 \expfs{x})$). Now, take $f$ to be defined as
\begin{align*}
f \parent{\tau,x} \equiv \expp{g \parent{X_{T}}|X_{T-\tau}=x}
\end{align*}
Then $f$ solves the following PIDE:
\begin{equation*}
\begin{cases}
\partial_{\tau}f(\tau,x)&=\lx f(\tau,x) \\
f(0,x)&=g(x), \qquad \qquad \parent{\tau,x}  \in [0,T] \times \R
\end{cases}
\end{equation*}
Observe that $f$ and $\Pi$ are related by
\begin{equation}
\Pi(\tau,S_0 \expfs{x})=\expfs{-r \tau}f(\tau,x)
\end{equation}
Consider a damped version of $f$ defined by $\fa(\tau,x)=\expfs{-\damp x}f(\tau,x)$; we see that $\partial_{\tau} \fa = \expfs{-\damp x} \lx f\parent{\tau,x} $.

There are different conventions for the Fourier transform. Here we consider the operator $\F$ such that 
\begin{align}
\F[f] \parent{\omega} \equiv \int_{\R}\expfs{i\omega x} f \parent{x} \d x
\label{eq:fourierDefinition}
\end{align}
defined for functions $f$ for which the previous integral is convergent. We also use $\ft{f} \parent{\omega}$ as a shorthand notation of $\F[f]\parent{\omega}$.
To recover the original function $f$, we define the inverse Fourier transform as 
$$\ift{f}(x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x}f(\omega) \d \omega$$
We have that $\ift{\ft{f}}(x)=f(x)$.

Applying $\F$ to $\fa$ we get $\ft{\fa}(\omega)=\ft{f} \parent{\omega + i \damp}$.
Observe also that the Fourier transform applied to $\lx f(\tau,x)$ gives $\charExp{-i\omega}\ft{f} \parent{\tau,\omega}$,
where $\charExp{\cdot}$ is the characteristic exponent of the process $X$, which satisfies $\expp{\expfs{zX_t}}=\expfs{t\charExp{z}}$.
The explicit expression for $\charExp{\cdot}$ is 
\begin{equation}
\label{characteristicDefinition}
\charExp{z}
=\parent{r-\frac{\sigma^2}2} z +\frac{\sigma^2}2 z^2 +\int_{\R} \left(\expfs{z y}-1-(\expfs{y}-1)z\right)\nu(\d y) \\[.5em]
\end{equation}
From the previous considerations it can be concluded that 
\begin{equation}
\partial_{\tau} \ft{\fa} = \charExp{\damp -i \omega} \ft{f} \parent{\omega - i \damp}
\end{equation}
Now $\ft{f}\parent{\omega - i \damp} = \ft{\fa} \parent{\omega}$ so $\ft{\fa}$ satisfies the following ODE
\begin{equation}
\begin{cases}
\frac{\partial_{\tau}\ft{\fa}\parent{\tau,\omega}}{\ft{\fa}(\tau,\omega)}&=\charExp{\damp -i \omega}\\[.5em]
\ft{\fa} \parent{0,\omega}&=\ft{\ga}\parent{\omega}
\end{cases}
\end{equation}

Solving the previous ODE explicitly, we obtain
\begin{equation}
\ft \fa \parent{\tau,\omega}
= \expfs{\tau \charExp{\damp-i \omega} } \ft{\ga} \parent{\omega}
\label{eq:theODE}
\end{equation}
Observe that the first factor in the right-hand side in the above 
equation is $\expp{\expfs{\parent{\damp-i \omega}X_{\tau}}}$, (ie, $\charFun{1}{-i\damp-\omega}$), where $\charFun{\tau}{\cdot}$ denotes the characteristic function of the
random variable $X_{\tau}$
\begin{align}
\charFun{\tau}{\omega} \equiv \expp{\tau \charExp{i \omega}} 
\end{align}

Now, to obtain the value function we employ
the inverse Fourier transformation, to obtain
\begin{equation}
\fa(\tau,x)=\ift{\ft{\fa}}(\tau,x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x} \ft{\fa}(\tau,\omega)d\omega
\label{eq:fadefinition}
\end{equation}
or
\begin{equation}
\fa(\tau,x)=\frac{1}{\pi}\int_{0}^{+\infty} \real{\expfs{-i\omega x} \ft{\fa}(\tau,\omega)}d\omega
\label{eq:fadefinition2}
\end{equation}

As it is typically not possible to compute the inverse Fourier transform analytically,
we approximate it by
discretising and truncating the integration domain
using trapezoidal quadrature \eqref{eq:fadefinition}.
Consider the following approximation:
\begin{align}
\label{eq:fdiscrete}
 \fadisc{\step,n} (\tau,x)&= \frac{\step}{2\pi} \sum_{k=-n}^{n-1} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}\\
 &=\frac{\step}{\pi} \sum_{k=0}^{n-1} \real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}
\end{align} 
Bounding and consequently minimising the error
in the approximation of $f(\tau,x)$ by $$\fdisc{\step,n}(\tau,x) \equiv \expfs{\damp x}\fadisc{\step,n}(\tau,x)$$
is the main focus of this paper and will be addressed in the following section.

\begin{remark}
Although we are mainly concerned with option pricing when 
the payoff function can be damped in order to guarantee
regularity in the $L^1$ sense, we note here that our main 
results are naturally extendable to include the Greeks of the option. 
Indeed, we have by \eqref{eq:theODE} that 
\begin{align}
 f \parent{t,x} = \frac{1}{2 \pi} \int_{\mathbb R} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
\end{align}
so the Delta and Gamma of the option equal
\begin{align}
\Delta \parent{t,x} &\equiv \frac{\partial f \parent{t,x}}{\partial x} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
\\
\Gamma \parent{t,x} &\equiv \frac{\partial^2 f \parent{t,x}}{\partial x^2} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega}^2 \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
%\\
%\Theta \parent{t,x}  &\equiv \frac{\partial f \parent{t,x}}{\partial t} = -\frac{1}{2 \pi} \int_{\mathbb R} \charExp{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} d\omega
\end{align}
Because the expressions involve partial derivatives with respect to only $x$,
the results in this work are applicable for the computation of $\Delta$ and $\Gamma$
through a modification of the payoff function:
\begin{align}
\ft g_{\damp, \Delta} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp - i \omega}
\\
\ft g_{\damp, \Gamma} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp-i \omega}^2
\end{align}
When the Fourier space payoff function manifests
exponential decay, the introduction of a coefficient that is polynomial in $\omega$
does not change the regularity of $\ft g$ in a way that would significantly change the following
analysis. Last, we note that since we do our analysis for PIDEs on a mesh of
$x$'s, one may also compute the option values in one go and obtain the Greeks with
little additional effort using a finite difference approach for the derivatives.
\end{remark}

\subsection{Evaluation of the method for multiple values of $x$ simultaneously}

The Fast Fourier Transform (FFT) algorithm provides an efficient 
way of computing \eqref{eq:fdiscrete} for an equidistantly
spaced mesh of values for $x$ simultaneously.
Examples of works that consider this widely extended tool are \cite{lord2008fast,jackson2008fourier,hurd2010fourier} and \cite{schmelzle2010option}.
%In Subsection \ref{subsection:FFT} we analyse this alternative.

Similarly, one may define the Fourier frequency $\omega$
as the conjugate variable of some external parameter on which
the payoff depends. Especially, for the practically relevant case
of call options, we can denote the log-strike as $k$ and treat
$x$ as a constant and write:
\begin{align}
\kft f_{k,\damp} \parent{\omega} \equiv
\int_\R \expfs{\parent{\damp +i \omega }k } f_k \parent{x} \d k
\end{align}

Using this convention, the time dependence is given by
\begin{align}
\label{eq:kftDef}
\kft f_{k,\damp} \parent{\tau,x} = 
\frac{\expfs{\parent{i \omega + \damp + 1} x} \charFun{\tau}{\omega - i \parent{\damp +1} }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
\end{align}
contrasted with the $x$-space solution
\begin{align}
\label{eq:xftDef}
\ft f_{k,\damp} \parent{\tau,x} = 
\frac{\expfs{\parent{i \omega - \damp + 1} k} \charFun{\tau}{\omega + i \damp }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
\end{align}
We note that for call option payoff to be in $L^1$,
we demand that $\damp$ in \eqref{eq:kftDef} is positive.
Omitting the exponential factors that contain the $x$ and $k$
dependence in \eqref{eq:kftDef} and \eqref{eq:xftDef} respectively,
we have that one can arrive from \eqref{eq:kftDef} to \eqref{eq:xftDef}
using the mapping $\damp \mapsto -\damp-1 $. Thanks to this,
much of the analysis regarding the $x$-space transformation
generalises in a straightforward manner to the $k$-space transform.

\section{Error bound}
\label{section:Bound}

The aim of this section is to compute a bound of the error when approximating 
the option price $f(\tau,x)$ by $\fadisc{\step,n} (\tau,x)$,
defined in \eqref{eq:fdiscrete}.
Considering
\begin{equation} \label{eq:serie}
\fadisc{\step}(\tau,x)= \frac{\step}{2\pi} \sum_{k\in \Z} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}
\end{equation}
the total error $\Error$ can be split into a sum of two terms: the quadrature and truncation errors.
The former is the error from the approximation of the integral in \eqref{eq:fadefinition}
by the infinite sum in \eqref{eq:serie}, while the latter is due to the truncation of the infinite sum. 
Using triangle inequality, we have
\begin{align} \label{eq:errordef}
\Error &:=\absval{f(\tau,x)-\fdisc{\step,n}(\tau,x)}\leq \EQ + \EF
\end{align}
with
\begin{align*}
\EQ &= \expfs{\damp x}\absval{\fa(\tau,x)-\fadisc{\step}(\tau,x)}\\
\EF &= \expfs{\damp x} \absval{\fadisc{\step}(\tau,x)-\fadisc{\step,n}(\tau,x)}
\end{align*}
Observe that each $\Error$, $\EQ$ and $\EF$ depend on
three kinds of parameters:
\begin{itemize}
 \item Parameters underlying the model and payoff such as volatility and strike price. We call these \emph{physical parameters.}
 \item Parameters relating to the numerical scheme such as $\damp$ and $n$.
 \item \emph{Auxiliary parameters} that will be introduced in the process of deriving the error bound. These parameters
 do not enter the computation of the option price, 
 but they need to be chosen appropriately to have as tight a bound as possible. 
\end{itemize}

We start by analysing the quadrature error.
\subsection{Quadrature error}

Denote by $\strip{a}$, with $a>0$, the strip of width $2a$ around the real line:
$$\strip{a} \equiv \left\{z\in \C \colon \absval{\im{z}}<a \right\}$$
The following theorem presents conditions under which the quadrature error goes to zero at a spectral rate as $\step$ goes to zero.
Later in this section, we discuss simpler conditions to verify the hypotheses and analyse in more detail the case 
when the process $X$ is a diffusive process or there are ``enough small jumps."

\begin{theorem}\label{theo:quadrature} 
Assume that for $a > 0$:
\begin{itemize}
\item[H1.] the characteristic function of the random variable $X_1$ has an analytic extension to the set $$\strip{a}-\damp i \equiv \left\{z\in \C \colon \absval{\im{z}+\damp}<a\right\}$$
\item[H2.] the Fourier transform of $\ga(x)$ is analytic in the strip $\strip{a}$ and
\item[H3.] there exists a continuous function $\gamma \in L^1(\R)$ such that  $\absval{\ft{\fa}(\tau,\omega+i \beta)}<\gamma(\omega)$ for all $\omega\in \R$ and for all $\beta \in [-a,a]$
\end{itemize}
Then the quadrature error is bounded by
\begin{equation*}
\ErrorQuadrature \leq \expfs{\damp x} \frac{M_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
\end{equation*}
where $M_{\damp,a}(\tau,x)$ is given by
\begin{align}\label{eq:defM}
M_{\damp,a}(\tau,x)&:=\sum_{\beta \in \{-a,a\}}\int_{\R}\absval{\expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)} \d \omega
\end{align}
$M_{\damp,a}(\tau,x)$ equals the Hardy norm (defined in \eqref{eq:hardynorm}) of the function $\omega \mapsto \expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)$, which
is finite.
\end{theorem}

The proof of Theorem \ref{theo:quadrature} is an application of Theorem 3.2.1 in \cite{stenger2012numerical},
whose relevant parts we include for ease of reading. Using the notation in \cite{stenger2012numerical}, $H^1_{\strip{a}}$ is the family of functions $w$ that are
 analytic in $\strip{a}$ and such that 
\begin{equation}\label{eq:hardynorm}
\nnorm{w}{H^1_{A_a}}:= \lim_{\varepsilon\to 0} \int_{\partial \strip{a}(\varepsilon)} \absval{w(z)}d\absval{z} < \infty
\end{equation}
 where 
 $$\strip{a}(\varepsilon) = \left\{z\in \C\colon \absval{\real{z}}<\frac1{\varepsilon},\ \absval{\im{z}} < a \parent{1-\varepsilon}\right\}$$
\begin{lemma}[Theorem 3.2.1 in \cite{stenger2012numerical}]\label{lem:stenger}
Let $w \in H^1_{A_a}$,
then define:
\begin{align}
I \parent{w} &= \int_\R w \parent{x} \d x \\
J \parent{w,h} &= h \ssum{j=-N}{N} w \parent{jh} \\
\zeta \parent{w,h} &= I \parent{w}- J \parent{w,h}
\end{align}
then
\begin{align}
\absval{\zeta \parent{w,h}} \leq \frac{e^{-\frac{\pi a}{h}} \nnorm{w}{H^1_{A_a}}}{2 \mathrm{sinh} \parent{\frac{\pi a}{h}}}
\end{align}
\end{lemma}

\begin{proof}[Proof of Theorem \ref{theo:quadrature}]
First observe that H1 and H2 imply that the function $w(z)=\expfs{-i x z + \step/2}\ft{\fa}(\tau,z+ \step/2)$ is analytic in $\strip{a}$. 
H3 allows us to use dominated convergence theorem to prove that $\nnorm{w}{H^1_{A_a}}$ is finite and coincides with $M_{\damp,a}(\tau,x)$. Applying Lemma \ref{lem:stenger} 
the proof is completed.
\end{proof}

Regarding the hypotheses of Theorem \ref{theo:quadrature}, the next propositions provide simpler conditions that imply H1 and H2 respectively.

\begin{proposition} \label{prop:analy1}
If $\alpha$, $a$ and $\nu$ are such that
\begin{equation}
\label{eq:alterH1}
\int_{y>1} \expfs{\parent{\damp+a}y}\nu(\d y) < \infty \quad \text{and}\quad \int_{y<-1} \expfs{\parent{\damp-a}y}\nu(\d y) < \infty
\end{equation}
then H1 in Theorem \ref{theo:quadrature} is fulfilled.

\begin{proof}
Denoting by $\charFun{1}{\cdot}$ the characteristic function of $X_1$, we want to prove that $z\mapsto \charFun{1}{z+\damp i}$ is analytic in $\strip{a}$. Considering that $\charFun{1}{z+\damp i}=\expfs{\charExp{iz-\damp}}$, the only non-trivial part of the proof is to verify that 
\begin{equation}
z\mapsto \int_{} p(z,y)\nu(\d y)
\end{equation}
is analytic in $\strip{a}$,
where $p\colon \strip{a} \times \R \to \C$ is given by
$$p(z,y)=\expfs{y \parent{i z-\damp}}-1-(\expfs{y}-1)\parent{i z-\damp}$$
To prove this fact, we show that we can apply the main result and the only theorem in \cite{mattner2001complex}, 
which, given a measure space $(\Omega,\mathcal{A},\mu)$ and an open subset $G\subseteq \C$, 
ensures the analyticity of $\int f(\cdot,\omega) d\mu(\omega)$, provided that $f\colon G\times \Omega \to \C$ satisfies: $f(z,\cdot)$ is $\mathcal{A}$-measurable for all $z\in G$; $f(\cdot,\omega)$ is holomorfic for all $\omega \in \Omega$; and $\int \absval{f(\cdot,\omega)} d\mu(\omega)$ is locally bounded.
In our case we consider the measure space to be $\R$ with the Borel $\sigma$-algebra and the Lebesgue measure, $G=\strip{a}$ and $f=p$. It is clear that 
$p(x,\cdot)$ is Borel measurable and $p(\cdot,y)$ is holomorphic. It remains to verify that  
\begin{equation*}
z \mapsto \int_{\R^*} \absval{p(z,y)}\nu(\d y)
\end{equation*} 
is locally bounded. 
To this end, we assume that $\real{z}<b$ (and, since $z\in \strip{a}$, $\im{z}<a$) and split the integration domain in $\absval{y}>1$ and ${0<\absval{y}\leq 1}$ to prove that both integrals are uniformly bounded.

Regarding the integral in ${\absval{y}> 1}$, we observe that
\begin{equation}
\absval{p(z,y)}\leq \expfs{y(\damp+\im{z})}+1+(\expfs{y}+1)(\damp+a+b)
\end{equation}
for $y<-1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp-a)}$ while for $y>1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp+a)}$. 
Using the previous bounds and the hypotheses together with \eqref{eq:nuhyp} and \eqref{eq:riskneutralexpint}, we obtain the needed bound.

For the integral in ${0<\absval{y}\leq 1}$, observe that, denoting $f(z,y)=\absval{p(z,y)}$, we have, $f(z,0)=0$ for every $z$, $\partial_y f(z,0)=0$ for every $z$, and $\absval{\partial_{yy}f(z,y)}< c$ for $z\in \strip{a}, \real{z}<b, \absval{y}<1$. From these observations we get that the McLaurin polynomial of degree one of $y\mapsto f(z,y)$ is null for every $z$, and we can bound $f(z,y)$ by the remainder term, which, in our region of interest, is bounded by $\frac{c}{2} y^2$, obtaining
\begin{equation}
\int_{0<\absval{y}\leq 1} \absval{p(z,y)}\nu(\d y)\leq \frac{c}2 \int_{0<\absval{y}\leq 1}y^2\nu(\d y)
\end{equation}
which is finite by hypothesis on $\nu$, which finishes the proof.
\end{proof}
\end{proposition}

\begin{proposition}
If for all $b<a$, the function $x\mapsto \expfs{b\absval{x}}{\ga(x)}$ is in $L^2(\R)$ then H2 in Theorem \ref{theo:quadrature} is fulfilled.
\end{proposition}
\begin{proof}
The proof is a direct application of Theorem IX.13 in \cite{reedsimon}
\end{proof}
We now turn our attention to a more restricted class of Levy processes. 
Namely, processes such that either $\sigma^2>0$ or there exists $\lambda \in (0,2)$ such that $\cl$ defined in \eqref{eq:PureJumpC} is strictly positive. 
For this class of processes, we can state our main result explicitly in terms of the characteristic triplet.

Given $\lambda \in (0,2)$, define $\cl$ as
\begin{equation}
\label{eq:PureJumpC}
\cl  = 
\textrm{inf}_{\kappa > 1} \left\{\kappa^{\lambda} 
 \int_{0<|y| < \frac{1}{\kappa}} y^2 \nu(\d y)\right\}
\end{equation}
Observe that $\cl\geq 0$ and, by our assumptions on the jump measure $\nu$, $\cl$ is finite. Furthermore, if $\lambda \in (0,2)$ is such
that
\begin{equation}
\label{eq:pureJumpAssumption}
\liminf_{\epsilon \downarrow 0} \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y) > 0
\end{equation}
then $\cl>0.$ To see this, observe that \eqref{eq:pureJumpAssumption} implies the existence of $\epsilon_0$ such that 
\begin{equation*}
\inf_{\epsilon \leq \epsilon_0}\left\{ \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} > 0
\end{equation*}
If $\epsilon_0<1$ observe that
\begin{equation*}
\inf_{\epsilon_0 \leq \epsilon \leq 1}\left\{ \frac{1}{\epsilon^{\lambda}}
 \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} \geq  \int_{0<|y| < \epsilon_0} y^2 \nu(\d y) > 0
\end{equation*}
where for the first inequality it was taken into account that $\frac{1}{\epsilon^{\lambda}}\geq 1$ and that the integral is increasing with $\epsilon$.
Combining the two previous infima and considering $\absval{\kappa}=\frac{1}{\epsilon}$ we get that $\cl>0.$

Furthermore, we note that for a \levy model with finite jump
intensity, such as the Black-Scholes and Merton models that satisfy the first of our
assumption, $\cl=0$ for all $\lambda \in (0,2)$.

\begin{theorem} \label{teo:boundM}
Assume that: $\damp$ and $a$ are such that \eqref{eq:alterH1} holds; $\ft{\ga} \in \linf{\strip{a}}$; and either $\sigma^2>0$ or $\cl>0$ for some $\lambda\in (0,2)$.
Then the quadrature error is bounded by
\begin{equation*}
\ErrorQuadrature \leq \expfs{\damp x} \frac{\tilde{M}_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
\end{equation*}
where 
\begin{equation}\label{eq:Mbound}
\tilde{M}_{\damp,a} \parent{\tau,x} 
=
\ssum{c\in \sset{-1,1}}{} 
\expfs{cax}
\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
\int_{\R}  \expfs{-\tau \left( \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}\right) } \d \omega
\end{equation}
Furthermore, if $\sigma^2>0$ we have
\begin{equation}
\label{eq:Mboundsigpositive}
\tilde{M}_{\damp,a} \parent{\tau,x}  \leq \frac{\sqrt{2\pi}}{\sigma\sqrt{\tau}}\ssum{c\in \sset{-1,1}}{} 
\expfs{cax}
\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
\end{equation}

\begin{proof}
Considering $h_{\damp,a}(\tau,x,\omega)$ defined by
\begin{equation}
h_{\damp,a} \parent{\tau,x,\omega} = \ssum{c \in \sset{-1,1}}{} \absval{\expfs{-i\parent{\omega + ica} x}\ft{\fa}(\tau,\omega+i ca)} 
\end{equation}
we have that
$$M_{\damp,a}(\tau,x) = \int_{\R}h_{\damp,a} \parent{\tau,x,\omega} d\omega$$
On the other hand, for  $\beta\in (-a,a)$:
\begin{equation}
\absval{\expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)} = \expfs{\beta x} \absval{\ft{\fa}(\tau,\omega+i \beta)}=\expfs{\beta x}\absval{\expfs{\tau \charExp{\damp+\beta-i\omega}}}\absval{ \ft{\ga}\parent{\omega+i\beta}}
\end{equation}
For the factor involving the characteristic exponent we have
\begin{equation}
\absval{\expfs{\tau \charExp{\damp+\beta-i\omega}}}=\expfs{\tau \real{\charExp{\damp+\beta-i\omega}}}
\end{equation}

Now, observe that
\begin{align}
\notag \real{\charExp{ \damp +\beta - i \omega}}&= \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
\label{eq:boundingPsiPureJump} &\quad +\int_{\rz}\parent{ \expfs{\parent{\damp + \beta} y} \cos(-y\omega)-1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y)
\end{align}

If $|\omega| \leq 1$ we bound $\cos(-y \omega)$ by 1, getting
\begin{align}
\notag \real{\charExp{ \damp +\beta - i \omega}} &\leq \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
\notag &\quad +\int_{\rz} \parent{\expfs{\parent{\damp + \beta} y} -1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y)  \\
& = \charExp{\damp+\beta} -\frac{\sigma^2}{2}\omega^2 \label{eq:Psibound}
\end{align}

Assume $| \omega| > 1$.
Using that for $|x| < 1$ it holds that $\textrm{cos}(x) < 1 - x^2 / 4$, we can bound the first term of the integral in the 
following manner:
\begin{align}
\notag \int_{\rz} \expfs{\parent{\damp + \beta} y} \cos(y\omega) \nu(\d y) & \leq
\int_{0<|y| < 1/|\omega|} \expfs{\parent{\damp + \beta} y} \left( 1 - \omega^2 y^2 / 4 \right)  \nu(\d y) \\
\notag & \quad + 
\int_{|y| \geq  1/|\omega|} \expfs{\parent{\damp + \beta} y}  \nu(\d y) \\
\notag & 
\leq \int_{\rz} \expfs{\parent{\damp + \beta} y}  \nu(\d y) 
- \frac{|\omega|^{2 - \lambda}}{4} | \omega |^{\lambda} \int_{0<|y| < 1 / |\omega|} y^2 \nu(\d y)  \\
\label{eq:extraTermPureJump} & 
\quad 
\leq \int_{\rz} \expfs{\parent{\damp + \beta} y}  \nu(\d y) 
- \frac{|\omega|^{2 - \lambda}}{4} \cl
\end{align}
Inserting \eqref{eq:extraTermPureJump} back into \eqref{eq:boundingPsiPureJump} we get
\begin{align*}
\notag 
\real{\charExp{ \damp +\beta - i \omega}} & \leq \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
\notag &\quad + \int_{\rz} \parent{\expfs{\parent{\damp + \beta} y} -1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y) \\
\notag & \quad \quad - \frac{|\omega|^{2 - \lambda}}{4} \cl    \\
& = \charExp{\damp+\beta} -\frac{\sigma^2}{2}\omega^2 - \frac{|\omega|^{2 - \lambda}}{4} \cl
\end{align*}

%We conclude that $\real{\Psi \parent{ \damp - i z}}\to -\infty$ uniformly in $A$ when $\absval{z}\to \infty$, and in consequence $h(z)\to 0$.
Taking the previous considerations and integrating in $\R$ with respect to $\omega$, we obtain \eqref{eq:Mbound}.

Finally, observing that  $\cl\geq 0$ and bounding it by 0, the bound \eqref{eq:Mboundsigpositive} is obtained by evaluating the integral.
\end{proof}
\end{theorem}

\begin{remark} \label{remark:callPut}
In the case of call options, hypothesis H2
implies a dependence between the strip-width parameter
$a$ and damping parameter $\damp$. We have that the
damped payoff of the call option is in $L^1 \parent{\R}$
if and only if $\damp > 1$ and hence the appropriate choice
of strip-width parameter is given by $0<a< \damp-1$.
A similar argument holds for the case of put options, for
which the Fourier-transformed damped payoff is identical
to the calls with the distinction that $\damp <0$.
In such case, we require $a < -\damp$.

The case of binary options whose payoff has finite support
($G \parent{x} = \mathbf{1}_{[x_-,x_+]} \parent{x}$) 
we can set any $a \in \R$ (ie, no damping is needed at all and
even if such damping is chosen, it has no effect on the appropriate
choice of $a$). 
\end{remark}

\begin{remark}
The bound we provide for the quadrature error is naturally positive and increasing in $\step$.
It decays to zero at a spectral rate as $\step$ decreases to 0.
\end{remark}

\subsection{Frequency truncation error}
The frequency truncation error is given by
\begin{align*}
\ErrorFrequency &= \frac{\expfs{\damp x}\step}{\pi} \absval{\sum_{k=n}^{\infty} \real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}}
\end{align*}
If a function $c\colon (\omega_0,\infty) \to (0,\infty)$ satisfies
\begin{equation} \label{eq:condFunc}
\absval{\real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}} \leq c\parent{\parent{k+\frac12}\step}
\end{equation}
for every natural number $k$, then we have that 
\begin{align*}
\ErrorFrequency &\leq \frac{\expfs{\damp x}\step}{\pi} \sum_{k=n}^{\infty}\absval{\real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}} \\
& \leq \frac{\expfs{\damp x}\step}{\pi} \sum_{k=n}^{\infty} c\parent{\parent{k+\frac12}\step}
\end{align*}
Furthermore, if $c$ is a non-increasing concave integrable function, we get 
\begin{equation}
\ErrorFrequency \leq \frac{\expfs{\damp x}}{\pi} \int_{n\step}^{\infty} c(\omega) \d\omega
\label{eq:truncationErrorIntegral}
\end{equation} 
When $\ft{\ga}\in \linf{[\omega_0,\infty)}$ and either $\sigma^2>0$ or $\cl>0$, then the function $c$ in \eqref{eq:condFunc} can be chosen as
\begin{equation} \label{eq:cutoffparticularSimpler}
c(\omega)=\norminf{\ft{\ga}}{[\omega_0,\infty)}\expfs{\tau \charExp{\damp}}\expfs{-\tau \parent{\frac{\sigma^2}{2} \omega^2 +\frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1} }}
\end{equation}

To prove that this function satisfies \eqref{eq:condFunc} we can use the same bound we found in the proof of Theorem \ref{teo:boundM}, with $\beta=0$, to obtain
\begin{equation*}
\real{\charExp{ \damp - i \omega}}\leq  \charExp{\damp} - \frac{\sigma^2}{2}\omega^2  - \frac{|\omega|^{2 - \lambda}}{4} \cl \indicator_{\absval{\omega}>1}
\end{equation*}
from where the result is straightforward.

\subsection{Bound for the full error}

In this section we summarize the bounds obtained for the error under different 
assumptions and analyse their central properties.

In general the bound provided in this paper are of the form
\begin{equation}\label{eq:fullErrorGeneral}
\est = \frac{\expfs{\damp x}}{\pi}\parent{ \frac{\bar{M}}{\expfs{2\pi a/\step}-1} +  \int_{n\step}^{\infty}  c\parent{\omega} \d \omega}
\end{equation}
where $\bar{M}$ is an upper bound of $M_{\damp,a}\parent{\tau,x}$ defined in \eqref{eq:defM} and $c$ is non-increasing, integrable and satisfies
\eqref{eq:condFunc}. Both $\bar{M}$ and $c$ may depend on the 
parameters of the model and the artificial parameters, but they are independent of $\step$ and $n$. Typically one can remove 
the dependence of some of the parameters, simplifying the expressions but obtaining
less tight bounds.

When analysing the behaviour of the bound one can observe that the term correspondent with the quadrature error decreases to zero
spectrally when $\step$ goes to 0. 
The second term goes to zero if $n\step$ diverges, but we are unable to determine
the rate of convergence without further assumptions.

Once an expression for the error bound is obtained, the problem of 
how to choose the parameters of the numerical method to minimise the bound arises, 
assuming a constraint on the computational effort one is willing to use. 
The computational effort of the numerical method depends
only on  $n$. For this reason we aim at finding the parameters that
minimise the bound for a fixed $n$.
The following result shows that the bound obtained, as a function of $\step$, 
has a unique local minimum, which is the global minimum.
\begin{proposition}
Fix $\alpha$, $a$, $n$, and $\lambda$ and consider the bound $\est$ as a
function of $\step$. There exists an optimal $\step^*\in [\frac{\omega_0}{n},\infty)$ such that $\est$ is
decreasing in $(\frac{\omega_0}{n},\step^*)$ and increasing in
$(\step^*,\infty)$; thus, a global minimum of $\est$ is attained at $\step^*$.

Furthermore, the optimal $\step$ is either the only point in which 
$\step\mapsto p\parent{n\step,b}-c(n\step)$, with $p$ defined in \eqref{eq:defp}, changes sign, or $\step=\frac{\omega_0}{n}$ if $p\parent{\omega_0,b}-c(\omega_0)>0$.

\begin{proof}
Let us simplify the notation by calling $y=n\step$, $b=2 \pi a n$ and $\tilde{\E}=\pi \expfs{-\damp x}\est$. We want to prove the existence of $y^*\colon y^*\geq \omega_0$ such that $\tilde{\E}(y)$ is decreasing for $\omega_0<y< y^*$ and increasing for $y> y^*$. We have 
$$\tilde{\E}(y)=\frac{\bar{M}}{\expfs{b/y}-1} +  \int_{y}^{\infty} c(\omega) \d\omega.$$
The first term is differentiable with respect to $y$ and goes to 0 if $y\to 0^+$. This allows us to express it as an integral of its derivative. We can then express $\tilde{\E}(y)$ as 
$$\tilde{\E}(y)=\tilde{\E}(\omega_0) + \int_{\omega_0}^y \parent{\frac{b \bar{M} \expfs{b/\omega}}{\parent{\expfs{b/\omega}-1}^2 \omega^2}-c(\omega)}d(\omega)$$
The first term on the right-hand side of the previous equation is constant. Now we move on to proving that the integrand is increasing with $y$ and it is positive if $y$ is large enough. Denote by
\begin{equation}\label{eq:defp}
p(y,b)=\frac{b \bar{M} \expfs{b/y}}{\parent{\expfs{b/y}-1}^2 y^2}
\end{equation}
Taking into account that $c$ is integrable, we can compute the limit of the integrand in $\infty$, obtaining
\begin{equation*}
\lim_{y \to +\infty}p(y,b)-c(y)=\frac{\bar{M}}{b}>0
\end{equation*} 
Let us prove that $p(y,b)$ is increasing with $y$ for all $b>0$, which renders $p(y,b)-c(y)$ also increasing with $y$. The derivative of $p$ with respect to $y$ is given by
$$\partial_{y} p(y,b)=\frac{b\bar{M}\expfs{b/y}\parent{(b/y)\expfs{b/y}-2\expfs{b/y}+b/y+2}}{y^3\parent{\expfs{b/y}-1}^3}$$
in which the denominator and the first factor in the numerator are clearly positive. To prove that the remainder factor is also positive, observe that 
$x\expfs{x}-2\expfs{x}+x+2>0$ if $x>0$.

\end{proof}
\end{proposition}

\subsection{Explicit error bounds}
\label{ss:bounds}
In the particular case when either $\sigma^2>0$ or 
$\cl>0$ for some $\lambda \in (0,2)$ we can give an explicit version 
of \eqref{eq:fullErrorGeneral}. Substituting $M$ by $\tilde{M}$ defined in Theorem \ref{teo:boundM} and $c$ by the function given in \eqref{eq:cutoffparticularSimpler} we obtain
\begin{align}%\tag{\ref{eq:fullErrorGeneral}*}
\label{eq:fullErrorExpression}
\est &= \estQuadrature + \estFrequency
\end{align}
where
\begin{align}
\label{eq:boundQ} \estQuadrature &= 
\ssum{c\in \sset{-1,1}}{} 
\frac{
\expfs{\damp x}
\expfs{cax}
\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}}{\pi \parent{ \expfs{\frac{2 \pi a}{\step}} -1}}
\int_{\R}  \expfs{-\tau \left( \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}\right) } \d \omega
\\[.5em]
\label{eq:boundF}
\estFrequency &= \frac{\expfs{\damp x}}{\pi}
\norminf{\ft{\ga}}{\R}
\expfs{\tau \charExp{\damp}} \int_{n\step}^{\infty} 
\expfs{-\tau \parent{\frac{\sigma^2}{2} \omega^2+\frac{|\omega|^{2 - \lambda}}{4} \cl \indicator_{\absval{\omega}>1}} } d\omega
\end{align}
This reproduces the essential features of Theorem 6.6
in \cite{fenglinetsky},
the bound \eqref{eq:boundF} can be further improved by substituting $\norminf{\ft{\ga}}{\R}$ by $\norminf{\ft{\ga}}{[n\step,\infty)}$.

\begin{remark}\label{rem:decuople}
Observe that the bound of both the quadrature and the cutoff error is given
by a product of one factor that depends exclusively on the payoff and 
another factor that depends on the asset dynamic. This property makes it easy to evaluate
the bound for a specific option under different dynamics of the asset price.
In Subsection \ref{subsec:call} we analyse the terms
that depend on the payoff function for the particular case of call options.
\end{remark}

\begin{remark}\label{rem:asynonasy}
From \eqref{eq:boundQ} it is evident that the speed of the exponential convergence
of the trapezoidal rule for analytic functions is dictated by the width of the strip in which
the function being transformed is analytic. Thus, in the limit of small error tolerances,
it is desirable to set $a$ as large as possible to obtain optimal rates.
However, non-asymptotic error tolerances are often practically relevant and in these
cases the tradeoff between optimal rates and the constant term $\absval{\ft{\ga}}$
becomes non-trivial. As an example, for the particular case of the Merton model,
we have that any finite value of $a$ will do. However, this improvement of the rate of
spectral convergence is more than compensated for by the divergence in the constant term.
\end{remark}

The integrals 
in \eqref{eq:boundQ} and \eqref{eq:boundF} can, in some cases, be computed analytically,
or bounded from above by a closed form expression.
Consider for instance dissipative models with finite jump intensity. 
These models are characterised by $\sigma^2>0$ and $\cl=0$. Thus the integrals can be expressed in terms of the cumulative normal distribution $\Phi$:
\begin{align}
\int_{\mathbb R} \expfs{-\tau \frac{\sigma^2 \omega^2}{2}}d\omega&=\sqrt{\frac{2 \pi}{\tau \sigma^2}},
\\
\int_{\varsigma}^\infty \expfs{-\tau \frac{\sigma^2 \omega^2}{2}} d \omega &= 
\sqrt{\frac{2 \pi}{\tau \sigma^2}}
\parent{1-\Phi \parent{\varsigma \sqrt{\tau \sigma^2}}}
\end{align}

Now we consider the case of pure-jump processes (ie, $\sigma^2 =0$) that satisfy the condition $\cl > 0$ for some $\lambda \in (0,2)$. 
In this case the integrals are expressible
in terms of the incomplete gamma function $\gamma$.
First, let us define the auxiliary integral:
\begin{align}
%%% Including the full derivation, remove intermediate stages if too dull.
I \parent{a,b} 
&\equiv
\expfs{-a} + a^{-\frac{1}{b}} \gamma \parent{\frac{1}{b},a}
\nonumber
\end{align}
for $a,b>0$. Using this, the 
%constant coefficients
integrals become:
\begin{equation}
\int_{\R} \expfs{- \tau \frac{\absval{\omega}^{2-\lambda}}{4} \cl \mathbf{1}_{\absval{\omega}>1}}
= 2 \parent{1+ I \parent{\frac{\tau \cl}{4},2-\lambda}}
\end{equation}

\begin{equation}
\int_{\varsigma}^\infty \expfs{- \tau \frac{\absval{\omega}^{2-\lambda}}{4} \cl \mathbf{1}_{\absval{\omega}>1}}
= \begin{cases}
I \parent{\frac{\tau \cl}{4},2-\lambda} + 1-\varsigma & \quad \varsigma <1\\
\varsigma I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda} &\quad \varsigma \geq 1
\end{cases}
\end{equation}
An example of a process for which the previous analysis works
is the CGMY model presented in \cite{cgmy_fine,carr2003stochastic}, for the regime $\cgmyY > 0$.

Lastly, when both $\cl$ and $\sigma^2$ are positive, 
the integrals in \eqref{eq:boundQ} and \eqref{eq:boundF} can be bounded by a simpler expression. 
Consider the two following auxiliary bounds for the same integral, in which $\varsigma\geq 1$:
\begin{align}
\int_{\varsigma}^\infty \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl} } d\omega
& \leq \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2} \int_{\varsigma}^\infty \expfs{-\tau \frac{|\omega|^{2 - \lambda}}{4} \cl} d\omega \\
& = \varsigma \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2}  I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda} \notag
\end{align}

\begin{align}
\int_{\varsigma}^\infty \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl} } d\omega
& \leq \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \int_{\varsigma}^\infty \expfs{-\tau \frac{\sigma^2}{2}\omega^2} d\omega \\
& =\sqrt{\frac{2 \pi}{\tau \sigma^2}} \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \parent{1-\Phi(\varsigma\sqrt{\tau \sigma^2})}\notag
\end{align}
We have that $b(\varsigma)$, defined as the minimum of the right hand sides of the two previous equations,
\begin{align*}
b(\varsigma)=\min & \left\{ \varsigma \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2}  I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda}, \sqrt{\frac{2 \pi}{\tau \sigma^2}} \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \parent{1-\Phi\parent{\varsigma\sqrt{\tau \sigma^2}}} \right\}
\end{align*}
is a bound for the integral. Bearing this in mind we have
%approximation:
\begin{align}
\int_{\R} \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}} } d\omega
& \leq 2 \Phi\parent{\sqrt{\tau\sigma^2}}-1 + 2 b(1)
\end{align}
and
\begin{align}
\int_{\varsigma}^{\infty} \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}} } d\omega
\leq & b\parent{\varsigma}
\end{align}
provided that $\varsigma\geq 1$.

\section{Computation and minimization of the bound}
\label{section:Numerics}

In this section, we present numerical examples on the
bound presented in the previous section using practical
models known from the literature.
We gauge the tightness of the bound compared to the
true error using both dissipative and pure-jump processes.
We also demonstrate the feasibility of using the expression of the bound
as a tool for choosing numerical parameters
for the Fourier inversion.

\subsection{Call option in variance gamma model}

The variance gamma model provides a test case
to evaluate the bound in the pure-jump setting. We note
that of the two numerical examples presented, it is the less regular one
in the sense that $\sigma^2=0$ and $C \parent{\lambda} = 0$ for $0 < \lambda < 2$,
indicating that Theorem \ref{teo:boundM} in particular is not applicable. 
%\marginpar{Do we really need this sentence? If so, what does it mean?
%$\exists ~ \tau^*>0: \charFun{\tau}{\cdot} \notin L_1\parent{\R} ~ \forall \tau<\tau^*$.
%It also is a case in which the characteristic function is not integrable for sufficiently short maturities, making approximation of the kind presented in (28) impossible. 
% }

The \levy measure of the VG model is given by:
\begin{align*} \label{eq:VG}
\nu_{\mathrm{VG}} \parent{\d y} 
= \d y \parent{
\mathbf{1}_{y>0} \frac{\cgmyC \expfs{-\cgmyM y}}{y}
- \mathbf{1}_{y<0} \frac{\cgmyC \expfs{\cgmyG y}}{y}
}
\end{align*}
and the corresponding characteristic function is given by
eq. (7) of \cite{madan1998variance}:
\begin{align*}
\charFun \tau \omega &= 
\parent{1 - i \theta \chi \omega + \frac{\sigma^2 \chi}{2}}^{-\frac{\tau}{\chi}}
\\
\cgmyC &= \chi^{-1}
\\
\cgmyG &= \parent{\sqrt{\frac{\theta^2 \chi^2}{4}+\frac{\sigma^2 \nu}{2}} -\frac{\theta \chi}{2}}^{-1}
\\
\cgmyM &= \parent{\sqrt{\frac{\theta^2 \chi^2}{4}+\frac{\sigma^2 \nu}{2}} +\frac{\theta \chi}{2}}^{-1}
\end{align*}

By Proposition \ref{prop:analy1} 
we get that 
\begin{align} 
a < \mathrm{min} \sset{\cgmyG-\damp,\cgmyM+\damp}
\end{align}
which, combined with the requirement that
$\ga \in L^1 \parent{\R}$ (cf, Remark \ref{remark:callPut}), implies:
\begin{align}
a &< \mathrm{min} \sset{\cgmyM-\damp,\cgmyG+\damp,\damp-1}
\\
a &< \mathrm{min} \sset{\cgmyM-\damp,\cgmyG+\damp,-\damp}
\nonumber
\end{align}
for calls and puts, respectively. We note that evaluation
of the integral in \eqref{eq:fadefinition} is possible also for $\damp \in (0,1)$
and for $\damp < 0$. In fact, there is a correspondence between shifts in the
integration countour and put-call parity. Integrals with $\damp<0$ give rise to put
option prices instead of calls. For an extended discussion of this, we refer to
\cite{lee2004option} or \cite{boyarchenko2011new}, in which 
conformal deformation of the
integration contour is exploited in order to achieve improved numerical accuracy.

In \cite{lee2004option} and in our calculations the parameters equal
$\cgmyM = 39.7840$, $\cgmyG = 20.2648$ and $\cgmyC=5.9311$.

Table \ref{tb:usvslee} presents the specific parameters and compares
the bound for the VG model with the results obtained by \cite{lee2004option}.
Based on the table, we note that for the VG model
presented in \cite{madan1998variance} we can achieve comparable or better error bounds when compared to the study
by Lee.

To evaluate the bound, we perform the integration of 
\eqref{eq:defM} and \eqref{eq:truncationErrorIntegral}
by relying on the Clenshaw-Curtis quadrature method
provided in the SciPy package.
To supplement Table \ref{tb:usvslee} for a wide range
of $n$,
we present the magnitude of the bound compared to the true error in Figure
\ref{fig:vgerrvsbound}.

In Figure \ref{fig:vgerrvsbound}, we see that the choice of numerical parameters
for the Fourier inversion has a strong influence on the error of the numerical method.
One does not in general have access to the true solution. 
Thus, the parameters need to be optimised 
with respect to the bound.
Recall that $\Error = \Error(\alpha,\Delta \omega,a, n)$ and 
$\est = \est(\alpha,\Delta \omega, n)$ denote the true and estimated
errors, respectively. Keeping the number of quadrature points $n$ fixed,
we let $\parent{\alpha_1,\Delta \omega_1,a_1}$ and $\parent{\alpha_2,\Delta \omega_2}$
denote the minimisers of the estimated and true errors, respectively
\begin{align}
\parent{\alpha_1,\Delta \omega_1,a_1}  &= \mathrm{arg ~ inf}~ \est 
\label{eq:pdef1}
\\
\parent{\alpha_2,\Delta \omega_2}  &= \mathrm{arg ~ inf} ~ \Error 
\label{eq:pdef2}
\end{align}
We further let $\mathcal{E}_1$ and $\mathcal{E}_2$ 
denote the true error as a function of the parameters
minimising the estimated and the true error, respectively
\begin{align}
\mathcal E_1 &= \Error \parent{\damp_1,\step_1} 
\label{eq:pdef3}
\\
\mathcal E_2 &= \Error \parent{\damp_2,\step_2}  
\label{eq:pdef4}
\end{align}
In Figure \ref{fig:vgerrvsbound} we see that the true error increases by approximately an 
order of magnitude when optimising to the bound instead of to the true error,
translating into a two-fold difference
in the number of quadrature points needed for a given tolerance. 
The difference between ${\Error}_1$ 
and the bound is approximately another order of magnitude and necessitates
another two-fold number of quadrature points compared to the theoretical minimum.

In Figure \ref{fig:parameterPlot}, we present the true error
\footnote{The reference value to compute the true error was
obtained by the numerical methods with $n$ and $\step$ such
that the level of accuracy is of the order $10^{-10}$.} for the Fourier method
for the two test cases in Table \ref{tb:usvslee}. 
 We note that while minimising error
bounds will produce sub-optimal results, the numerical parameters that minimise
the bound are a good approximation of the true optimal parameters. 
This, of course,
is a consequence of the error bound having qualitatively similar behaviour as
the true error, especially as one gets further away from the true optimal parameters.

\begin{remark}
In practice, the Hardy norm in coefficient $M$ reduces to evaluating an
$L^1$ norm along the two boundaries of the strip of width $2a$.
We find that, for practical purposes, the performance of the
Clenshaw-Curtis quadrature of the
QUADPACK library provided by SciPy library is more than adequate,
enabling the evaluation of the bound in a fraction of a second.

For example, the evaluations of the bounds in Table
\ref{tb:usvslee} take only around $0.3$ seconds to evaluate
on Mid 2014 Macbook Pro equipped with a
2.6 GHz Intel Core i5 processor, this without attempting to optimize
or parallelize the implementation and while checking for input sanity
factors such as the evaluation of the characteristic function in a domain
that is a subset in the permitted strip.

We believe that through optimizing routines, skipping sanity checks for inputs
and using a lower-level computation routines this can be optimized even further,
guaranteeing a fast performance even when numerous evaluations are needed.

\end{remark}

\begin{table}
\begin{center}\begin{tabular}{ccccccc}
\hline
 & $K$ & $80$ & $90$ & $100$ & $110$ & $120$\\
\hline
\multirow{2}{*}{$12 \tau=1$} & $\damp$ & $ -16.9 $ & $ -13.8 $ & $ 21.6 $ & $ 29.10 $ & $ 36.3 $ \\ 
 & $a$ & $ 3.33 $ & $ 6.45 $ & $ 18.1 $ & $ 9.77 $ & $ 3.52 $ \\ 
\multirow{2}{*}{$N =32$} & $\cutoff$ & $ 229 $ & $ 229 $ & $ 363 $ & $ 363 $ & $ 424 $ \\ 
 & $\overline{\mathcal E}$ & $ 3.35 \times 10^{-4} $ & $ 0.00334 $ & $ 0.00562 $ & $ 3.97 \times 10^{-4} $ & $ 7.33 \times 10^{-6} $ \\ 
 & $\overline{\mathcal E}^*$ & $ 6 \times 10^{-4} $ & $ 0.0032 $ & $ 0.0058 $ & $ 6 \times 10^{-4} $ & $ 1 \times 10^{-4} $ \\ 
\hline
\multirow{2}{*}{$12 \tau=4$} & $\damp$ & $ -13.8 $ & $ -13.8 $ & $ 22.1 $ & $ 23.7 $ & $ 29.10 $ \\ 
 & $a$ & $ 6.11 $ & $ 6.11 $ & $ 17.9 $ & $ 15.2 $ & $ 8.75 $ \\ 
\multirow{2}{*}{$N =8$} & $\cutoff$ & $ 62.4 $ & $ 42.4 $ & $ 84.9 $ & $ 126 $ & $ 126 $ \\ 
 & $\overline{\mathcal E}$ & $ 3.99 \times 10^{-4} $ & $ 0.00312 $ & $ 0.00398 $ & $ 3.57 \times 10^{-4} $ & $ 1.33 \times 10^{-5} $ \\ 
& $\overline{\mathcal E}^*$ & $ 1.3 \times 10^{-3} $ & $ 0.0057 $ & $ 0.0055 $ & $ 9 \times 10^{-4} $ & $ 1 \times 10^{-4} $ \\

\hline
\end{tabular}\end{center}
\caption{\label{tb:usvslee}The error bound for European call/put options in the VG model for select examples.
Refererence result $\overline{\mathcal E}^*$ from \cite{lee2004option}}
\end{table}

\begin{remark}
Like many other authors, we note the exceptional \emph{guaranteed} accuracy of the FT-method,
with only dozens of quadrature points. This is partially a result of the regularity of the European option price.
Numerous Fourier-based methods have been developed for pricing path-dependent options.
One might, for the sake of generality of implementation, 
be tempted to use these methods for European options as well, correcting for the lack of early exercise
opportunities. This certainly can be done, but due to the
weakened regularity, the required numbers of quadrature points are easily in the thousands, even when no rigorous
bound for the error is required.

We raise one point of comparison,
the the European option pricing example in Table 2 of \cite{jackson2008fourier},
which indicates a number of quadrature points for pricing the option in the range of thousands.
With the method introduced, to guarantee $\est \approx 10^{-3}$,
even with no optimisation, $n=64$ turns out to be sufficient.
\end{remark}

\subsection{Call options under Kou dynamics}

\begin{table}
\begin{center}\begin{tabular}{ccccccc}
\hline
 & $K$ & $80$ & $90$ & $100$ & $110$ & $120$\\
\hline
 & $\overline{\mathcal E}$ & $ 2.67 \times 10^{-4} $ & $  3.49 \times 10^{-4} $ & $  4.43 \times 10^{-4} $ & $ 5.52 \times 10^{-4} $ & $ 6.77 \times 10^{-4} $ \\ 
  & $\damp$ & $ -1.57 $ & $  -1.57 $ & $  -1.57$ & $ -1.57 $ & $ -1.57$ \\ 
  & $\cutoff$ & $ 22.9 $ & $  22.8 $ & $  22.6$ & $ 22.5 $ & $22.4$ \\ 
\hline
  &$\overline{\mathcal E}^*$   & $ 0.34 $ & $  0.26 $ & $  0.21$ & $ 0.17 $ & $0.13$ \\ 
 & $\overline{\mathcal E}^{\dagger}$  & $ 6.87 \times 10^{-4} $ & $  1.90 \times 10^{-3} $ & $  2.82 \times 10^{-3} $ & $ 2.72 \times 10^{-3} $ & $ 2.29 \times 10^{-3} $ \\ 
 \hline
\end{tabular}
\end{center}
\caption{\label{tb:usvsleekou}
Numerical performance of the bound for the Kou model, with the test case in
\cite{toivanen2007numerical} (see also \cite{d2005robust}) with the number
of quadrature points set to $n=32$.
The point of comparison $\overline{\mathcal E}^*$ refers to the corresponding bound computed
with the method described in Chapters $6.1$ to $6.4$ of \cite{lee2004option}.
In the $\overline{\mathcal E}^{\dagger}$, the cutoff error has been evaluated using a 
computationally more intensive Clenshaw-Curtis quadrature instead of an asymptotic argument
with an exponentially decaying upper bound for the option price.
}
\end{table}

For contrast with the pure-jump process presented above, we also test the performance
of the bound for Kou model and present relevant results in \ref{tb:usvsleekou}. This model differs from the first example not only by being dissipative
but also in regularity, in the sense that the maximal width of the domain $A_a$ is, for the 
case at hand, considerably narrower. The \levy measure in the Kou model is given by
\begin{align*}
\nu_{\mathrm{Kou}} \parent{\d y} = \lambda \parent{ p \expfs{-\eta_1 y} \mathbf{1}_{y>0} + q \expfs{\eta_2 y} \mathbf{1}_{y<0} }
\end{align*}
with $p+q=1$.
For the characterisation given in \cite{toivanen2007numerical} the
values are set as
\begin{align*}
\lambda = 0.1, ~~ r = 0.05 ~~ \tau = 0.25 ~~ S_0 = 100 \\
p = 0.3445, ~~ \eta_1 = 3.0465 ~~ \eta_2 = 3.0775
\end{align*}
from the expression of the characteristic exponent (see  \cite{kou2004option})
\begin{align*}
\charExp{z} = z \parent{r -\frac{\sigma^2}{2} - \lambda \zeta} + \frac{z^2 \sigma^2}{2} 
+ \lambda \parent{\frac{p \eta_1}{\eta_1-z} + \frac{q \eta_2}{\eta_2 +z} -1}
\end{align*}
it is straightforward to see
\begin{align*}
A_a \subset \sset{z \in \C : \mathrm{Im} z \in ( -3.0465, 3.0775)}
\end{align*}
This range is considerably narrower than that considered earlier.
In the case of transforming the option prices in strike space, the
relevant expressions for option prices as well as the
error bounds contain a factor exponential in $k$. The practical implication
of this is that for deep out of the money calls, it is often beneficial to exploit
the put-call-parity and to compute deep in the money calls. However, in the case
at hand, the strip width does not permit such luxury. As a consequence,
the parameters that minimize the bound are near-identical through a wide range of moneyness,
suggesting use of FFT algorithm to evaluate the option prices at once for a range of strikes.

\subsection{Binary option in the Merton model}
For the particular case of Merton model, the \levy measure is given by
\begin{align*}
\nu_{\mathrm{Merton}} \parent{\d y}
=
\frac{\lambda}{\sqrt{2 \pi \sigma^2}} \expf{-\frac{\parent{y-r_j}^2}{2\sigma_j^2}}
\end{align*}
and the characteristic exponent correspondingly by
\begin{align*}
\charExp{z}_{\mathrm{Merton}}
=
z \parent{r-\frac{\sigma^2 z}{2}}
+ \frac{\sigma^2 z^2}{2}
+ \lambda
\parent{\expfs{z r_j + \frac{\sigma_j^2 z^2}{2}}-1 - z \parent{\expfs{r_j +\frac{\sigma_j^2}{2} }-1} }
\end{align*}
we may employ a fast semi-closed form evaluation
of the relevant integrals instead of resorting to quadrature methods.
We choose the Merton model as an example of bounding the error of the numerical method for such
a model. The parameters are adopted from 
the estimated parameters for S\&P 500 Index from 
\cite{andersen2000jump}:
\begin{align*}
S_0 = 100,
~~~
\lambda = 0.089,
~~
\sigma = 0.1765,
~~~
r = 0.05,
~~
r_j = -0.8898,
~~~
\sigma_j = 0.4505
\end{align*}

In Figure \ref{fig:mertonPlot}, we present the bound and true error for the Merton model
to demonstrate the bound on another dissipative model. The option presented is a binary option
with finite support on $[95,105]$; no damping was needed or used. We note that
like in the case of the pure-jump module presented above, our bound reproduces the
qualitative behaviour of the true error. The configuration resulting from optimising the bound
is a good approximation of the true error. Such behaviour is consistent through the range of
$n$ of the most practical relevance.

\subsection{Call options}
\label{subsec:call}
In Subsection \ref{ss:bounds} explicit expressions to bound $\Error$ are provided. 
To evaluate these bounds it is necessary to compute $\norminf{\ft{\ga}}{\R}$ and $\norminf{\ft{\ga}}{\strip{a}}$. According to Remark \ref{rem:decuople}, once
we compute these values we could use them for any model, provided that 
they satisfy the conditions considered there.

The payoff of perhaps the most practical relevance is that of
a call option. 
Consider $g$ defined by:
\begin{align*}
 g\parent{x} = \parent{S_0 \expfs{x}-K}^+ = S_0 \parent{\expfs{x}-\expfs{k}}^+
\end{align*}
for which a selection of a damping parameter $\damp>0$ is necessary
to have the damped payoff in $L^1 \parent{\R}$ and to ensure the existence of a
Fourier transformation. In this case we have% \cite{lordfourier}:
\begin{align}
\hat{\ga}
%\fcallga
\parent{\omega} &= S_0 \int_\R \expf{ \parent{1-\damp + i \omega}x} - \expf{k+\parent{i \omega- \damp}x}
\d x
\\
 %&= S_0 \evaldiff{ 
 %\frac{\expf{\parent{1-\damp + i \omega}x}}{1+ i \omega - \damp}
%-
 %\frac{\expf{k + \parent{i \omega - \damp}x}}{i \omega - \damp}
   %}{k}{\infty}
   %\\
   &=
   \frac{ S_0 \expf{\parent{1- \damp + i \omega} k } }{\parent{1 + i \omega- \damp} \parent{i \omega - \damp}}
%\label{eq:dampedghatcall}
%%%TODO: check!!!!
\end{align}
and
\begin{align}
 \absval{\ft{\ga}(\omega) }^2
 =
 \frac{S_0^2 \expfs{2 \parent{1-\alpha}k}}{\parent{\alpha^2 + \omega^2 } \parent{\parent{1-\alpha}^2 + \omega^2}}
\end{align}
It is easy to see that the previous expression decreases as $|\omega|$ increase. 
This yields
\begin{align}
\norminf{\ft{\ga}}{\R}
=\absval{\ft{\ga}(0) } =
\frac{S_0 \expfs{\parent{1-\alpha}k}}{\alpha^2-\alpha}
\end{align} 
and 
\begin{align}
\norminf{\ft{\ga}}{[\varsigma,\infty)}
=\absval{\ft{\ga}(\varsigma) }
\end{align} 
The maximisation of $\absval{\ft \ga}$ in the strip $\strip{a}$ of the complex plane is more subtle.
%illustration of the magnitude of the term is plotted in Figure \ref{fig:gHatSup}. 
Denoting $\ft \ga(\eta,\rho)=\ft \ga(\eta + i \rho)$, we look for critical points that
satisfy 
% Local extrema of $\absval{\ft \ga}$, denoting $\omega = \eta + i \rho$ are characterised
$\partial_{\eta} \absval{\ft \ga}=0$. 
This gives
\begin{align}
4 \eta^3 + 2 \eta \parent{4 \rho \alpha + 2 \alpha^2 - 2 \rho -2 \alpha + \rho^2 +1}=0.
\end{align}
For $\rho$ fixed, $\absval{\ft \ga}$ has a vanishing derivative with respect to $\eta$
at a maximum of three points. Of the three roots of the derivative, only the one characterised
by $\eta=0$ is a local maximum, giving us that for call options
\begin{align}
\norminf{\ft{\ga}}{\strip{a}} = \sup_{\rho \in [-a,a]} \absval{\ft \ga(0,\rho)} 
\label{eq:complexsup}
\end{align}
Now, observe that $\absval{\ft \ga(0,\rho)}$ is a differentiable real function of $\rho$, whose derivative is given by the following polynomial of second degree:
\begin{align}
 p \parent{\rho} \equiv k \parent{\rho +\alpha - 2 \rho \alpha - \alpha^2 - \rho^2}
 - 2 \alpha - 2 \rho +1
\end{align}
We conclude that
\begin{align}
\norminf{\ft{\ga}}{\strip{a}} = \max_{\rho \in B} \left\{ \absval{\ft \ga(0,\rho)}\right\} 
\label{eq:complexsup2}
\end{align}
where $B$ is the set of no more than four elements consisting of $a$; $-a$; and the real roots of $p$ that fall in $(-a,a)$.

\begin{remark}
So far, we have assumed the number of quadrature points $n$ to be constant.
In real life applications, however, this is often not the case. Typically the user might want to choose a minimal
$n$ that is sufficient to guarantee error that lies within a pre-defined error tolerance.

In such a case, we propose the following, very simplistic scheme for optimising numerical parameters
and choosing the appropriate $n$ to satisfy error smaller than $\epsilon$:
\begin{enumerate}
\item Select $n=n_0$ and optimize to find the relevant configuration
\item See if $\EQ + \EF <\epsilon$, if not, increase $n$ by choosing it from a pre-determined
increasing sequence $n=n_j$ and repeat procedure.
\end{enumerate}
Especially in using FFT algorithms to evaluate the Fourier transforms,
we propose $n_j = 2^j n_0$. We further note that typically the optimal configuration
for the optimizing configuration for $n_{j+1}$ quadrature points does not differ too
dramatically from the configuration that optimizes bounds for $n_j$.
\end{remark}

\section{Conclusion}
\label{section:Conclusion}

We have presented a decomposition of the error committed in
the numerical evaluation of the inverse Fourier transform needed
in asset pricing for exponential \levy models into truncation and
quadrature errors. For a wide class of exponential \levy models, we
have presented an $L^\infty$-bound for the error.

The error bound differs from the earlier work presented in \cite{lee2004option}
in the sense that it does not rely on the asymptotic behaviour of the option
payoff at extreme strikes or option prices, allowing pricing a wide variety
of non-standard payoff functions such as the ones in \cite{suh2008class}.
The bound, however, does not take into account path-dependent options. We
argue that the error for the methods that allow evaluating american, bermudan,
or knockoff options are considerably more cumbersome and produce significantly
larger errors so that in implementations where performance is important, such as
calibration, using American option pricing methods for European options is not justified.

The bound also provides a general framework in which
the truncation error is evaluated using a quadrature method that remains invariant
regardless of the asymptotic behaviour of the option price function.
The structure of the bound allows for
a modular implementation that decomposes the error components arising from
the dynamics of the system and the payoff into a product form for a large
class of models, including all dissipative models. On select examples, we
also demonstrate the performance that is comparable or superior to the
relevant points of comparison.

We have focused on the minimization of the bound as a proxy for minimizing numerical error.
Doing this, one obtains, for a given parametrization of a model, a rigorous $L^\infty$
bound for the error committed in solving the European option price. We have shown that the 
bound reproduces the qualitative behaviour of the actual error. This supports the argument for
selecting numerical parameters in a way that minimizes the bound, giving evidence that
this selection will, besides guaranteeing numerical precision, be close to the actual minimizing
configuration that is not often achievable at an acceptable computational cost.

The bound can be used in the primitive setting of establishing a strict error
bound for the numerical estimation of option prices for a given set of physical and
numerical parameters or as a part of a numerical
scheme, whereby the end user wishes to estimate an option price either on
a single point or in a domain up to a predetermined error tolerance.

In the future, the error bounds presented can be used in efforts requiring
multiple evaluations of Fourier transformations. Examples of such applications
include multi-dimensional Fourier transformations, possibly in sparse tensor grids,
as well as time-stepping algorithms for American and Bermudan options.
Such applications are sensitive towards the error bound
being used, as any numerical scheme will be required to run multiple times, either
in high dimension or for multiple time steps (or both).

\section{Acknowledgements}

H\"app\"ol\"a, Crocce are members and Tempone is the director KAUST Strategic Research Initiative
for Uncertainty quantification.

We wish to thank an anonymous referee for their critique that significantly improved the quality of the manuscript.

\section{Declarations of Interest}

The authors report no conflicts of interest. The authors alone are responsible for the content and writing of the paper.

\bibliographystyle{agsm}
\bibliography{references}

\appendix

\section{Truncation bound in Lee's scheme}

For the strike-space transformed option price we have
\begin{align*}
\tilde \fa \parent{\omega}
=
\frac{\charFun 1 {\omega - i \parent{\damp + 1}}}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
\end{align*}
In the particular case of the Kou model, we have 
\begin{align*}
\charFun 1 \omega
=
\expfs{\tau
\parent{
i \omega \parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
- \frac{\omega^2 \sigma^2}{2}
+ \lambda
\parent{
\frac{p \eta_1 }{\eta_1 - i \omega}
+\frac{q \eta_2}{\eta_2 + i \omega}
-1
}
}
}
\end{align*}
Thus,
\begin{align*}
\absval{ \charFun 1\omega}
=
\expfs{
\tau
\parent{ 
\parent{\damp+1}
\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
+
\frac{\parent{\damp+1}^2 - \omega^2}{2}
}
}
\expfs{ \tau \lambda \mathrm{Re} \parent{
\frac{p \eta_1}{\eta_1-\damp-1 - i \omega}
\frac{q \eta_2}{\eta_2 +\damp +1 + i \omega}
-1
 }}
\end{align*}
We bound the first jump term resulting from upward jumps as
\begin{align*}
\frac{p \eta_1}{ \eta_1 - \damp -1 - i \omega}
= \frac{p \eta_1 \parent{\eta_1 - \damp -1 +i \omega}}{\parent{\eta_1 - \damp -1}^2 + \omega^2}
\end{align*}
from which it follows
\begin{align*}
\mathrm{Re}
\parent{
\frac{p \eta_1}{ \eta_1 - \damp -1 - i \omega}
}
\leq
\frac{p \eta_1 \parent{\eta_1 - \damp -1 }}{\parent{\eta_1 - \damp -1}^2 }
\end{align*}
similar inequality holds for the term resulting from downward jumps giving us
that
\begin{align*}
\absval{ \charFun 1 \omega}
\leq 
\expfs{
\tau
\parent{ 
\parent{\damp+1}
\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
+
\frac{ \sigma^2 \parent{\damp+1}^2 - \sigma^2 \omega^2}{2}
}
}
\expfs{
\tau \lambda
\parent{
\frac{p \eta_1 \parent{\eta_1 - \damp -1}}{ \parent{\eta_1 -\damp -1}^2}
+
\frac{q \eta_2 \parent{\eta_2 + \damp +1}}{\parent{\eta_2+ \damp +1}^2}
+1
}
}
\end{align*}
The term $\expf{- \frac{\omega^2}{2}}$ can be bounded from above
by an exponential term $\expf{-\gamma \omega}$:
\begin{align}
\expf{- \frac{\sigma^2 \omega^2}{2}}
\label{eq:tailBound}
& \leq
\expf{-\gamma \omega+ K}
\\
- \sigma^2 \omega^2 + 2 \gamma \omega - 2 K & \leq 0
\nonumber
\end{align}
the condition for the quadratic to have no solutions
is given by
\begin{align*}
 \gamma^2 - 2 K  \sigma^2 \leq 0,
\end{align*}
setting $K =\frac{1}{2}$, $\gamma = \sigma$
giving us the bound in eq. \eqref{eq:tailBound}.

This puts us in place to apply theorem 6.1 of \cite{lee2004option}, with
\begin{align*}
\Phi \parent{u} &= 
\frac{
\expf{
\tau
\parent{ 
\parent{\damp+1}
\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
+
\frac{ \sigma^2 \parent{\damp+1}^2 }{2}
}
}
\expfs{
\tau \lambda
\parent{
\frac{p \eta_1 \parent{\eta_1 - \damp -1}}{ \parent{\eta_1 -\damp -1}^2}
+
\frac{q \eta_2 \parent{\eta_2 + \damp +1}}{\parent{\eta_2+ \damp +1}^2}
+1
}
}}
{\damp^2 \parent{\damp+1}^2}
\\
\gamma & = \sigma
\end{align*}

\end{document}

### 2016-03-27 tests/HJM/echo_test_cmd.py

@@ -46,14 +46,14 @@ args = parser.parse_known_args()[0]
 # Note that 0.6931 is np.log(2)
 base = "run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
 -qoi_seed {seed} \
--mimc_dim 1 -mimc_w 1 -mimc_s 1 -mimc_gamma 2 -mimc_beta 2 \
+-mimc_dim 2 -mimc_w 2 2 -mimc_s 0.5 1 -mimc_gamma 1 1 -mimc_beta 2 2 \
 -mimc_bayesian {bayesian} "
 
 cmd_multi = "python " + base + "-mimc_verbose False -db True -db_tag {tag} " + " -db_host {} ".format(args.db_host)
 cmd_single = "python " + base + " -mimc_verbose True -db False "
 
 if not args.multi:
-    print(cmd_single.format(seed=0, bayesian=False, TOL=0.0001))
+    print(cmd_single.format(seed=0, bayesian=False, TOL=args.TOL_min))
 else:
     realizations = args.N_iter # number of runs for each tolerance
     TOL_max = args.TOL_max
@@ -62,5 +62,5 @@ else:
     TOLs = np.logspace(np.log10(TOL_max),np.log10(TOL_min),TOL_N)
     for TOL in TOLs:
         for i in range(0, realizations):
-            print cmd_multi.format(bayesian=False, tag="HJM_Example_2_newrate", TOL=TOL,
+            print cmd_multi.format(bayesian=False, tag="HJM_Example_2_newnewrate", TOL=TOL,
                                    seed=np.random.randint(2**32-1))

### 2016-03-27 tests/HJM/run.py

@@ -77,7 +77,7 @@ def main():
     return mimcRun.data.calcEg()
 
 def mySampleQoI(run, inds):
-    return HJM.hoLeeExample3(inds)
+    return HJM.hoLeeExample2(inds)
 
 
 if __name__ == "__main__":

### 2016-03-27 tests/HJM/manuscript.tex
still compiling
and still working
now the friggin bjork paper works!
working with a reasonable text

@@
 
-\section{Introduction}
-\label{section:Introduction}
+We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models.
+The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.
 
-%\subsection{Levy processes and finances}
-\levy processes form a rich field within mathematical
-finance. They allow modelling of asset prices with possibly discontinuous dynamics. An early and probably the best known model involving
-a \levy process is
-the \cite{Merton1976125} model, which generalises the \cite{black1973pricing} model. More recently, we have seen more complex models allowing
-for more general dynamics of the asset price. Examples of such models include
-the \cite{kou2002jump} model (see also \cite{Dotsis20073584}), the Normal Inverse Gaussian
-model (\cite{barndorff1997normal, rydberg1997normal}),
-the Variance Gamma model (\cite{madan1990variance, madan1998variance}),
-and the Carr-Geman-Madan-Yor (CGMY) model
-(\cite{cgmy_fine,carr2003stochastic}). For a good exposition on jump processes in finance we refer to \cite{conttankov}
-(also see \cite{raible2000levy} and \cite{eberlein2001application}).
+We are focused on the evolution of the forward curve $f \parent{t , \tau}$ for $t \in [0,t_{max}]$, $\tau \in [0,\tau_{max}]$ with $\tau_{max}>t_{max}$.
+In order to define a mesh for the numerical solution of the problem, let us define $\bar \ell \in \mathbb{N}^3$.
 
-%\subsection{European option PIDE}
-Prices of European options whose underlying asset is driven
-by the \levy process are solutions to partial integro-differential
-Equations (PIDEs) (\cite{nualart2001backward,briani_chioma_natalini,almendral2005numerical,kiessling2011diffusion})
-that generalise
-the Black-Scholes equation by incorporating a
-non-local integral term to account for the discontinuities in
-the asset price. This approach has also been extended to
-cases where the option price features path dependence,
-for instance in \cite{boyarchenko2002barrier} \cite{d2004penalty} and \cite{lord2008fast}.
-
-%\subsection{Fourier methods}
-The \levy-Khintchine formula provides an explicit 
-representation of the characteristic function of a 
-\levy process (cf, \cite{tankov2004financial}).
-As a consequence, one can derive an exact expression for the 
-Fourier transform of the solution of the relevant PIDE.
-Using the inverse fast Fourier transform (iFFT) method, one
-may efficiently compute the option price for a range of asset prices
-simultaneously.
-Furthermore, in the case of European call options, one may
-use the duality property presented by \cite{dupire1997pricing}
-and iFFT to efficiently compute option prices for a wide range of strike prices.
-
-%\subsection{Need for an error bound and state of the art (Lee)}
-Despite the popularity of Fourier methods for option pricing, few works
-can be found on the error analysis and related parameter selection for these methods. 
-A bound for the error not only provides an interval for the precise
-value of the option, but also suggests a method to select the parameters of the numerical method.
-An important
-work in this direction is the one by \cite{lee2004option} in which several
-payoff functions are considered for a rather general set of models, whose
-characteristic function is assumed to be known.
-\cite{fenglinetsky} presents the framework and theoretical approach
-for the error analysis, and establishes polynomial convergence rates
-for approximations of the option prices.
-For a more
-contemporary review on the error committed in various
-FT-related methods we refer the reader to \cite{boyarchenko2011new},
-that extends the classical \emph{flat} Fourier methods by deforming the integration countours
-on the complex plane, studying discretely monitored barrier options
-studied in \cite{de2014pricing}.
-
-%\subsection{Our contribution and comparison with Lee}
-In this work, we present a methodology for studying
-and bounding  the error committed
-when using FT methods to compute option prices.
-We also provide a systematic way of choosing 
-the parameters of the numerical method, in a way that minimises
-the strict error bound, thus guaranteeing adherence to a pre-described
-error tolerance. 
-We focus on exponential \levy processes that may be of either
-diffusive or pure jump. Our contribution is to derive a strict error bound for a Fourier
-transform method when pricing options under risk-neutral 
-\levy dynamics. We derive a simplified bound that separates the contributions of
-the payoff and of the process in an easily processed and extensible
-product form that is independent of the asymptotic behaviour of the
-option price at extreme prices and at strike parameters.
-We also provide a proof for the existence of optimal
-parameters of the numerical computation that minimise the presented
-error bound.
-When comparing our work with Lee's work we find that Lee's work is more general than ours in that he studies a wider range
-of processes, on the other hand, our results apply to a larger class of payoffs.
-On test examples of practical relevance, we also find that the bound presented produces comparable or better results than the ones previously presented in the literature, with acceptable computational cost.
-
-%\subsection{Organization of the paper}
-The paper is organised in the following sections:
-In Section \ref{section:Method}
-we introduce the PIDE setting in the context of
-risk-neutral asset pricing; we show the Fourier representation of the relevant
-PIDE for asset pricing with \levy processes and use that
-representation for derivative pricing.
-In Section \ref{section:Bound}
-we derive a representation
-for the numerical error and divide it into
-quadrature and cutoff contributions. We also describe the
-methodology for choosing numerical
-parameters to obtain minimal error bounds
-for the FT method.
-The derivation is supported by numerical examples
-using relevant test cases with both diffusive and
-pure-jump \levy processes in Section \ref{section:Numerics}.
-Numerics are followed by conclusions in Section \ref{section:Conclusion}.
-
-\section{Fourier method for option pricing}
-
-\label{section:Method}
-Consider an asset whose price at time $t$ is modelled 
-by the stochastic process $S=(S_t)$ defined by $S_t=S_0 \expfs{X_t}$,
-where $X=(X_t) \in \R$ is assumed to be
-a \levy process whose jump measure $\nu$ satisfies
-\begin{align} \label{eq:nuhyp}
-\int_{\rz} \min\{y^2,1\} \nu \parent{dy} < \infty
-\end{align}
-
-Assuming the risk-neutral dynamic for $S_t$, the price at time $t=T-\tau$ of a European option
-with payoff $G$ and maturity time $T$ is given by
+Let us first discretise the time interval as
 \begin{align*}
-\Pi(\tau,s)=\expfs{-r\tau} \expp{G\parent{S_T}|S_{T-\tau}=s}
+\Delta t_{\bar \ell} &= 2^{-\ell_1} t_{max},
+\\
+t_{\bar \ell, n} &= n \Delta t_{\bar \ell}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_1}].
 \end{align*}
-where $r$ is the short rate that we assume to be constant and $\tau\colon 0\leq\tau\leq T$ is the time to maturity. 
-Extensions to non-constant deterministic short rates are straightforward.
-
-The infinitesimal generator of a \levy
-process $X$ is given by \citep[see][]{applebaum2004levy}
-\begin{align}
-\lx f(x)&\equiv 
-\lim_{h\to 0} \frac{\expp{f(X_{t+h})|X_t=x}-f(x)}h \notag \\
-&= \gamma f'(x)+\frac12\sigma^2f''(x)+\int_{\rz} \left(f(x+y)-f(x)-y1_{|y|\leq 1} f'(x)\right)\nu(\d y)
-\end{align}
-where $(\gamma,\sigma^2,\nu)$ is the characteristic triple of the \levy process.
-The risk-neutral assumption on $(S_t)$ implies
-\begin{equation} \label{eq:riskneutralexpint}
-\int_{\absval{y}>1}\expfs{y} \nu(\d y)<\infty
-\end{equation}
-and fixes the drift term (see \cite{kiessling2011diffusion}) $\gamma$ of the \levy process to
-% TODO: find more references to this
-\begin{align} 
-\label{drift_equation}
-\gamma = r-\frac12 \sigma^2 - \int_{\rz} \parent{\expfs{y}-1-y1_{|y|\leq 1} }\nu(\d y)
-\end{align}
-Thus, the infinitesimal generator of $X$ may be written under the risk-neutral assumption as
-\begin{align}
-\lx f(x)&= \left(r-\frac{\sigma^2}2\right) f'(x)+\frac{\sigma^2}2 f''(x)+\int_{\rz} \left(f(x+y)-f(x)-(\expfs{y}-1)f'(x)\right)\nu(\d y)
-\end{align}
-Consider $g$ as the reward function in log prices (ie, defined by 
-$g(x)=G(S_0 \expfs{x})$). Now, take $f$ to be defined as
+Since there is an additional requirement that $t_{max}$ be in
+the mesh of maturities, one point in the mesh in $\tau \in [0,\tau_{max}]$. We let $\ell_1$ and $\ell_2$ define the number of discretisation points to the right and to the left of the fixed
+point, respectively.
 \begin{align*}
-f \parent{\tau,x} \equiv \expp{g \parent{X_{T}}|X_{T-\tau}=x}
-\end{align*}
-Then $f$ solves the following PIDE:
-\begin{equation*}
-\begin{cases}
-\partial_{\tau}f(\tau,x)&=\lx f(\tau,x) \\
-f(0,x)&=g(x), \qquad \qquad \parent{\tau,x}  \in [0,T] \times \R
-\end{cases}
-\end{equation*}
-Observe that $f$ and $\Pi$ are related by
-\begin{equation}
-\Pi(\tau,S_0 \expfs{x})=\expfs{-r \tau}f(\tau,x)
-\end{equation}
-Consider a damped version of $f$ defined by $\fa(\tau,x)=\expfs{-\damp x}f(\tau,x)$; we see that $\partial_{\tau} \fa = \expfs{-\damp x} \lx f\parent{\tau,x} $.
-
-There are different conventions for the Fourier transform. Here we consider the operator $\F$ such that 
-\begin{align}
-\F[f] \parent{\omega} \equiv \int_{\R}\expfs{i\omega x} f \parent{x} \d x
-\label{eq:fourierDefinition}
-\end{align}
-defined for functions $f$ for which the previous integral is convergent. We also use $\ft{f} \parent{\omega}$ as a shorthand notation of $\F[f]\parent{\omega}$.
-To recover the original function $f$, we define the inverse Fourier transform as 
-$$\ift{f}(x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x}f(\omega) \d \omega$$
-We have that $\ift{\ft{f}}(x)=f(x)$.
-
-Applying $\F$ to $\fa$ we get $\ft{\fa}(\omega)=\ft{f} \parent{\omega + i \damp}$.
-Observe also that the Fourier transform applied to $\lx f(\tau,x)$ gives $\charExp{-i\omega}\ft{f} \parent{\tau,\omega}$,
-where $\charExp{\cdot}$ is the characteristic exponent of the process $X$, which satisfies $\expp{\expfs{zX_t}}=\expfs{t\charExp{z}}$.
-The explicit expression for $\charExp{\cdot}$ is 
-\begin{equation}
-\label{characteristicDefinition}
-\charExp{z}
-=\parent{r-\frac{\sigma^2}2} z +\frac{\sigma^2}2 z^2 +\int_{\R} \left(\expfs{z y}-1-(\expfs{y}-1)z\right)\nu(\d y) \\[.5em]
-\end{equation}
-From the previous considerations it can be concluded that 
-\begin{equation}
-\partial_{\tau} \ft{\fa} = \charExp{\damp -i \omega} \ft{f} \parent{\omega - i \damp}
-\end{equation}
-Now $\ft{f}\parent{\omega - i \damp} = \ft{\fa} \parent{\omega}$ so $\ft{\fa}$ satisfies the following ODE
-\begin{equation}
-\begin{cases}
-\frac{\partial_{\tau}\ft{\fa}\parent{\tau,\omega}}{\ft{\fa}(\tau,\omega)}&=\charExp{\damp -i \omega}\\[.5em]
-\ft{\fa} \parent{0,\omega}&=\ft{\ga}\parent{\omega}
-\end{cases}
-\end{equation}
-
-Solving the previous ODE explicitly, we obtain
-\begin{equation}
-\ft \fa \parent{\tau,\omega}
-= \expfs{\tau \charExp{\damp-i \omega} } \ft{\ga} \parent{\omega}
-\label{eq:theODE}
-\end{equation}
-Observe that the first factor in the right-hand side in the above 
-equation is $\expp{\expfs{\parent{\damp-i \omega}X_{\tau}}}$, (ie, $\charFun{1}{-i\damp-\omega}$), where $\charFun{\tau}{\cdot}$ denotes the characteristic function of the
-random variable $X_{\tau}$
-\begin{align}
-\charFun{\tau}{\omega} \equiv \expp{\tau \charExp{i \omega}} 
-\end{align}
-
-Now, to obtain the value function we employ
-the inverse Fourier transformation, to obtain
-\begin{equation}
-\fa(\tau,x)=\ift{\ft{\fa}}(\tau,x)=\frac{1}{2\pi}\int_{\R} \expfs{-i\omega x} \ft{\fa}(\tau,\omega)d\omega
-\label{eq:fadefinition}
-\end{equation}
-or
-\begin{equation}
-\fa(\tau,x)=\frac{1}{\pi}\int_{0}^{+\infty} \real{\expfs{-i\omega x} \ft{\fa}(\tau,\omega)}d\omega
-\label{eq:fadefinition2}
-\end{equation}
-
-As it is typically not possible to compute the inverse Fourier transform analytically,
-we approximate it by
-discretising and truncating the integration domain
-using trapezoidal quadrature \eqref{eq:fadefinition}.
-Consider the following approximation:
-\begin{align}
-\label{eq:fdiscrete}
- \fadisc{\step,n} (\tau,x)&= \frac{\step}{2\pi} \sum_{k=-n}^{n-1} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}\\
- &=\frac{\step}{\pi} \sum_{k=0}^{n-1} \real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}
-\end{align} 
-Bounding and consequently minimising the error
-in the approximation of $f(\tau,x)$ by $$\fdisc{\step,n}(\tau,x) \equiv \expfs{\damp x}\fadisc{\step,n}(\tau,x)$$
-is the main focus of this paper and will be addressed in the following section.
-
-\begin{remark}
-Although we are mainly concerned with option pricing when 
-the payoff function can be damped in order to guarantee
-regularity in the $L^1$ sense, we note here that our main 
-results are naturally extendable to include the Greeks of the option. 
-Indeed, we have by \eqref{eq:theODE} that 
-\begin{align}
- f \parent{t,x} = \frac{1}{2 \pi} \int_{\mathbb R} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
-\end{align}
-so the Delta and Gamma of the option equal
-\begin{align}
-\Delta \parent{t,x} &\equiv \frac{\partial f \parent{t,x}}{\partial x} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
-\\
-\Gamma \parent{t,x} &\equiv \frac{\partial^2 f \parent{t,x}}{\partial x^2} = \frac{1}{2 \pi} \int_{\mathbb R} \parent{\damp-i \omega}^2 \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} \d \omega
-%\\
-%\Theta \parent{t,x}  &\equiv \frac{\partial f \parent{t,x}}{\partial t} = -\frac{1}{2 \pi} \int_{\mathbb R} \charExp{\damp-i \omega} \expfs{\parent{\damp - i \omega}x} \ft \fa \parent{\tau, \omega} d\omega
-\end{align}
-Because the expressions involve partial derivatives with respect to only $x$,
-the results in this work are applicable for the computation of $\Delta$ and $\Gamma$
-through a modification of the payoff function:
-\begin{align}
-\ft g_{\damp, \Delta} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp - i \omega}
-\\
-\ft g_{\damp, \Gamma} \parent{\omega} =& \ft g_{\damp} \parent{\omega} \parent{\damp-i \omega}^2
-\end{align}
-When the Fourier space payoff function manifests
-exponential decay, the introduction of a coefficient that is polynomial in $\omega$
-does not change the regularity of $\ft g$ in a way that would significantly change the following
-analysis. Last, we note that since we do our analysis for PIDEs on a mesh of
-$x$'s, one may also compute the option values in one go and obtain the Greeks with
-little additional effort using a finite difference approach for the derivatives.
-\end{remark}
-
-\subsection{Evaluation of the method for multiple values of $x$ simultaneously}
-
-The Fast Fourier Transform (FFT) algorithm provides an efficient 
-way of computing \eqref{eq:fdiscrete} for an equidistantly
-spaced mesh of values for $x$ simultaneously.
-Examples of works that consider this widely extended tool are \cite{lord2008fast,jackson2008fourier,hurd2010fourier} and \cite{schmelzle2010option}.
-%In Subsection \ref{subsection:FFT} we analyse this alternative.
-
-Similarly, one may define the Fourier frequency $\omega$
-as the conjugate variable of some external parameter on which
-the payoff depends. Especially, for the practically relevant case
-of call options, we can denote the log-strike as $k$ and treat
-$x$ as a constant and write:
-\begin{align}
-\kft f_{k,\damp} \parent{\omega} \equiv
-\int_\R \expfs{\parent{\damp +i \omega }k } f_k \parent{x} \d k
-\end{align}
-
-Using this convention, the time dependence is given by
-\begin{align}
-\label{eq:kftDef}
-\kft f_{k,\damp} \parent{\tau,x} = 
-\frac{\expfs{\parent{i \omega + \damp + 1} x} \charFun{\tau}{\omega - i \parent{\damp +1} }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
-\end{align}
-contrasted with the $x$-space solution
-\begin{align}
-\label{eq:xftDef}
-\ft f_{k,\damp} \parent{\tau,x} = 
-\frac{\expfs{\parent{i \omega - \damp + 1} k} \charFun{\tau}{\omega + i \damp }}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
-\end{align}
-We note that for call option payoff to be in $L^1$,
-we demand that $\damp$ in \eqref{eq:kftDef} is positive.
-Omitting the exponential factors that contain the $x$ and $k$
-dependence in \eqref{eq:kftDef} and \eqref{eq:xftDef} respectively,
-we have that one can arrive from \eqref{eq:kftDef} to \eqref{eq:xftDef}
-using the mapping $\damp \mapsto -\damp-1 $. Thanks to this,
-much of the analysis regarding the $x$-space transformation
-generalises in a straightforward manner to the $k$-space transform.
-
-\section{Error bound}
-\label{section:Bound}
-
-The aim of this section is to compute a bound of the error when approximating 
-the option price $f(\tau,x)$ by $\fadisc{\step,n} (\tau,x)$,
-defined in \eqref{eq:fdiscrete}.
-Considering
-\begin{equation} \label{eq:serie}
-\fadisc{\step}(\tau,x)= \frac{\step}{2\pi} \sum_{k\in \Z} \expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}
-\end{equation}
-the total error $\Error$ can be split into a sum of two terms: the quadrature and truncation errors.
-The former is the error from the approximation of the integral in \eqref{eq:fadefinition}
-by the infinite sum in \eqref{eq:serie}, while the latter is due to the truncation of the infinite sum. 
-Using triangle inequality, we have
-\begin{align} \label{eq:errordef}
-\Error &:=\absval{f(\tau,x)-\fdisc{\step,n}(\tau,x)}\leq \EQ + \EF
-\end{align}
-with
-\begin{align*}
-\EQ &= \expfs{\damp x}\absval{\fa(\tau,x)-\fadisc{\step}(\tau,x)}\\
-\EF &= \expfs{\damp x} \absval{\fadisc{\step}(\tau,x)-\fadisc{\step,n}(\tau,x)}
-\end{align*}
-Observe that each $\Error$, $\EQ$ and $\EF$ depend on
-three kinds of parameters:
-\begin{itemize}
- \item Parameters underlying the model and payoff such as volatility and strike price. We call these \emph{physical parameters.}
- \item Parameters relating to the numerical scheme such as $\damp$ and $n$.
- \item \emph{Auxiliary parameters} that will be introduced in the process of deriving the error bound. These parameters
- do not enter the computation of the option price, 
- but they need to be chosen appropriately to have as tight a bound as possible. 
-\end{itemize}
-
-We start by analysing the quadrature error.
-\subsection{Quadrature error}
-
-Denote by $\strip{a}$, with $a>0$, the strip of width $2a$ around the real line:
-$$\strip{a} \equiv \left\{z\in \C \colon \absval{\im{z}}<a \right\}$$
-The following theorem presents conditions under which the quadrature error goes to zero at a spectral rate as $\step$ goes to zero.
-Later in this section, we discuss simpler conditions to verify the hypotheses and analyse in more detail the case 
-when the process $X$ is a diffusive process or there are ``enough small jumps."
-
-\begin{theorem}\label{theo:quadrature} 
-Assume that for $a > 0$:
-\begin{itemize}
-\item[H1.] the characteristic function of the random variable $X_1$ has an analytic extension to the set $$\strip{a}-\damp i \equiv \left\{z\in \C \colon \absval{\im{z}+\damp}<a\right\}$$
-\item[H2.] the Fourier transform of $\ga(x)$ is analytic in the strip $\strip{a}$ and
-\item[H3.] there exists a continuous function $\gamma \in L^1(\R)$ such that  $\absval{\ft{\fa}(\tau,\omega+i \beta)}<\gamma(\omega)$ for all $\omega\in \R$ and for all $\beta \in [-a,a]$
-\end{itemize}
-Then the quadrature error is bounded by
-\begin{equation*}
-\ErrorQuadrature \leq \expfs{\damp x} \frac{M_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
-\end{equation*}
-where $M_{\damp,a}(\tau,x)$ is given by
-\begin{align}\label{eq:defM}
-M_{\damp,a}(\tau,x)&:=\sum_{\beta \in \{-a,a\}}\int_{\R}\absval{\expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)} \d \omega
-\end{align}
-$M_{\damp,a}(\tau,x)$ equals the Hardy norm (defined in \eqref{eq:hardynorm}) of the function $\omega \mapsto \expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)$, which
-is finite.
-\end{theorem}
-
-The proof of Theorem \ref{theo:quadrature} is an application of Theorem 3.2.1 in \cite{stenger2012numerical},
-whose relevant parts we include for ease of reading. Using the notation in \cite{stenger2012numerical}, $H^1_{\strip{a}}$ is the family of functions $w$ that are
- analytic in $\strip{a}$ and such that 
-\begin{equation}\label{eq:hardynorm}
-\nnorm{w}{H^1_{A_a}}:= \lim_{\varepsilon\to 0} \int_{\partial \strip{a}(\varepsilon)} \absval{w(z)}d\absval{z} < \infty
-\end{equation}
- where 
- $$\strip{a}(\varepsilon) = \left\{z\in \C\colon \absval{\real{z}}<\frac1{\varepsilon},\ \absval{\im{z}} < a \parent{1-\varepsilon}\right\}$$
-\begin{lemma}[Theorem 3.2.1 in \cite{stenger2012numerical}]\label{lem:stenger}
-Let $w \in H^1_{A_a}$,
-then define:
-\begin{align}
-I \parent{w} &= \int_\R w \parent{x} \d x \\
-J \parent{w,h} &= h \ssum{j=-N}{N} w \parent{jh} \\
-\zeta \parent{w,h} &= I \parent{w}- J \parent{w,h}
-\end{align}
-then
-\begin{align}
-\absval{\zeta \parent{w,h}} \leq \frac{e^{-\frac{\pi a}{h}} \nnorm{w}{H^1_{A_a}}}{2 \mathrm{sinh} \parent{\frac{\pi a}{h}}}
-\end{align}
-\end{lemma}
-
-\begin{proof}[Proof of Theorem \ref{theo:quadrature}]
-First observe that H1 and H2 imply that the function $w(z)=\expfs{-i x z + \step/2}\ft{\fa}(\tau,z+ \step/2)$ is analytic in $\strip{a}$. 
-H3 allows us to use dominated convergence theorem to prove that $\nnorm{w}{H^1_{A_a}}$ is finite and coincides with $M_{\damp,a}(\tau,x)$. Applying Lemma \ref{lem:stenger} 
-the proof is completed.
-\end{proof}
-
-Regarding the hypotheses of Theorem \ref{theo:quadrature}, the next propositions provide simpler conditions that imply H1 and H2 respectively.
-
-\begin{proposition} \label{prop:analy1}
-If $\alpha$, $a$ and $\nu$ are such that
-\begin{equation}
-\label{eq:alterH1}
-\int_{y>1} \expfs{\parent{\damp+a}y}\nu(\d y) < \infty \quad \text{and}\quad \int_{y<-1} \expfs{\parent{\damp-a}y}\nu(\d y) < \infty
-\end{equation}
-then H1 in Theorem \ref{theo:quadrature} is fulfilled.
-
-\begin{proof}
-Denoting by $\charFun{1}{\cdot}$ the characteristic function of $X_1$, we want to prove that $z\mapsto \charFun{1}{z+\damp i}$ is analytic in $\strip{a}$.
-Considering that $\charFun{1}{z+\damp i}=\expfs{\charExp{iz-\damp}}$, the only non-trivial part of the proof is to verify that
-\begin{equation}
-z\mapsto \int_{} p(z,y)\nu(\d y)
-\end{equation}
-is analytic in $\strip{a}$,
-where $p\colon \strip{a} \times \R \to \C$ is given by
-$$p(z,y)=\expfs{y \parent{i z-\damp}}-1-(\expfs{y}-1)\parent{i z-\damp}$$
-To prove this fact, we show that we can apply the main result and the only theorem in \cite{mattner2001complex}, 
-which, given a measure space $(\Omega,\mathcal{A},\mu)$ and an open subset $G\subseteq \C$, 
-ensures the analyticity of $\int f(\cdot,\omega) d\mu(\omega)$, provided that $f\colon G\times \Omega \to \C$ satisfies: $f(z,\cdot)$ is $\mathcal{A}$-measurable for all $z\in G$; $f(\cdot,\omega)$ is holomorfic for all $\omega \in \Omega$; and $\int \absval{f(\cdot,\omega)} d\mu(\omega)$ is locally bounded.
-In our case we consider the measure space to be $\R$ with the Borel $\sigma$-algebra and the Lebesgue measure, $G=\strip{a}$ and $f=p$. It is clear that 
-$p(x,\cdot)$ is Borel measurable and $p(\cdot,y)$ is holomorphic. It remains to verify that  
-\begin{equation*}
-z \mapsto \int_{\R^*} \absval{p(z,y)}\nu(\d y)
-\end{equation*} 
-is locally bounded. 
-To this end, we assume that $\real{z}<b$ (and, since $z\in \strip{a}$, $\im{z}<a$) and split the integration domain in $\absval{y}>1$ and ${0<\absval{y}\leq 1}$ to prove that both integrals are uniformly bounded.
-
-Regarding the integral in ${\absval{y}> 1}$, we observe that
-\begin{equation}
-\absval{p(z,y)}\leq \expfs{y(\damp+\im{z})}+1+(\expfs{y}+1)(\damp+a+b)
-\end{equation}
-for $y<-1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp-a)}$ while for $y>1$ we have $\expfs{y(\damp+\im{z})}<\expfs{y(\damp+a)}$. 
-Using the previous bounds and the hypotheses together with \eqref{eq:nuhyp} and \eqref{eq:riskneutralexpint}, we obtain the needed bound.
-
-For the integral in ${0<\absval{y}\leq 1}$, observe that, denoting $f(z,y)=\absval{p(z,y)}$, we have, $f(z,0)=0$ for every $z$, $\partial_y f(z,0)=0$ for every $z$, and $\absval{\partial_{yy}f(z,y)}< c$ for $z\in \strip{a}, \real{z}<b, \absval{y}<1$.
-From these observations we get that the McLaurin polynomial of degree one of $y\mapsto f(z,y)$ is null for every $z$, and we can bound $f(z,y)$ by the remainder term, which, in our region of interest, is bounded by $\frac{c}{2} y^2$, obtaining
-\begin{equation}
-\int_{0<\absval{y}\leq 1} \absval{p(z,y)}\nu(\d y)\leq \frac{c}2 \int_{0<\absval{y}\leq 1}y^2\nu(\d y)
-\end{equation}
-which is finite by hypothesis on $\nu$, which finishes the proof.
-\end{proof}
-\end{proposition}
-
-\begin{proposition}
-If for all $b<a$, the function $x\mapsto \expfs{b\absval{x}}{\ga(x)}$ is in $L^2(\R)$ then H2 in Theorem \ref{theo:quadrature} is fulfilled.
-\end{proposition}
-\begin{proof}
-The proof is a direct application of Theorem IX.13 in \cite{reedsimon}
-\end{proof}
-We now turn our attention to a more restricted class of Levy processes. 
-Namely, processes such that either $\sigma^2>0$ or there exists $\lambda \in (0,2)$ such that $\cl$ defined in \eqref{eq:PureJumpC} is strictly positive. 
-For this class of processes, we can state our main result explicitly in terms of the characteristic triplet.
-
-Given $\lambda \in (0,2)$, define $\cl$ as
-\begin{equation}
-\label{eq:PureJumpC}
-\cl  = 
-\textrm{inf}_{\kappa > 1} \left\{\kappa^{\lambda} 
- \int_{0<|y| < \frac{1}{\kappa}} y^2 \nu(\d y)\right\}
-\end{equation}
-Observe that $\cl\geq 0$ and, by our assumptions on the jump measure $\nu$, $\cl$ is finite. Furthermore, if $\lambda \in (0,2)$ is such
-that
-\begin{equation}
-\label{eq:pureJumpAssumption}
-\liminf_{\epsilon \downarrow 0} \frac{1}{\epsilon^{\lambda}}
- \int_{0<|y| < \epsilon} y^2 \nu(\d y) > 0
-\end{equation}
-then $\cl>0.$ To see this, observe that \eqref{eq:pureJumpAssumption} implies the existence of $\epsilon_0$ such that 
-\begin{equation*}
-\inf_{\epsilon \leq \epsilon_0}\left\{ \frac{1}{\epsilon^{\lambda}}
- \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} > 0
-\end{equation*}
-If $\epsilon_0<1$ observe that
-\begin{equation*}
-\inf_{\epsilon_0 \leq \epsilon \leq 1}\left\{ \frac{1}{\epsilon^{\lambda}}
- \int_{0<|y| < \epsilon} y^2 \nu(\d y)\right\} \geq  \int_{0<|y| < \epsilon_0} y^2 \nu(\d y) > 0
-\end{equation*}
-where for the first inequality it was taken into account that $\frac{1}{\epsilon^{\lambda}}\geq 1$ and that the integral is increasing with $\epsilon$.
-Combining the two previous infima and considering $\absval{\kappa}=\frac{1}{\epsilon}$ we get that $\cl>0.$
-
-Furthermore, we note that for a \levy model with finite jump
-intensity, such as the Black-Scholes and Merton models that satisfy the first of our
-assumption, $\cl=0$ for all $\lambda \in (0,2)$.
-
-\begin{theorem} \label{teo:boundM}
-Assume that: $\damp$ and $a$ are such that \eqref{eq:alterH1} holds; $\ft{\ga} \in \linf{\strip{a}}$; and either $\sigma^2>0$ or $\cl>0$ for some $\lambda\in (0,2)$.
-Then the quadrature error is bounded by
-\begin{equation*}
-\ErrorQuadrature \leq \expfs{\damp x} \frac{\tilde{M}_{\damp,a}(\tau,x) }{2\pi\parent{\expfs{2\pi a/\step}-1}}
-\end{equation*}
-where 
-\begin{equation}\label{eq:Mbound}
-\tilde{M}_{\damp,a} \parent{\tau,x} 
-=
-\ssum{c\in \sset{-1,1}}{} 
-\expfs{cax}
-\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
-\int_{\R}  \expfs{-\tau \left( \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}\right) } \d \omega
-\end{equation}
-Furthermore, if $\sigma^2>0$ we have
-\begin{equation}
-\label{eq:Mboundsigpositive}
-\tilde{M}_{\damp,a} \parent{\tau,x}  \leq \frac{\sqrt{2\pi}}{\sigma\sqrt{\tau}}\ssum{c\in \sset{-1,1}}{} 
-\expfs{cax}
-\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}
-\end{equation}
-
-\begin{proof}
-Considering $h_{\damp,a}(\tau,x,\omega)$ defined by
-\begin{equation}
-h_{\damp,a} \parent{\tau,x,\omega} = \ssum{c \in \sset{-1,1}}{} \absval{\expfs{-i\parent{\omega + ica} x}\ft{\fa}(\tau,\omega+i ca)} 
-\end{equation}
-we have that
-$$M_{\damp,a}(\tau,x) = \int_{\R}h_{\damp,a} \parent{\tau,x,\omega} d\omega$$
-On the other hand, for  $\beta\in (-a,a)$:
-\begin{equation}
-\absval{\expfs{-i\parent{\omega + i\beta} x}\ft{\fa}(\tau,\omega+i \beta)} = \expfs{\beta x} \absval{\ft{\fa}(\tau,\omega+i \beta)}=\expfs{\beta x}\absval{\expfs{\tau \charExp{\damp+\beta-i\omega}}}\absval{ \ft{\ga}\parent{\omega+i\beta}}
-\end{equation}
-For the factor involving the characteristic exponent we have
-\begin{equation}
-\absval{\expfs{\tau \charExp{\damp+\beta-i\omega}}}=\expfs{\tau \real{\charExp{\damp+\beta-i\omega}}}
-\end{equation}
-
-Now, observe that
-\begin{align}
-\notag \real{\charExp{ \damp +\beta - i \omega}}&= \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
-\label{eq:boundingPsiPureJump} &\quad +\int_{\rz}\parent{ \expfs{\parent{\damp + \beta} y} \cos(-y\omega)-1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y)
-\end{align}
-
-If $|\omega| \leq 1$ we bound $\cos(-y \omega)$ by 1, getting
-\begin{align}
-\notag \real{\charExp{ \damp +\beta - i \omega}} &\leq \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
-\notag &\quad +\int_{\rz} \parent{\expfs{\parent{\damp + \beta} y} -1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y)  \\
-& = \charExp{\damp+\beta} -\frac{\sigma^2}{2}\omega^2 \label{eq:Psibound}
-\end{align}
-
-Assume $| \omega| > 1$.
-Using that for $|x| < 1$ it holds that $\textrm{cos}(x) < 1 - x^2 / 4$, we can bound the first term of the integral in the 
-following manner:
-\begin{align}
-\notag \int_{\rz} \expfs{\parent{\damp + \beta} y} \cos(y\omega) \nu(\d y) & \leq
-\int_{0<|y| < 1/|\omega|} \expfs{\parent{\damp + \beta} y} \left( 1 - \omega^2 y^2 / 4 \right)  \nu(\d y) \\
-\notag & \quad + 
-\int_{|y| \geq  1/|\omega|} \expfs{\parent{\damp + \beta} y}  \nu(\d y) \\
-\notag & 
-\leq \int_{\rz} \expfs{\parent{\damp + \beta} y}  \nu(\d y) 
-- \frac{|\omega|^{2 - \lambda}}{4} | \omega |^{\lambda} \int_{0<|y| < 1 / |\omega|} y^2 \nu(\d y)  \\
-\label{eq:extraTermPureJump} & 
-\quad 
-\leq \int_{\rz} \expfs{\parent{\damp + \beta} y}  \nu(\d y) 
-- \frac{|\omega|^{2 - \lambda}}{4} \cl
-\end{align}
-Inserting \eqref{eq:extraTermPureJump} back into \eqref{eq:boundingPsiPureJump} we get
-\begin{align*}
-\notag 
-\real{\charExp{ \damp +\beta - i \omega}} & \leq \parent{\damp + \beta} \parent{r-\frac{\sigma^2}{2}}+\frac{\sigma^2}{2}\parent{\parent{\damp +\beta}^2 - \omega^2} \\
-\notag &\quad + \int_{\rz} \parent{\expfs{\parent{\damp + \beta} y} -1-\parent{\damp + \beta} \parent{\expfs{y}-1}} \nu(\d y) \\
-\notag & \quad \quad - \frac{|\omega|^{2 - \lambda}}{4} \cl    \\
-& = \charExp{\damp+\beta} -\frac{\sigma^2}{2}\omega^2 - \frac{|\omega|^{2 - \lambda}}{4} \cl
+\Delta \tau_{\bar \ell,1} &= 2^{-\ell_3} t_{max},  \\
+\Delta \tau_{\bar \ell,2} &= 2^{-\ell_2} \parent{\tau_{max}-t_{max}}, \\
+\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_3}] \\
+\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [2^{\ell_3},2^{\ell_3}+1,\cdots, 2^{\ell_3}+2^{\ell 2}].
 \end{align*}
 
-%We conclude that $\real{\Psi \parent{ \damp - i z}}\to -\infty$ uniformly in $A$ when $\absval{z}\to \infty$, and in consequence $h(z)\to 0$.
-Taking the previous considerations and integrating in $\R$ with respect to $\omega$, we obtain \eqref{eq:Mbound}.
-
-Finally, observing that  $\cl\geq 0$ and bounding it by 0, the bound \eqref{eq:Mboundsigpositive} is obtained by evaluating the integral.
-\end{proof}
-\end{theorem}
-
-\begin{remark} \label{remark:callPut}
-In the case of call options, hypothesis H2
-implies a dependence between the strip-width parameter
-$a$ and damping parameter $\damp$. We have that the
-damped payoff of the call option is in $L^1 \parent{\R}$
-if and only if $\damp > 1$ and hence the appropriate choice
-of strip-width parameter is given by $0<a< \damp-1$.
-A similar argument holds for the case of put options, for
-which the Fourier-transformed damped payoff is identical
-to the calls with the distinction that $\damp <0$.
-In such case, we require $a < -\damp$.
-
-The case of binary options whose payoff has finite support
-($G \parent{x} = \mathbf{1}_{[x_-,x_+]} \parent{x}$) 
-we can set any $a \in \R$ (ie, no damping is needed at all and
-even if such damping is chosen, it has no effect on the appropriate
-choice of $a$). 
-\end{remark}
-
-\begin{remark}
-The bound we provide for the quadrature error is naturally positive and increasing in $\step$.
-It decays to zero at a spectral rate as $\step$ decreases to 0.
-\end{remark}
-
-\subsection{Frequency truncation error}
-The frequency truncation error is given by
-\begin{align*}
-\ErrorFrequency &= \frac{\expfs{\damp x}\step}{\pi} \absval{\sum_{k=n}^{\infty} \real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}}
-\end{align*}
-If a function $c\colon (\omega_0,\infty) \to (0,\infty)$ satisfies
-\begin{equation} \label{eq:condFunc}
-\absval{\real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}} \leq c\parent{\parent{k+\frac12}\step}
-\end{equation}
-for every natural number $k$, then we have that 
-\begin{align*}
-\ErrorFrequency &\leq \frac{\expfs{\damp x}\step}{\pi} \sum_{k=n}^{\infty}\absval{\real{\expfs{-i{\parent{k+\frac12}\step} x} \ft{\fa}\parent{\tau,{\parent{k+\frac12}\step}}}} \\
-& \leq \frac{\expfs{\damp x}\step}{\pi} \sum_{k=n}^{\infty} c\parent{\parent{k+\frac12}\step}
-\end{align*}
-Furthermore, if $c$ is a non-increasing concave integrable function, we get 
-\begin{equation}
-\ErrorFrequency \leq \frac{\expfs{\damp x}}{\pi} \int_{n\step}^{\infty} c(\omega) \d\omega
-\label{eq:truncationErrorIntegral}
-\end{equation} 
-When $\ft{\ga}\in \linf{[\omega_0,\infty)}$ and either $\sigma^2>0$ or $\cl>0$, then the function $c$ in \eqref{eq:condFunc} can be chosen as
-\begin{equation} \label{eq:cutoffparticularSimpler}
-c(\omega)=\norminf{\ft{\ga}}{[\omega_0,\infty)}\expfs{\tau \charExp{\damp}}\expfs{-\tau \parent{\frac{\sigma^2}{2} \omega^2 +\frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1} }}
-\end{equation}
-
-To prove that this function satisfies \eqref{eq:condFunc} we can use the same bound we found in the proof of Theorem \ref{teo:boundM}, with $\beta=0$, to obtain
-\begin{equation*}
-\real{\charExp{ \damp - i \omega}}\leq  \charExp{\damp} - \frac{\sigma^2}{2}\omega^2  - \frac{|\omega|^{2 - \lambda}}{4} \cl \indicator_{\absval{\omega}>1}
-\end{equation*}
-from where the result is straightforward.
-
-\subsection{Bound for the full error}
-
-In this section we summarize the bounds obtained for the error under different 
-assumptions and analyse their central properties.
-
-In general the bound provided in this paper are of the form
-\begin{equation}\label{eq:fullErrorGeneral}
-\est = \frac{\expfs{\damp x}}{\pi}\parent{ \frac{\bar{M}}{\expfs{2\pi a/\step}-1} +  \int_{n\step}^{\infty}  c\parent{\omega} \d \omega}
-\end{equation}
-where $\bar{M}$ is an upper bound of $M_{\damp,a}\parent{\tau,x}$ defined in \eqref{eq:defM} and $c$ is non-increasing, integrable and satisfies
-\eqref{eq:condFunc}. Both $\bar{M}$ and $c$ may depend on the 
-parameters of the model and the artificial parameters, but they are independent of $\step$ and $n$. Typically one can remove 
-the dependence of some of the parameters, simplifying the expressions but obtaining
-less tight bounds.
-
-When analysing the behaviour of the bound one can observe that the term correspondent with the quadrature error decreases to zero
-spectrally when $\step$ goes to 0. 
-The second term goes to zero if $n\step$ diverges, but we are unable to determine
-the rate of convergence without further assumptions.
-
-Once an expression for the error bound is obtained, the problem of 
-how to choose the parameters of the numerical method to minimise the bound arises, 
-assuming a constraint on the computational effort one is willing to use. 
-The computational effort of the numerical method depends
-only on  $n$. For this reason we aim at finding the parameters that
-minimise the bound for a fixed $n$.
-The following result shows that the bound obtained, as a function of $\step$, 
-has a unique local minimum, which is the global minimum.
-\begin{proposition}
-Fix $\alpha$, $a$, $n$, and $\lambda$ and consider the bound $\est$ as a
-function of $\step$. There exists an optimal $\step^*\in [\frac{\omega_0}{n},\infty)$ such that $\est$ is
-decreasing in $(\frac{\omega_0}{n},\step^*)$ and increasing in
-$(\step^*,\infty)$; thus, a global minimum of $\est$ is attained at $\step^*$.
-
-Furthermore, the optimal $\step$ is either the only point in which 
-$\step\mapsto p\parent{n\step,b}-c(n\step)$, with $p$ defined in \eqref{eq:defp}, changes sign, or $\step=\frac{\omega_0}{n}$ if $p\parent{\omega_0,b}-c(\omega_0)>0$.
-
-\begin{proof}
-Let us simplify the notation by calling $y=n\step$, $b=2 \pi a n$ and $\tilde{\E}=\pi \expfs{-\damp x}\est$.
-We want to prove the existence of $y^*\colon y^*\geq \omega_0$ such that $\tilde{\E}(y)$ is decreasing for $\omega_0<y< y^*$ and increasing for $y> y^*$.
-We have
-$$\tilde{\E}(y)=\frac{\bar{M}}{\expfs{b/y}-1} +  \int_{y}^{\infty} c(\omega) \d\omega.$$
-The first term is differentiable with respect to $y$ and goes to 0 if $y\to 0^+$. This allows us to express it as an integral of its derivative. We can then express $\tilde{\E}(y)$ as 
-$$\tilde{\E}(y)=\tilde{\E}(\omega_0) + \int_{\omega_0}^y \parent{\frac{b \bar{M} \expfs{b/\omega}}{\parent{\expfs{b/\omega}-1}^2 \omega^2}-c(\omega)}d(\omega)$$
-The first term on the right-hand side of the previous equation is constant. Now we move on to proving that the integrand is increasing with $y$ and it is positive if $y$ is large enough. Denote by
-\begin{equation}\label{eq:defp}
-p(y,b)=\frac{b \bar{M} \expfs{b/y}}{\parent{\expfs{b/y}-1}^2 y^2}
-\end{equation}
-Taking into account that $c$ is integrable, we can compute the limit of the integrand in $\infty$, obtaining
-\begin{equation*}
-\lim_{y \to +\infty}p(y,b)-c(y)=\frac{\bar{M}}{b}>0
-\end{equation*} 
-Let us prove that $p(y,b)$ is increasing with $y$ for all $b>0$, which renders $p(y,b)-c(y)$ also increasing with $y$. The derivative of $p$ with respect to $y$ is given by
-$$\partial_{y} p(y,b)=\frac{b\bar{M}\expfs{b/y}\parent{(b/y)\expfs{b/y}-2\expfs{b/y}+b/y+2}}{y^3\parent{\expfs{b/y}-1}^3}$$
-in which the denominator and the first factor in the numerator are clearly positive. To prove that the remainder factor is also positive, observe that 
-$x\expfs{x}-2\expfs{x}+x+2>0$ if $x>0$.
-
-\end{proof}
-\end{proposition}
-
-\subsection{Explicit error bounds}
-\label{ss:bounds}
-In the particular case when either $\sigma^2>0$ or 
-$\cl>0$ for some $\lambda \in (0,2)$ we can give an explicit version 
-of \eqref{eq:fullErrorGeneral}. Substituting $M$ by $\tilde{M}$ defined in Theorem \ref{teo:boundM} and $c$ by the function given in \eqref{eq:cutoffparticularSimpler} we obtain
-\begin{align}%\tag{\ref{eq:fullErrorGeneral}*}
-\label{eq:fullErrorExpression}
-\est &= \estQuadrature + \estFrequency
+Using this notation, we have the following rates of convergence, following the Multi-Index notation:
+\begin{align}
+\bar \beta &= [2,2,2]^T \\
+\bar \gamma &= [1,1,1]^T .
 \end{align} 
-where
+Choosing the Simpson method for integration, we have the order of quadrature $p_Q=3$, which brings the quadrature error to 
 \begin{align}
-\label{eq:boundQ} \estQuadrature &= 
-\ssum{c\in \sset{-1,1}}{} 
-\frac{
-\expfs{\damp x}
-\expfs{cax}
-\expfs{\tau \Psi \parent{ca}} \absval{\ft{\ga} \parent{ca}}}{\pi \parent{ \expfs{\frac{2 \pi a}{\step}} -1}}
-\int_{\R}  \expfs{-\tau \left( \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}\right) } \d \omega
-\\[.5em]
-\label{eq:boundF}
-\estFrequency &= \frac{\expfs{\damp x}}{\pi}
-\norminf{\ft{\ga}}{\R}
-\expfs{\tau \charExp{\damp}} \int_{n\step}^{\infty} 
-\expfs{-\tau \parent{\frac{\sigma^2}{2} \omega^2+\frac{|\omega|^{2 - \lambda}}{4} \cl \indicator_{\absval{\omega}>1}} } d\omega
+\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}} \leq C \parent{\Delta \tau_{\bar \ell}}^{2 \kappa p_Q} = 
+\tilde C 2^{-6 \kappa \mathrm{min} \parent{\ell_3,\ell_2}}.
 \end{align}
-This reproduces the essential features of Theorem 6.6
-in \cite{fenglinetsky},
-the bound \eqref{eq:boundF} can be further improved by substituting $\norminf{\ft{\ga}}{\R}$ by $\norminf{\ft{\ga}}{[n\step,\infty)}$.
-
-\begin{remark}\label{rem:decuople}
-Observe that the bound of both the quadrature and the cutoff error is given
-by a product of one factor that depends exclusively on the payoff and 
-another factor that depends on the asset dynamic. This property makes it easy to evaluate
-the bound for a specific option under different dynamics of the asset price.
-In Subsection \ref{subsec:call} we analyse the terms
-that depend on the payoff function for the particular case of call options.
-\end{remark}
-
-\begin{remark}\label{rem:asynonasy}
-From \eqref{eq:boundQ} it is evident that the speed of the exponential convergence
-of the trapezoidal rule for analytic functions is dictated by the width of the strip in which
-the function being transformed is analytic. Thus, in the limit of small error tolerances,
-it is desirable to set $a$ as large as possible to obtain optimal rates.
-However, non-asymptotic error tolerances are often practically relevant and in these
-cases the tradeoff between optimal rates and the constant term $\absval{\ft{\ga}}$
-becomes non-trivial. As an example, for the particular case of the Merton model,
-we have that any finite value of $a$ will do. However, this improvement of the rate of
-spectral convergence is more than compensated for by the divergence in the constant term.
-\end{remark}
-
-The integrals 
-in \eqref{eq:boundQ} and \eqref{eq:boundF} can, in some cases, be computed analytically,
-or bounded from above by a closed form expression.
-Consider for instance dissipative models with finite jump intensity. 
-These models are characterised by $\sigma^2>0$ and $\cl=0$. Thus the integrals can be expressed in terms of the cumulative normal distribution $\Phi$:
+On the other hand, the error of the simpson method is the sum of errors for $\tau < t_{max}$ and $\tau \leq t_{max}$.
+The error for these two domains are $\bigo{2^{-2 \ell_2}}$ and $\bigo {2^{-2 \ell_3}}$, respectively.
+Combining together, we get the weak and the strong rate
+for the Simpson quadrature:
 \begin{align}
-\int_{\mathbb R} \expfs{-\tau \frac{\sigma^2 \omega^2}{2}}d\omega&=\sqrt{\frac{2 \pi}{\tau \sigma^2}},
+\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2}} \leq&  \tilde C_Q 2^{-6 \mathrm{min} \parent{\ell_3,\ell_2}}
 \\
-\int_{\varsigma}^\infty \expfs{-\tau \frac{\sigma^2 \omega^2}{2}} d \omega &= 
-\sqrt{\frac{2 \pi}{\tau \sigma^2}}
-\parent{1-\Phi \parent{\varsigma \sqrt{\tau \sigma^2}}}
-\end{align}
-
-Now we consider the case of pure-jump processes (ie, $\sigma^2 =0$) that satisfy the condition $\cl > 0$ for some $\lambda \in (0,2)$. 
-In this case the integrals are expressible
-in terms of the incomplete gamma function $\gamma$.
-First, let us define the auxiliary integral:
-\begin{align}
-%%% Including the full derivation, remove intermediate stages if too dull.
-I \parent{a,b} 
-&\equiv
-\expfs{-a} + a^{-\frac{1}{b}} \gamma \parent{\frac{1}{b},a}
-\nonumber
-\end{align}
-for $a,b>0$. Using this, the 
-%constant coefficients
-integrals become:
-\begin{equation}
-\int_{\R} \expfs{- \tau \frac{\absval{\omega}^{2-\lambda}}{4} \cl \mathbf{1}_{\absval{\omega}>1}}
-= 2 \parent{1+ I \parent{\frac{\tau \cl}{4},2-\lambda}}
-\end{equation}
-
-\begin{equation}
-\int_{\varsigma}^\infty \expfs{- \tau \frac{\absval{\omega}^{2-\lambda}}{4} \cl \mathbf{1}_{\absval{\omega}>1}}
-= \begin{cases}
-I \parent{\frac{\tau \cl}{4},2-\lambda} + 1-\varsigma & \quad \varsigma <1\\
-\varsigma I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda} &\quad \varsigma \geq 1
-\end{cases}
-\end{equation}
-An example of a process for which the previous analysis works
-is the CGMY model presented in \cite{cgmy_fine,carr2003stochastic}, for the regime $\cgmyY > 0$.
-
-Lastly, when both $\cl$ and $\sigma^2$ are positive, 
-the integrals in \eqref{eq:boundQ} and \eqref{eq:boundF} can be bounded by a simpler expression. 
-Consider the two following auxiliary bounds for the same integral, in which $\varsigma\geq 1$:
-\begin{align}
-\int_{\varsigma}^\infty \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl} } d\omega
-& \leq \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2} \int_{\varsigma}^\infty \expfs{-\tau \frac{|\omega|^{2 - \lambda}}{4} \cl} d\omega \\
-& = \varsigma \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2}  I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda} \notag
+\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}} \leq& C_Q 2^{-2 \mathrm{min} \parent{\ell_2,\ell_3}}.
 \end{align}
@@
 \begin{align}
-\int_{\varsigma}^\infty \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl} } d\omega
-& \leq \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \int_{\varsigma}^\infty \expfs{-\tau \frac{\sigma^2}{2}\omega^2} d\omega \\
-& =\sqrt{\frac{2 \pi}{\tau \sigma^2}} \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \parent{1-\Phi(\varsigma\sqrt{\tau \sigma^2})}\notag
+\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}}
 \end{align}
-We have that $b(\varsigma)$, defined as the minimum of the right hand sides of the two previous equations,
-\begin{align*}
-b(\varsigma)=\min & \left\{ \varsigma \expfs{-\tau \frac{\sigma^2}{2}\varsigma^2}  I \parent{\frac{\tau \varsigma^{2-\lambda}\cl}{4},2-\lambda}, \sqrt{\frac{2 \pi}{\tau \sigma^2}} \expfs{-\tau \frac{\varsigma^{2 - \lambda}}{4} \cl} \parent{1-\Phi\parent{\varsigma\sqrt{\tau \sigma^2}}} \right\}
-\end{align*}
-is a bound for the integral. Bearing this in mind we have
-%approximation:
+using Theorem 3.3 of \cite{bjork2013monte}.
+
+On the other hand, theorem 3.2 and equation (3.54) states that
 \begin{align}
-\int_{\R} \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}} } d\omega
-& \leq 2 \Phi\parent{\sqrt{\tau\sigma^2}}-1 + 2 b(1)
+\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{2 \kappa}  } ^{\frac{1}{2 \kappa}} \leq C_{5}^{CV} \parent{ \parent{\Delta t}^{\frac{1}{2}} + \Delta \tau }.
 \end{align}
-and
+
+The weak rate for the discretisation error can be bounded by eq. (4.6):
 \begin{align}
-\int_{\varsigma}^{\infty} \expfs{-\tau \parent{ \frac{\sigma^2}{2}\omega^2 + \frac{|\omega|^{2 - \lambda}}{4} \cl\indicator_{\absval{\omega}>1}} } d\omega
-\leq & b\parent{\varsigma}
+\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{}  }
+= 
+ \bigo{\parent{\Delta t}^2 + \parent{\Delta \tau}^2}
+ = \bigo{2^{-2 \ell_1} + 2^{-2 \mathrm {min} \parent{\ell_2,\ell_3} }}
 \end{align}
-provided that $\varsigma\geq 1$.
 
-\section{Computation and minimization of the bound}
-\label{section:Numerics}
+Due to the fact that the discretisation errors relating to $\tau$
+depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
+to assume $\ell_3=\ell_2$. This gives us a two-dimensional
+Multi-Index Monte Carlo formulation with the following effective parameters
+\begin{align}
+\bar \beta &= [2,2]^T ,\\
+\bar \gamma &= [1,1]^T , \\
+\bar s &= [2^{-1},1], \\
+\bar w & = [2,2].
+\end{align}
 
-In this section, we present numerical examples on the
-bound presented in the previous section using practical
-models known from the literature.
-We gauge the tightness of the bound compared to the
-true error using both dissipative and pure-jump processes.
-We also demonstrate the feasibility of using the expression of the bound
-as a tool for choosing numerical parameters
-for the Fourier inversion.
+On the other hand, one may set $\ell_1=\ell_2=\ell_3=\ell$
+and arrive at a one-dimensional MLMC formulation with a weak rate of $2$, strong rate of $1$, and computational cost per realisation of $W_\ell=2^{2\ell}$.
+In the sequel, we will present the results of these two variations for comparison.
 
-\subsection{Call option in variance gamma model}
+\section{Multi-Index Monte Carlo}
 
-The variance gamma model provides a test case
-to evaluate the bound in the pure-jump setting. We note
-that of the two numerical examples presented, it is the less regular one
-in the sense that $\sigma^2=0$ and $C \parent{\lambda} = 0$ for $0 < \lambda < 2$,
-indicating that Theorem \ref{teo:boundM} in particular is not applicable. 
-%\marginpar{Do we really need this sentence? If so, what does it mean?
-%$\exists ~ \tau^*>0: \charFun{\tau}{\cdot} \notin L_1\parent{\R} ~ \forall \tau<\tau^*$.
-%It also is a case in which the characteristic function is not integrable for sufficiently short maturities, making approximation of the kind presented in (28) impossible. 
-% }
-
-The \levy measure of the VG model is given by:
-\begin{align*} \label{eq:VG}
-\nu_{\mathrm{VG}} \parent{\d y} 
-= \d y \parent{
-\mathbf{1}_{y>0} \frac{\cgmyC \expfs{-\cgmyM y}}{y}
-- \mathbf{1}_{y<0} \frac{\cgmyC \expfs{\cgmyG y}}{y}
-}
-\end{align*}
-and the corresponding characteristic function is given by
-eq. (7) of \cite{madan1998variance}:
-\begin{align*}
-\charFun \tau \omega &= 
-\parent{1 - i \theta \chi \omega + \frac{\sigma^2 \chi}{2}}^{-\frac{\tau}{\chi}}
-\\
-\cgmyC &= \chi^{-1}
-\\
-\cgmyG &= \parent{\sqrt{\frac{\theta^2 \chi^2}{4}+\frac{\sigma^2 \nu}{2}} -\frac{\theta \chi}{2}}^{-1}
-\\
-\cgmyM &= \parent{\sqrt{\frac{\theta^2 \chi^2}{4}+\frac{\sigma^2 \nu}{2}} +\frac{\theta \chi}{2}}^{-1}
-\end{align*}
-
-By Proposition \ref{prop:analy1} 
-we get that 
-\begin{align} 
-a < \mathrm{min} \sset{\cgmyG-\damp,\cgmyM+\damp}
-\end{align}
-which, combined with the requirement that
-$\ga \in L^1 \parent{\R}$ (cf, Remark \ref{remark:callPut}), implies:
-\begin{align}
-a &< \mathrm{min} \sset{\cgmyM-\damp,\cgmyG+\damp,\damp-1}
-\\
-a &< \mathrm{min} \sset{\cgmyM-\damp,\cgmyG+\damp,-\damp}
-\nonumber
-\end{align}
-for calls and puts, respectively. We note that evaluation
-of the integral in \eqref{eq:fadefinition} is possible also for $\damp \in (0,1)$
-and for $\damp < 0$. In fact, there is a correspondence between shifts in the
-integration countour and put-call parity. Integrals with $\damp<0$ give rise to put
-option prices instead of calls. For an extended discussion of this, we refer to
-\cite{lee2004option} or \cite{boyarchenko2011new}, in which 
-conformal deformation of the
-integration contour is exploited in order to achieve improved numerical accuracy.
-
-In \cite{lee2004option} and in our calculations the parameters equal
-$\cgmyM = 39.7840$, $\cgmyG = 20.2648$ and $\cgmyC=5.9311$.
-
-Table \ref{tb:usvslee} presents the specific parameters and compares
-the bound for the VG model with the results obtained by \cite{lee2004option}.
-Based on the table, we note that for the VG model
-presented in \cite{madan1998variance} we can achieve comparable or better error bounds when compared to the study
-by Lee.
-
-To evaluate the bound, we perform the integration of 
-\eqref{eq:defM} and \eqref{eq:truncationErrorIntegral}
-by relying on the Clenshaw-Curtis quadrature method
-provided in the SciPy package.
-To supplement Table \ref{tb:usvslee} for a wide range
-of $n$,
-we present the magnitude of the bound compared to the true error in Figure
-\ref{fig:vgerrvsbound}.
-
-In Figure \ref{fig:vgerrvsbound}, we see that the choice of numerical parameters
-for the Fourier inversion has a strong influence on the error of the numerical method.
-One does not in general have access to the true solution. 
-Thus, the parameters need to be optimised 
-with respect to the bound.
-Recall that $\Error = \Error(\alpha,\Delta \omega,a, n)$ and 
-$\est = \est(\alpha,\Delta \omega, n)$ denote the true and estimated
-errors, respectively. Keeping the number of quadrature points $n$ fixed,
-we let $\parent{\alpha_1,\Delta \omega_1,a_1}$ and $\parent{\alpha_2,\Delta \omega_2}$
-denote the minimisers of the estimated and true errors, respectively
-\begin{align}
-\parent{\alpha_1,\Delta \omega_1,a_1}  &= \mathrm{arg ~ inf}~ \est 
-\label{eq:pdef1}
-\\
-\parent{\alpha_2,\Delta \omega_2}  &= \mathrm{arg ~ inf} ~ \Error 
-\label{eq:pdef2}
-\end{align}
-We further let $\mathcal{E}_1$ and $\mathcal{E}_2$ 
-denote the true error as a function of the parameters
-minimising the estimated and the true error, respectively
-\begin{align}
-\mathcal E_1 &= \Error \parent{\damp_1,\step_1} 
-\label{eq:pdef3}
-\\
-\mathcal E_2 &= \Error \parent{\damp_2,\step_2}  
-\label{eq:pdef4}
-\end{align}
-In Figure \ref{fig:vgerrvsbound} we see that the true error increases by approximately an 
-order of magnitude when optimising to the bound instead of to the true error,
-translating into a two-fold difference
-in the number of quadrature points needed for a given tolerance. 
-The difference between ${\Error}_1$ 
-and the bound is approximately another order of magnitude and necessitates
-another two-fold number of quadrature points compared to the theoretical minimum.
-
-In Figure \ref{fig:parameterPlot}, we present the true error
-\footnote{The reference value to compute the true error was
-obtained by the numerical methods with $n$ and $\step$ such
-that the level of accuracy is of the order $10^{-10}$.} for the Fourier method
-for the two test cases in Table \ref{tb:usvslee}. 
- We note that while minimising error
-bounds will produce sub-optimal results, the numerical parameters that minimise
-the bound are a good approximation of the true optimal parameters. 
-This, of course,
-is a consequence of the error bound having qualitatively similar behaviour as
-the true error, especially as one gets further away from the true optimal parameters.
-
-\begin{remark}
-In practice, the Hardy norm in coefficient $M$ reduces to evaluating an
-$L^1$ norm along the two boundaries of the strip of width $2a$.
-We find that, for practical purposes, the performance of the
-Clenshaw-Curtis quadrature of the
-QUADPACK library provided by SciPy library is more than adequate,
-enabling the evaluation of the bound in a fraction of a second.
-
-For example, the evaluations of the bounds in Table
-\ref{tb:usvslee} take only around $0.3$ seconds to evaluate
-on Mid 2014 Macbook Pro equipped with a
-2.6 GHz Intel Core i5 processor, this without attempting to optimize
-or parallelize the implementation and while checking for input sanity
-factors such as the evaluation of the characteristic function in a domain
-that is a subset in the permitted strip.
-
-We believe that through optimizing routines, skipping sanity checks for inputs
-and using a lower-level computation routines this can be optimized even further,
-guaranteeing a fast performance even when numerous evaluations are needed.
-
-\end{remark}
-
-\begin{table}
-\begin{center}\begin{tabular}{ccccccc}
-\hline
- & $K$ & $80$ & $90$ & $100$ & $110$ & $120$\\
-\hline
-\multirow{2}{*}{$12 \tau=1$} & $\damp$ & $ -16.9 $ & $ -13.8 $ & $ 21.6 $ & $ 29.10 $ & $ 36.3 $ \\ 
- & $a$ & $ 3.33 $ & $ 6.45 $ & $ 18.1 $ & $ 9.77 $ & $ 3.52 $ \\ 
-\multirow{2}{*}{$N =32$} & $\cutoff$ & $ 229 $ & $ 229 $ & $ 363 $ & $ 363 $ & $ 424 $ \\ 
- & $\overline{\mathcal E}$ & $ 3.35 \times 10^{-4} $ & $ 0.00334 $ & $ 0.00562 $ & $ 3.97 \times 10^{-4} $ & $ 7.33 \times 10^{-6} $ \\ 
- & $\overline{\mathcal E}^*$ & $ 6 \times 10^{-4} $ & $ 0.0032 $ & $ 0.0058 $ & $ 6 \times 10^{-4} $ & $ 1 \times 10^{-4} $ \\ 
-\hline
-\multirow{2}{*}{$12 \tau=4$} & $\damp$ & $ -13.8 $ & $ -13.8 $ & $ 22.1 $ & $ 23.7 $ & $ 29.10 $ \\ 
- & $a$ & $ 6.11 $ & $ 6.11 $ & $ 17.9 $ & $ 15.2 $ & $ 8.75 $ \\ 
-\multirow{2}{*}{$N =8$} & $\cutoff$ & $ 62.4 $ & $ 42.4 $ & $ 84.9 $ & $ 126 $ & $ 126 $ \\ 
- & $\overline{\mathcal E}$ & $ 3.99 \times 10^{-4} $ & $ 0.00312 $ & $ 0.00398 $ & $ 3.57 \times 10^{-4} $ & $ 1.33 \times 10^{-5} $ \\ 
-& $\overline{\mathcal E}^*$ & $ 1.3 \times 10^{-3} $ & $ 0.0057 $ & $ 0.0055 $ & $ 9 \times 10^{-4} $ & $ 1 \times 10^{-4} $ \\
-
-\hline
-\end{tabular}\end{center}
-\caption{\label{tb:usvslee}The error bound for European call/put options in the VG model for select examples.
-Refererence result $\overline{\mathcal E}^*$ from \cite{lee2004option}}
-\end{table}
-
-\begin{remark}
-Like many other authors, we note the exceptional \emph{guaranteed} accuracy of the FT-method,
-with only dozens of quadrature points. This is partially a result of the regularity of the European option price.
-Numerous Fourier-based methods have been developed for pricing path-dependent options.
-One might, for the sake of generality of implementation, 
-be tempted to use these methods for European options as well, correcting for the lack of early exercise
-opportunities. This certainly can be done, but due to the
-weakened regularity, the required numbers of quadrature points are easily in the thousands, even when no rigorous
-bound for the error is required.
-
-We raise one point of comparison,
-the the European option pricing example in Table 2 of \cite{jackson2008fourier},
-which indicates a number of quadrature points for pricing the option in the range of thousands.
-With the method introduced, to guarantee $\est \approx 10^{-3}$,
-even with no optimisation, $n=64$ turns out to be sufficient.
-\end{remark}
-
-\subsection{Call options under Kou dynamics}
-
-\begin{table}
-\begin{center}\begin{tabular}{ccccccc}
-\hline
- & $K$ & $80$ & $90$ & $100$ & $110$ & $120$\\
-\hline
- & $\overline{\mathcal E}$ & $ 2.67 \times 10^{-4} $ & $  3.49 \times 10^{-4} $ & $  4.43 \times 10^{-4} $ & $ 5.52 \times 10^{-4} $ & $ 6.77 \times 10^{-4} $ \\ 
-  & $\damp$ & $ -1.57 $ & $  -1.57 $ & $  -1.57$ & $ -1.57 $ & $ -1.57$ \\ 
-  & $\cutoff$ & $ 22.9 $ & $  22.8 $ & $  22.6$ & $ 22.5 $ & $22.4$ \\ 
-\hline
-  &$\overline{\mathcal E}^*$   & $ 0.34 $ & $  0.26 $ & $  0.21$ & $ 0.17 $ & $0.13$ \\ 
- & $\overline{\mathcal E}^{\dagger}$  & $ 6.87 \times 10^{-4} $ & $  1.90 \times 10^{-3} $ & $  2.82 \times 10^{-3} $ & $ 2.72 \times 10^{-3} $ & $ 2.29 \times 10^{-3} $ \\ 
- \hline
-\end{tabular}
-\end{center}
-\caption{\label{tb:usvsleekou}
-Numerical performance of the bound for the Kou model, with the test case in
-\cite{toivanen2007numerical} (see also \cite{d2005robust}) with the number
-of quadrature points set to $n=32$.
-The point of comparison $\overline{\mathcal E}^*$ refers to the corresponding bound computed
-with the method described in Chapters $6.1$ to $6.4$ of \cite{lee2004option}.
-In the $\overline{\mathcal E}^{\dagger}$, the cutoff error has been evaluated using a 
-computationally more intensive Clenshaw-Curtis quadrature instead of an asymptotic argument
-with an exponentially decaying upper bound for the option price.
-}
-\end{table}
-
-For contrast with the pure-jump process presented above, we also test the performance
-of the bound for Kou model and present relevant results in \ref{tb:usvsleekou}. This model differs from the first example not only by being dissipative
-but also in regularity, in the sense that the maximal width of the domain $A_a$ is, for the 
-case at hand, considerably narrower. The \levy measure in the Kou model is given by
-\begin{align*}
-\nu_{\mathrm{Kou}} \parent{\d y} = \lambda \parent{ p \expfs{-\eta_1 y} \mathbf{1}_{y>0} + q \expfs{\eta_2 y} \mathbf{1}_{y<0} }
-\end{align*}
-with $p+q=1$.
-For the characterisation given in \cite{toivanen2007numerical} the
-values are set as
-\begin{align*}
-\lambda = 0.1, ~~ r = 0.05 ~~ \tau = 0.25 ~~ S_0 = 100 \\
-p = 0.3445, ~~ \eta_1 = 3.0465 ~~ \eta_2 = 3.0775
-\end{align*}
-from the expression of the characteristic exponent (see  \cite{kou2004option})
-\begin{align*}
-\charExp{z} = z \parent{r -\frac{\sigma^2}{2} - \lambda \zeta} + \frac{z^2 \sigma^2}{2} 
-+ \lambda \parent{\frac{p \eta_1}{\eta_1-z} + \frac{q \eta_2}{\eta_2 +z} -1}
-\end{align*}
-it is straightforward to see
-\begin{align*}
-A_a \subset \sset{z \in \C : \mathrm{Im} z \in ( -3.0465, 3.0775)}
-\end{align*}
-This range is considerably narrower than that considered earlier.
-In the case of transforming the option prices in strike space, the
-relevant expressions for option prices as well as the
-error bounds contain a factor exponential in $k$. The practical implication
-of this is that for deep out of the money calls, it is often beneficial to exploit
-the put-call-parity and to compute deep in the money calls. However, in the case
-at hand, the strip width does not permit such luxury. As a consequence,
-the parameters that minimize the bound are near-identical through a wide range of moneyness,
-suggesting use of FFT algorithm to evaluate the option prices at once for a range of strikes.
-
-\subsection{Binary option in the Merton model}
-For the particular case of Merton model, the \levy measure is given by
-\begin{align*}
-\nu_{\mathrm{Merton}} \parent{\d y}
-=
-\frac{\lambda}{\sqrt{2 \pi \sigma^2}} \expf{-\frac{\parent{y-r_j}^2}{2\sigma_j^2}}
-\end{align*}
-and the characteristic exponent correspondingly by
-\begin{align*}
-\charExp{z}_{\mathrm{Merton}}
-=
-z \parent{r-\frac{\sigma^2 z}{2}}
-+ \frac{\sigma^2 z^2}{2}
-+ \lambda
-\parent{\expfs{z r_j + \frac{\sigma_j^2 z^2}{2}}-1 - z \parent{\expfs{r_j +\frac{\sigma_j^2}{2} }-1} }
-\end{align*}
-we may employ a fast semi-closed form evaluation
-of the relevant integrals instead of resorting to quadrature methods.
-We choose the Merton model as an example of bounding the error of the numerical method for such
-a model. The parameters are adopted from 
-the estimated parameters for S\&P 500 Index from 
-\cite{andersen2000jump}:
-\begin{align*}
-S_0 = 100,
-~~~
-\lambda = 0.089,
-~~
-\sigma = 0.1765,
-~~~
-r = 0.05,
-~~
-r_j = -0.8898,
-~~~
-\sigma_j = 0.4505
-\end{align*}
-
-In Figure \ref{fig:mertonPlot}, we present the bound and true error for the Merton model
-to demonstrate the bound on another dissipative model. The option presented is a binary option
-with finite support on $[95,105]$; no damping was needed or used. We note that
-like in the case of the pure-jump module presented above, our bound reproduces the
-qualitative behaviour of the true error. The configuration resulting from optimising the bound
-is a good approximation of the true error. Such behaviour is consistent through the range of
-$n$ of the most practical relevance.
-
-\subsection{Call options}
-\label{subsec:call}
-In Subsection \ref{ss:bounds} explicit expressions to bound $\Error$ are provided. 
-To evaluate these bounds it is necessary to compute $\norminf{\ft{\ga}}{\R}$ and $\norminf{\ft{\ga}}{\strip{a}}$. According to Remark \ref{rem:decuople}, once
-we compute these values we could use them for any model, provided that 
-they satisfy the conditions considered there.
-
-The payoff of perhaps the most practical relevance is that of
-a call option. 
-Consider $g$ defined by:
-\begin{align*}
- g\parent{x} = \parent{S_0 \expfs{x}-K}^+ = S_0 \parent{\expfs{x}-\expfs{k}}^+
-\end{align*}
-for which a selection of a damping parameter $\damp>0$ is necessary
-to have the damped payoff in $L^1 \parent{\R}$ and to ensure the existence of a
-Fourier transformation. In this case we have% \cite{lordfourier}:
-\begin{align}
-\hat{\ga}
-%\fcallga
-\parent{\omega} &= S_0 \int_\R \expf{ \parent{1-\damp + i \omega}x} - \expf{k+\parent{i \omega- \damp}x}
-\d x
-\\
- %&= S_0 \evaldiff{ 
- %\frac{\expf{\parent{1-\damp + i \omega}x}}{1+ i \omega - \damp}
-%-
- %\frac{\expf{k + \parent{i \omega - \damp}x}}{i \omega - \damp}
-   %}{k}{\infty}
-   %\\
-   &=
-   \frac{ S_0 \expf{\parent{1- \damp + i \omega} k } }{\parent{1 + i \omega- \damp} \parent{i \omega - \damp}}
-%\label{eq:dampedghatcall}
-%%%TODO: check!!!!
-\end{align}
-and
-\begin{align}
- \absval{\ft{\ga}(\omega) }^2
- =
- \frac{S_0^2 \expfs{2 \parent{1-\alpha}k}}{\parent{\alpha^2 + \omega^2 } \parent{\parent{1-\alpha}^2 + \omega^2}}
-\end{align}
-It is easy to see that the previous expression decreases as $|\omega|$ increase. 
-This yields
-\begin{align}
-\norminf{\ft{\ga}}{\R}
-=\absval{\ft{\ga}(0) } =
-\frac{S_0 \expfs{\parent{1-\alpha}k}}{\alpha^2-\alpha}
-\end{align} 
-and 
-\begin{align}
-\norminf{\ft{\ga}}{[\varsigma,\infty)}
-=\absval{\ft{\ga}(\varsigma) }
-\end{align} 
-The maximisation of $\absval{\ft \ga}$ in the strip $\strip{a}$ of the complex plane is more subtle.
-%illustration of the magnitude of the term is plotted in Figure \ref{fig:gHatSup}. 
-Denoting $\ft \ga(\eta,\rho)=\ft \ga(\eta + i \rho)$, we look for critical points that
-satisfy 
-% Local extrema of $\absval{\ft \ga}$, denoting $\omega = \eta + i \rho$ are characterised
-$\partial_{\eta} \absval{\ft \ga}=0$. 
-This gives
-\begin{align}
-4 \eta^3 + 2 \eta \parent{4 \rho \alpha + 2 \alpha^2 - 2 \rho -2 \alpha + \rho^2 +1}=0.
-\end{align}
-For $\rho$ fixed, $\absval{\ft \ga}$ has a vanishing derivative with respect to $\eta$
-at a maximum of three points. Of the three roots of the derivative, only the one characterised
-by $\eta=0$ is a local maximum, giving us that for call options
-\begin{align}
-\norminf{\ft{\ga}}{\strip{a}} = \sup_{\rho \in [-a,a]} \absval{\ft \ga(0,\rho)} 
-\label{eq:complexsup}
-\end{align}
-Now, observe that $\absval{\ft \ga(0,\rho)}$ is a differentiable real function of $\rho$, whose derivative is given by the following polynomial of second degree:
-\begin{align}
- p \parent{\rho} \equiv k \parent{\rho +\alpha - 2 \rho \alpha - \alpha^2 - \rho^2}
- - 2 \alpha - 2 \rho +1
-\end{align}
-We conclude that
-\begin{align}
-\norminf{\ft{\ga}}{\strip{a}} = \max_{\rho \in B} \left\{ \absval{\ft \ga(0,\rho)}\right\} 
-\label{eq:complexsup2}
-\end{align}
-where $B$ is the set of no more than four elements consisting of $a$; $-a$; and the real roots of $p$ that fall in $(-a,a)$.
-
-\begin{remark}
-So far, we have assumed the number of quadrature points $n$ to be constant.
-In real life applications, however, this is often not the case. Typically the user might want to choose a minimal
-$n$ that is sufficient to guarantee error that lies within a pre-defined error tolerance.
-
-In such a case, we propose the following, very simplistic scheme for optimising numerical parameters
-and choosing the appropriate $n$ to satisfy error smaller than $\epsilon$:
-\begin{enumerate}
-\item Select $n=n_0$ and optimize to find the relevant configuration
-\item See if $\EQ + \EF <\epsilon$, if not, increase $n$ by choosing it from a pre-determined
-increasing sequence $n=n_j$ and repeat procedure.
-\end{enumerate}
-Especially in using FFT algorithms to evaluate the Fourier transforms,
-we propose $n_j = 2^j n_0$. We further note that typically the optimal configuration
-for the optimizing configuration for $n_{j+1}$ quadrature points does not differ too
-dramatically from the configuration that optimizes bounds for $n_j$.
-\end{remark}
-
-\section{Conclusion}
-\label{section:Conclusion}
-
-We have presented a decomposition of the error committed in
-the numerical evaluation of the inverse Fourier transform needed
-in asset pricing for exponential \levy models into truncation and
-quadrature errors. For a wide class of exponential \levy models, we
-have presented an $L^\infty$-bound for the error.
-
-The error bound differs from the earlier work presented in \cite{lee2004option}
-in the sense that it does not rely on the asymptotic behaviour of the option
-payoff at extreme strikes or option prices, allowing pricing a wide variety
-of non-standard payoff functions such as the ones in \cite{suh2008class}.
-The bound, however, does not take into account path-dependent options. We
-argue that the error for the methods that allow evaluating american, bermudan,
-or knockoff options are considerably more cumbersome and produce significantly
-larger errors so that in implementations where performance is important, such as
-calibration, using American option pricing methods for European options is not justified.
-
-The bound also provides a general framework in which
-the truncation error is evaluated using a quadrature method that remains invariant
-regardless of the asymptotic behaviour of the option price function.
-The structure of the bound allows for
-a modular implementation that decomposes the error components arising from
-the dynamics of the system and the payoff into a product form for a large
-class of models, including all dissipative models. On select examples, we
-also demonstrate the performance that is comparable or superior to the
-relevant points of comparison.
-
-We have focused on the minimization of the bound as a proxy for minimizing numerical error.
-Doing this, one obtains, for a given parametrization of a model, a rigorous $L^\infty$
-bound for the error committed in solving the European option price. We have shown that the 
-bound reproduces the qualitative behaviour of the actual error. This supports the argument for
-selecting numerical parameters in a way that minimizes the bound, giving evidence that
-this selection will, besides guaranteeing numerical precision, be close to the actual minimizing
-configuration that is not often achievable at an acceptable computational cost.
-
-The bound can be used in the primitive setting of establishing a strict error
-bound for the numerical estimation of option prices for a given set of physical and
-numerical parameters or as a part of a numerical
-scheme, whereby the end user wishes to estimate an option price either on
-a single point or in a domain up to a predetermined error tolerance.
-
-In the future, the error bounds presented can be used in efforts requiring
-multiple evaluations of Fourier transformations. Examples of such applications
-include multi-dimensional Fourier transformations, possibly in sparse tensor grids,
-as well as time-stepping algorithms for American and Bermudan options.
-Such applications are sensitive towards the error bound
-being used, as any numerical scheme will be required to run multiple times, either
-in high dimension or for multiple time steps (or both).
-
-\section{Acknowledgements}
-
-H\"app\"ol\"a, Crocce are members and Tempone is the director KAUST Strategic Research Initiative
-for Uncertainty quantification.
-
-We wish to thank an anonymous referee for their critique that significantly improved the quality of the manuscript.
-
-\section{Declarations of Interest}
-
-The authors report no conflicts of interest. The authors alone are responsible for the content and writing of the paper.
+\section{Multi-Level Monte Carlo}
 
@@
 
-\section{Truncation bound in Lee's scheme}
-
-For the strike-space transformed option price we have
-\begin{align*}
-\tilde \fa \parent{\omega}
-=
-\frac{\charFun 1 {\omega - i \parent{\damp + 1}}}{\parent{i \omega + \damp} \parent{i \omega + \damp + 1}}
-\end{align*}
-In the particular case of the Kou model, we have 
-\begin{align*}
-\charFun 1 \omega
-=
-\expfs{\tau
-\parent{
-i \omega \parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
-- \frac{\omega^2 \sigma^2}{2}
-+ \lambda
-\parent{
-\frac{p \eta_1 }{\eta_1 - i \omega}
-+\frac{q \eta_2}{\eta_2 + i \omega}
--1
-}
-}
-}
-\end{align*}
-Thus,
-\begin{align*}
-\absval{ \charFun 1\omega}
-=
-\expfs{
-\tau
-\parent{ 
-\parent{\damp+1}
-\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
-+
-\frac{\parent{\damp+1}^2 - \omega^2}{2}
-}
-}
-\expfs{ \tau \lambda \mathrm{Re} \parent{
-\frac{p \eta_1}{\eta_1-\damp-1 - i \omega}
-\frac{q \eta_2}{\eta_2 +\damp +1 + i \omega}
--1
- }}
-\end{align*}
-We bound the first jump term resulting from upward jumps as
-\begin{align*}
-\frac{p \eta_1}{ \eta_1 - \damp -1 - i \omega}
-= \frac{p \eta_1 \parent{\eta_1 - \damp -1 +i \omega}}{\parent{\eta_1 - \damp -1}^2 + \omega^2}
-\end{align*}
-from which it follows
-\begin{align*}
-\mathrm{Re}
-\parent{
-\frac{p \eta_1}{ \eta_1 - \damp -1 - i \omega}
-}
-\leq
-\frac{p \eta_1 \parent{\eta_1 - \damp -1 }}{\parent{\eta_1 - \damp -1}^2 }
-\end{align*}
-similar inequality holds for the term resulting from downward jumps giving us
-that
-\begin{align*}
-\absval{ \charFun 1 \omega}
-\leq 
-\expfs{
-\tau
-\parent{ 
-\parent{\damp+1}
-\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
-+
-\frac{ \sigma^2 \parent{\damp+1}^2 - \sigma^2 \omega^2}{2}
-}
-}
-\expfs{
-\tau \lambda
-\parent{
-\frac{p \eta_1 \parent{\eta_1 - \damp -1}}{ \parent{\eta_1 -\damp -1}^2}
-+
-\frac{q \eta_2 \parent{\eta_2 + \damp +1}}{\parent{\eta_2+ \damp +1}^2}
-+1
-}
-}
-\end{align*}
-The term $\expf{- \frac{\omega^2}{2}}$ can be bounded from above
-by an exponential term $\expf{-\gamma \omega}$:
-\begin{align}
-\expf{- \frac{\sigma^2 \omega^2}{2}}
-\label{eq:tailBound}
-& \leq
-\expf{-\gamma \omega+ K}
-\\
-- \sigma^2 \omega^2 + 2 \gamma \omega - 2 K & \leq 0
-\nonumber
-\end{align}
-the condition for the quadratic to have no solutions
-is given by
-\begin{align*}
- \gamma^2 - 2 K  \sigma^2 \leq 0,
-\end{align*}
-setting $K =\frac{1}{2}$, $\gamma = \sigma$
-giving us the bound in eq. \eqref{eq:tailBound}.
-
-This puts us in place to apply theorem 6.1 of \cite{lee2004option}, with
-\begin{align*}
-\Phi \parent{u} &= 
-\frac{
-\expf{
-\tau
-\parent{ 
-\parent{\damp+1}
-\parent{r - \frac{\sigma^2}{2} - \lambda \zeta}
-+
-\frac{ \sigma^2 \parent{\damp+1}^2 }{2}
-}
-}
-\expfs{
-\tau \lambda
-\parent{
-\frac{p \eta_1 \parent{\eta_1 - \damp -1}}{ \parent{\eta_1 -\damp -1}^2}
-+
-\frac{q \eta_2 \parent{\eta_2 + \damp +1}}{\parent{\eta_2+ \damp +1}^2}
-+1
-}
-}}
-{\damp^2 \parent{\damp+1}^2}
-\\
-\gamma & = \sigma
-\end{align*}
-
 \end{document}

### 2016-03-27 tests/HJM/echo_test_cmd.py
first runs to the document running, scripts set

@@ -62,5 +62,5 @@ else:
     TOLs = np.logspace(np.log10(TOL_max),np.log10(TOL_min),TOL_N)
     for TOL in TOLs:
         for i in range(0, realizations):
-            print cmd_multi.format(bayesian=False, tag="HJM_Example_2_newnewrate", TOL=TOL,
+            print cmd_multi.format(bayesian=False, tag="HJM_Example_2_c_rate", TOL=TOL,
                                    seed=np.random.randint(2**32-1))

### 2016-03-27 mimclib/plot.py
Added support for some MIMC plots. The rate lines only work for isotropic beta

@@ -342,13 +342,13 @@ ax is in instance of matplotlib.axes
         xy = [[r.TOL, r.totalTime] for r in runs_data]
     elif work_estimate:
         xy = [[r.TOL, np.sum(r.run.data.M*r.run.Wl_estimate),
-               r.run.Wl_estimate[-1] *
+               np.max(r.run.Wl_estimate) *
                np.maximum(r.run.params.M0, r.run.params.Ca *
                           r.run.all_data.calcVl()[0] * r.TOL**-2.)]
               for r in runs_data]
     else:
         xy = [[r.TOL, np.sum(r.run.data.t),
-               r.run.all_data.calcTl()[-1] *
+               np.max(r.run.all_data.calcTl()) *
                np.maximum(r.run.params.M0, r.run.params.Ca *
                           r.run.all_data.calcVl()[0] * r.TOL**-2.)]
               for r in runs_data]
@@ -475,17 +475,18 @@ ax is in instance of matplotlib.axes
 
 
 def __add_legend(ax, handles=None, labels=None, alpha=0.5,
-                 outside=False, loc='best', *args, **kwargs):
+                 outside=None, loc='best', *args, **kwargs):
     if not handles:
         handles, labels = ax.get_legend_handles_labels()
         if not handles:
             return
-    if outside:
+    if outside is not None and len(handles) >= outside:
         # Shrink current axis by 20%
         box = ax.get_position()
         ax.set_position([box.x0, box.y0, box.width * 0.8, box.height])
-        ax.legend(handles, labels, loc='center left', fancybox=True,
-                  shadow=True, bbox_to_anchor=(1, 0.5)).draggable(True)
+        ax.legend(handles, labels, loc='center left', fancybox=False,
+                  frameon=False, shadow=False,
+                  bbox_to_anchor=(1, 0.5)).draggable(True)
     else:
         ax.legend(handles, labels, loc=loc, fancybox=True,
                   shadow=True).draggable(True)
@@ -516,6 +517,8 @@ def genPDFBooklet(runs_data, fileName=None, exact=None, **kwargs):
         params = next(r.run.params for r in runs_data if r.TOL == maxTOL)
 
     dim = params.dim
+    legend_outside = kwargs.pop("legend_outside", 5)
+
     has_gamma_rate = hasattr(params, 'gamma')
     has_w_rate = hasattr(params, 'w')
     has_s_rate = hasattr(params, 's')
[…]

### 2016-03-27 tests/HJM/HJM.py
rates

@@ -3,6 +3,103 @@
 import numpy as np
 import scipy as sp
 import matplotlib.pyplot as plt
+import scipy.integrate as spint
+
+
+def rateTest2D(Nref=6,M=100,r0=0.05,sig=0.01):
+    
+    """
+    Test the convergence rates in two different dimensions.
+    """
+
+    # First check convergence along first difference
+
+    plotX = range(1,Nref+1)
+    plotYStrong = []
+    plotYWeak = []
+    plotYStrongE = []
+    plotYWeakE = []
+
+    plotRate = lambda ells,rate: 2**(np.array(ells)*(-1.0*rate))
+    
+    for ell in plotX:
+        sample = [hoLeeExample2([[ell,Nref],[ell-1,Nref]]) for foo in range(M)]
+        sample = [abs(foo[1]-foo[0]) for foo in sample]
+        plotYWeak.append(np.mean(sample))
+        plotYWeakE.append(np.sqrt(np.var(sample)/M))
+        sample = [foo**2 for foo in sample]
+        plotYStrong.append(np.mean(sample))
+        plotYStrongE.append(np.sqrt(np.var(sample)/M))
+
+    plotYStrong = np.array(plotYStrong)
+    plotYStrongE = np.array(plotYStrongE)
+    plotYWeak = np.array(plotYWeak)
+    plotYWeakE = np.array(plotYWeakE)
+
+    plt.figure()
+    plt.semilogy(plotX,plotYWeak,'b-')
+    plt.semilogy(plotX,plotYWeak-plotYWeakE,'b--')
+    plt.semilogy(plotX,plotYWeak+plotYWeakE,'b--')
+    plt.semilogy(plotX,0.1*plotRate(plotX,1.0),'r-')
+    plt.grid(1)
+    plt.title('Weak error, $\Delta t$ direction')
+    plt.xlabel('$\ell$')
+    plt.ylabel('$E_w$')
+    plt.savefig('dim_1_weak.pdf')
+
+    plt.figure()
+    plt.semilogy(plotX,plotYStrong,'b-')
+    plt.semilogy(plotX,plotYStrong+plotYStrongE,'b--')
+    plt.semilogy(plotX,plotYStrong-plotYStrongE,'b--')
+    plt.semilogy(plotX,1.0e-3*(plotRate(plotX,1.0))**2,'r-')
+    plt.grid(1)
[…]

### 2016-03-28 tests/HJM/HJM.py
generic HJM template

@@ -116,14 +116,68 @@ def hoLeeExample(inds,t_max=1.0,tau_max=2.0,r0=0.05,sig=0.01,verbose=False):
     thi = lambda tau: 0.1*(1-np.exp(-1*tau))
     f0 = lambda tau: r0-sig*sig*0.5*tau*tau+thi(tau)
 
+    F = lambda x: 1.0-x
+    G = lambda x: 1.0*x
+    Psi = lambda x: 1.0*x
+    U = lambda x: 0.0*x
+
+    d1 = lambda s: sig*sig*s
+
+    drift = lambda s: d1(s)
+
+    v1 = lambda s: sig*np.ones(np.shape(s))
+
+    vols = [v1,]
+
+    identifierString = 'Evaluating the Ho Lee example.\n'
+    identifierString += 'r0: %f, vol %f , t_max %f , tau_max %f .'
+
+    return multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=t_max,tau_max=tau_max,identifierString=identifierString,verbose=verbose)
+    
+def twoFactorGaussianExample(inds,t_max=1.0,tau_max=3.0,b0=0.0759,b1=-0.0439,k=0.4454,a2=0.5,s1=0.02,s2=0.01,K=0.5,verbose=False):
+    
+    '''
+    Compute the two factor Gaussian Example in Beck-Tempone-Szepessy-Zouraris
+    '''
+    
+
+    f0 = lambda tau: b0+b1*np.exp(-1.0*k*tau)
+
+    F = lambda x: np.exp(-1.0*x)
+    G = lambda x: np.fmax(np.exp(-1.0*x)-K)
+    Psi = lambda x: 1.0*x
+    U = lambda x: 0.0*x
+    
+    d1 = lambda s: s1*s1*s
+    d20 = lambda s: np.exp(-0.5*a2*s)
+    d2 = lambda s: 2*s2*s2/a2*d20(s)*(1.0-d20(s))
+    
+    drift = lambda s: d1(s)+d2(s) 
+
+    v1 = lambda s: s1*np.ones(np.shape(s))
+    v2 = lambda s: s2*d20(s)
+
+    vols = [v1,v2]
+
+    identifierString = 'Evaluating the Two Factor Gaussian example.\n'
+    identifierString += 's1: %f, s2: %f, b0: %f, tau_max: %f, t_max: %f\n'%(s1,s2,b0,tau_max,t_max)
[…]

### 2016-03-28 tests/HJM/echo_test_cmd.py

@@ -46,7 +46,7 @@ args = parser.parse_known_args()[0]
 # Note that 0.6931 is np.log(2)
 base = "run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
 -qoi_seed {seed} \
--mimc_dim 2 -mimc_w 2 2 -mimc_s 0.5 1 -mimc_gamma 1 1 -mimc_beta 2 2 \
+-mimc_dim 2 -mimc_w 1 2 -mimc_s 1 2 -mimc_gamma 1 1 -mimc_beta 2 2 \
 -mimc_bayesian {bayesian} "
 
 cmd_multi = "python " + base + "-mimc_verbose False -db True -db_tag {tag} " + " -db_host {} ".format(args.db_host)
@@ -62,5 +62,5 @@ else:
     TOLs = np.logspace(np.log10(TOL_max),np.log10(TOL_min),TOL_N)
     for TOL in TOLs:
         for i in range(0, realizations):
-            print cmd_multi.format(bayesian=False, tag="HJM_Example_2_c_rate", TOL=TOL,
+            print cmd_multi.format(bayesian=False, tag="HJM_Example_saferate_2dim", TOL=TOL,
                                    seed=np.random.randint(2**32-1))

### 2016-04-02 mimclib/plot.py
Another small bug in plotting higher moments

@@ -242,8 +242,8 @@ def __calc_moments(runs_data, seed=None, direction=None):
     dim = runs_data[0].run.data.dim
     seed = np.array(seed) if seed is not None else np.zeros(dim, dtype=np.uint32)
     direction = np.array(direction) if direction is not None else np.ones(dim, dtype=np.uint32)
-
-    psums = np.zeros((0, 2))
+    moments = runs_data[0].run.data.psums.shape[1]
+    psums = np.zeros((0, moments))
     Tl = np.zeros(0)
     M = np.zeros(0)
     for curRun in runs_data:
@@ -258,10 +258,10 @@ def __calc_moments(runs_data, seed=None, direction=None):
             cur = cur + direction
         L = len(inds)
         if L > len(M):
-            psums.resize((L, 2), refcheck=False)
+            psums.resize((L, moments), refcheck=False)
             M.resize(L, refcheck=False)
             Tl.resize(L, refcheck=False)
-        psums[:L, :] += curRun.run.data.psums[inds, :2]
+        psums[:L, :] += curRun.run.data.psums[inds, :]
         M[:L] += curRun.run.data.M[inds]
         Tl[:L] += curRun.run.data.t[inds]
 

### 2016-04-02 tests/gbm/echo_test_cmd.py
add moments

@@ -14,7 +14,7 @@ parser.add_argument("-db_host", type=str, action="store",
 args = parser.parse_known_args()[0]
 
 base = "run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
--qoi_sigma 0.1 -qoi_mu 1 -qoi_seed {seed} \
+-qoi_sigma 0.1 -qoi_mu 1 -qoi_seed {seed} -mimc_moments 4 \
 -mimc_dim 1 -mimc_w 1 -mimc_s 1 -mimc_gamma 1 -mimc_beta 2 \
 -mimc_theta {theta} -mimc_bayesian {bayesian} "
 
@@ -28,6 +28,6 @@ else:
     TOLs = 0.1*np.sqrt(2.)**-np.arange(0., 10.)
     for TOL in TOLs:
         for i in range(0, realizations):
-            print cmd_multi.format(bayesian=False, tag="GBM_test", TOL=TOL,
+            print cmd_multi.format(bayesian=False, tag="GBM_test_kurtosis", TOL=TOL,
                                    seed=np.random.randint(2**32-1), theta="0.2")
 

### 2016-04-02 tests/Schoebel_Zhu_Hull_White/manuscript.tex

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{textcomp}
\usepackage{mathrsfs }
\usepackage[section]{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{patchplots}
\usepackage{subfigure}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{multirow}

%line suggested by my compiler (Fabian)
\pgfplotsset{compat=1.9}

% put figures at the end of the document, while I am writing is very annoying to have the figures in the middle of the text! 
% \usepackage[nomarkers,figuresonly]{endfloat}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\input{macros.tex}

\title[MLMC for timer options]{Multi-Level Monte Carlo evaluation of timer options.}
\author{Juho H\"app\"ol\"a}
\address{\kaustaddress}
\email{juho.happola@iki.fi}

\date{\today}    
                                    % Activate to display a given date or no date
\begin{document}

\begin{abstract}
Abstract here!!!
\end{abstract}

\maketitle
%\tableofcontents

\section{Introduction}

Timer options are an interesting class of derivative assets that have gained some popularity recently. The central feature of timer options is their stochastic exercise time.

Let the asset price be modelled by the following SDE
\begin{align}
dX_t &= r_t X dt + \beta \parent{t,X_t, \sigma_t} dW_t
\nonumber \\
X_0 &= x_0
\end{align}
with the volatility $\sigma_t$, and interest rate $r_t$
given by adapted processes.

Unlike in plain vanilla options, we are interested in evaluating
the expected value of a payoff functional $g$
\begin{align}
\expp{e^{-\int_0^\tau r_s ds} g \parent{X_\tau}}
\end{align}
with the hitting time $\tau$ being defined as
\begin{align}
\label{eq:hittingTimeDefinition}
\tau \equiv \mathop{\mathrm{inf.}}_{\iota}
\sset{\iota: \int_0^\iota \sigma_s^2 ds \geq V}.
\end{align}

We note, that for processes that have almost surely $\sigma_s=\sigma$,
the option prices reduce to those of a fixed-maturity plain-vanilla option.

The case of stochastic volatility has been studied earlier
in the Forward-Euler Monte Carlo setting by \cite{bernard2011pricing}. For a selected class of stochastic
volatility models, \cite{zeng2015fast} presented a fast Hilbert transform based solution.

\section{Model}

For a rather general study framework of studying the pricing of timer options, we choose the Sch\"obel-Zhu-Hull-White model introduced in \cite{grzelak2012extension}, with stochastic interest rates and volatility:
\begin{align}
d S_t &=& r_t S_t dt &+& \sigma_t^p S_t dW_t^x 
\nonumber
\\
d r_t &=& \lambda \parent{\theta_t -r_t} dt  &+& \eta dW_t^r
\\
d \sigma_t &=& \kappa \parent{\overline \sigma - \sigma_t} dt
 &+& \gamma \sigma^{1-p} dW_t^\sigma
\nonumber,
\end{align}
where the three Brownian motions $W^x$, $W^r$ and $W^\sigma$
are correlated through the constant instantaneous correlation matrix
$\Sigma$.

\section{Numerical example}

To set a reasonable test case in which to test the convergence of
Multi-Level Monte Carlo, we may select the a reasonable asset volatility
to twenty per cent:
\begin{align*}
\overline \sigma = \sigma_0 = 0.2.
\end{align*}
Setting $p=1$, the
long-term variance of volatility is given by
\begin{align*}
\frac{\gamma^2 }{2 \kappa} = \frac{\bar \sigma}{4},
\end{align*}
we set
\begin{align*}
\kappa &= 2, \\
\bar \sigma &= \gamma.
\end{align*}
Likewise, for the interest rate, we set 
\begin{align*}
\theta_t &= 0.01, \\
\lambda &=1,
\eta = 0.1.
\end{align*}

We may select the maturity time $T=\frac{1}{12}$ and which
gives us
\begin{align*}
V = \frac{\bar \sigma^2}{12}.
\end{align*}

For the discretisation, we select the drift-implicit Euler method
which guarantees stability for longer time steps when using longer
time steps for the Ornstein-Uhlenbeck process.
\begin{align}
\sigma_{n+1} &= \frac{\kappa \overline \sigma}{1+\Delta t}
+ \gamma \sigma_{n}^{1-p} \Delta W_n^{\sigma}
\\
r_{n+1} &= \frac{\lambda \overline \theta_n}{1+ \Delta t}
+ \eta \Delta W_n^r
\\
\log X_{n+1} &=
 \parent{r - \frac{\sigma^{2p}_{n+1}}{2}} \Delta t
+ \sigma_n^{p} \Delta W_n^x .
\end{align}
To evaluate the accumulated variance, we discretise the integral in eq. \eqref{eq:hittingTimeDefinition} as
\begin{align}
\nu_{n+1} = \nu_n + \sigma_n^2 \Delta t,
\end{align}
and similarly for the discounting process
\begin{align}
d_{n+1} = d_n + r_n \Delta t.
\end{align}

For the correlation matrix, we initially set $\Sigma = \mathbf 1_3$.

Given a multi-level index $\ell \geq 0$, we set
\begin{align*}
\Delta t^{\ell} = 2^{-\ell-5} T,
\end{align*}
i.e. We start with a setting that for $\gamma=0$ case
has $2^5$ time steps and increase the number of time steps geometrically.

We simulate realisations $\overline X^\ell$, $\overline X^{\ell-1}$ for
$1<\ell<L$ in pairs using the same brownian paths and run the discretised Euler integration until $n^*,k^*$ $\nu_{n^*}^\ell \geq V$ and $\nu_{k^*}^{\ell+1} \geq V$. Finally, we form the estimator for the option price as
\begin{align}
\mathcal A \parent{g} = \ssum{m=0}{M_0} \frac{e^{d_{n^*}^{\parent{0}}} g \parent{X_{k^*}} }{M_0}
+
\ssum{\ell=1}{L} \ssum{m=0}{M_\ell}
\frac{e^{d_{n^*}^{\parent{0}}} g \parent{X_{n^*}} - e^{d_{k^*}^{\parent{0}}} g \parent{X_{k^*}} }{M_0}
\end{align}

\section{Convergence results (???)}

\bibliographystyle{agsm}
\bibliography{references}

\appendix

\end{document}

### 2016-04-25 mimclib/plot.py
Fixed theory curve fitting for number of levels

@@ -15,17 +15,19 @@ def public(sym):
 
 @public
 class FunctionLine2D(plt.Line2D):
-    def __init__(self, fn, data=None, **kwargs):
+    def __init__(self, fn, log_data=True, data=None, **kwargs):
         self.flip = kwargs.pop('flip', False)
         self.fn = fn
         if data is not None:
             x = np.array([d[0] for d in data])
             y = np.array([d[1] for d in data])
             if len(x) > 0 and len(y) > 0:
-                const = [np.mean(y/fn(x)), 0]
-                # const = np.polyfit(fn(x), y, 1)
-                # print(const, np.mean(y/fn(x)))
-                self.fn = lambda x, cc=const, ff=fn: cc[0] * ff(x) + cc[1]
+                if log_data:
+                    const = [np.mean(y/fn(x)), 0]
+                    self.fn = lambda x, cc=const, ff=fn: cc[0] * ff(x) + cc[1]
+                else:
+                    const = [np.mean(y-fn(x)), 0]
+                    self.fn = lambda x, cc=const, ff=fn: ff(x) + cc[0]
 
         super(FunctionLine2D, self).__init__([], [], **kwargs)
 
@@ -813,6 +815,7 @@ def genPDFBooklet(runs_data, fileName=None, exact=None, **kwargs):
             label = r'${}\log\left(\textrm{{TOL}}^{{-1}}\right)$'.format(formatPower(rate))
             ax.add_line(FunctionLine2D(lambda x, r=rate: -rate*np.log(x),
                                        data=line_data,
+                                       log_data=False,
                                        linestyle='--', c='k',
                                        label=label))
     except:

### 2016-04-25 tests/HJM/manuscript.tex
first runs to the document running, scripts set
generic HJM template
add empirical rates to notes

@@
 \usepgfplotslibrary{patchplots}
-\usepackage{subfigure}
+%\usepackage{subfigure}
+\usepackage{subcaption}
+
 \usepackage{hyperref}
@@
 \begin{abstract}
-Abstract here!!!
+Notes of MIMC and MLMC evaluation of HJM models.
 \end{abstract}
@@
 
-We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models.
+\section{Intro}
+
+We are interested in analysing the Forward Euler Finite Difference evaluation of forward curve models of HJM models.
 The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.
@@
 In order to define a mesh for the numerical solution of the problem, let us define $\bar \ell \in \mathbb{N}^3$.
+
+Let our quantity of interest be defined as the expectation of a functional $\mathcal F$
+of the forward curve:
+\begin{align*}
+\mathcal F \parent f =& F \parent{\int_0 ^{t_{max}} f \parent{s,s} ds}
+G \parent{ \int_{\tau_a}^{\tau_{max}} \Psi \parent{f \parent{t_{max},\tau}} d\tau }
+\\
+&+
+\int_{0}^{t_{max}} F \parent{\int_0^{s} f \parent{s',s'} ds'} U \parent{f\parent{s,s}} ds.
+\end{align*}
 
@@
 \bar \gamma &= [1,1]^T , \\
-\bar s &= [2^{-1},1], \\
-\bar w & = [2,2].
+\bar s &= [1,2], \\
+\bar w & = [1,2].
 \end{align}
@@
 
-\section{Multi-Index Monte Carlo}
+\section{Numerical tests}
 
-\section{Multi-Level Monte Carlo}
+For the simplest possible example, let our model be Gaussian with
+the following exponential covariance structure
+\begin{align}
+\expp{df \parent{t,\tau} df \parent{t, \tau'}}
+=
+\sigma^2 \expf{- \kappa \absval{\tau-\tau'}} dt.
+\end{align}
+
+Let our first test case be arguably simplest infinite dimensional one possible,
+\begin{align*}
+F \parent{x} &= 1-x,\\
+U \parent{x} &= 0, \\
+G \parent{x} &= x, \\
+\Psi \parent{x} &= x, \\
+f \parent{0,\tau} &= 0, \\
+\kappa &= 2.
+\end{align*}
+
+In \ref{fig:empInfDimRates} we get see that we have rate one for both 
+the weak and the strong rate along both dimensions for this text example.
+
+\begin{figure}
+\includegraphics[scale=0.7]{figure_1.pdf}
+\caption{Sample realisation of a forward curve with the discretisation level 
+$\overline \ell = [5,5,5]^T$.}
+\end{figure}
+
+It remains to deduce the cost of simulating the cost of a realisation. In order to
+simulate Gaussian random variables one needs the Cholesky decomposition of the covariance matrix.
+The cost to generate this $N_\tau \times N_\tau$-matrix is $N_\tau^3$, with $N_\tau = 2^{\ell_3} 2^{\ell_2}$.
+Using the Cholesky decomposition, the cost to take a a time step becomes $N_\tau^2$.
+Putting all together, the cost to generate $M$ Monte Carlo realisations with $N_t$ time steps and $N_\tau$ discretisation steps along the maturity direction becomes
+\begin{align*}
+C \parent{\Delta t, \Delta \tau, M} = N_\tau^3 + M N_\tau^2 N_t.
+\end{align*}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{dim_1_strong.pdf}
+        %\caption{A gull}
+        \label{fig:gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{dim_1_weak.pdf}
+        %\caption{A tiger}
+        \label{fig:tiger}
+    \end{subfigure}
+    \\
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+    %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{dim_2_strong.pdf}
+        %\caption{A mouse}
+        \label{fig:mouse}
+    \end{subfigure}
+        \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{dim_2_weak.pdf}
+        %\caption{A mouse}
+        \label{fig:mouse}
+    \end{subfigure}
+    \caption{\label{fig:empInfDimRates}Empirical strong and weak convergence rate for the $\Delta t$ and $\Delta \tau$ discretisation errors.}
+\end{figure}
 

### 2016-04-25 tests/HJM/HJM.py

@@ -4,8 +4,26 @@ import numpy as np
 import scipy as sp
 import matplotlib.pyplot as plt
 import scipy.integrate as spint
+import scipy.linalg as spal
 import time
+import matplotlib as mpl
+from mpl_toolkits.mplot3d import Axes3D
 
+def hash(x):
+    return '%f_%f_%f_%f'%(x[0],x[-1],x[1]-x[0],x[-1]-x[-2])
+
+def expCovar(xs,kappa,chol=False):
+    '''
+    Given a spatial mesh x, return an exponential
+    covariance matrix with inverse correlation length
+    kappa.
+    '''
+
+    x,y = np.meshgrid(xs,xs)
+    rv = np.exp(-1.0*kappa*abs(x-y))
+    if chol:
+        return spal.cholesky(rv)
+    return rv
 
 def hoLeeExample3(inds,t_max=1.0,tau_max=2.0,r0=0.05,sig=0.01,verbose=False):
     return hoLeeExample([[foo[0]]*3 for foo in inds],t_max=t_max,tau_max=tau_max,r0=r0,sig=sig,verbose=verbose)
@@ -71,6 +89,158 @@ def twoFactorGaussianExample(inds,t_max=1.0,tau_max=3.0,b0=0.0759,b1=-0.0439,k=0
 
     return multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=t_max,tau_max=tau_max,identifierString=identifierString,verbose=verbose)    
 
+def infDimHjmModel(inds,F,G,U,Psi,f0,kappa,t_max=1.0,tau_max=2.0,identifierString='Infinite HJM Model',verbose=False,maxLev=30):
+
+    '''
+    Solve an infinite-dimensional HJM model, crude example
+    '''
+
+    ts = [time.time(),]
+
+    if verbose:
+        print('Evaluating the Two Factor Gaussian example.')
+        print(identifierString)
+        for ind in inds:
+            print(ind)
+
+    # largest values of the discretisation numbers                                                                           
+    N_t = max([foo[0] for foo in inds])
+    N_tau_1 = max([foo[1] for foo in inds])
+    N_tau_2 = max([foo[2] for foo in inds])
+
+    if N_t+max(N_tau_1,N_tau_2) > maxLev:
[…]

### 2016-04-25 tests/Schoebel_Zhu_Hull_White/szhw.cpp

@@ -3,7 +3,7 @@
 #include <math.h>
 #include <iostream>
 #include <random>
-#include <quadmath.h>
+//#include <quadmath.h>
 
 #ifndef Pi
 #define Pi 3.141592653589793238462643

### 2016-05-04 tests/HJM/manuscript.tex
three and three halves weak rate N_f

--- 2016-03-27 tests/HJM/manuscript.tex
@@
 
-We are interested in analysing the Forward Euler Finite Element evaluation of forward curve models of HJM models.
-The main approach was represented with a convergence analysis in \cite{bjork2013monte}, whose notation we largely follow.
+\section{HJM model in Fourier space}
 
-We are focused on the evolution of the forward curve $f \parent{t , \tau}$ for $t \in [0,t_{max}]$, $\tau \in [0,\tau_{max}]$ with $\tau_{max}>t_{max}$.
-In order to define a mesh for the numerical solution of the problem, let us define $\bar \ell \in \mathbb{N}^3$.
+Let us focus on the following HJM type SDE:
+\begin{align}
+d f \parent{t,\tau} &= \alpha \parent{t,\tau} dt + \beta \parent{t,\tau} d \overline W \parent t \\
+f \parent{0,\tau} &= f_0 \parent \tau ,
+\end{align}
+with the diffusion term given as the infinite-dimensional expansion of
+eq. (1.1) in \cite{bjork2013monte}:
+\begin{align}
+\beta \parent{t, \tau} d \overline W \parent t = \ssum{n=0}{\infty} c_n \parent{\sin \parent{\frac{n \pi \tau}{L}} + \cos \parent{\frac{n \pi \tau}{L}}} dW_n \parent t , 
+\end{align}
+with $W_n$ independent Brownian motions and $L>>1$. This gives the following Fourier decomposition for the covariance
+of interests:
+\begin{align}
+&\expp{ \beta \parent{t, \tau_1} d \overline W \parent t \beta \parent{t, \tau_2} d \overline W \parent t}
+\\
+= & \ssum{n=0}{\infty} c_n^2 \parent{ \cos \parent{\frac{n \pi \parent{\tau_1-\tau_2 }}{L}}}
+\end{align}
 
-Let us first discretise the time interval as
+To keep the exponential HJM model risk neutral, we fix the drift of the equation as
 \begin{align*}
-\Delta t_{\bar \ell} &= 2^{-\ell_1} t_{max},
+\alpha \parent{t, \tau} 
+=&
+c_0^2 \parent{\tau-t}
 \\
-t_{\bar \ell, n} &= n \Delta t_{\bar \ell}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_1}].
-\end{align*}
-Since there is an additional requirement that $t_{max}$ be in
-the mesh of maturities, one point in the mesh in $\tau \in [0,\tau_{max}]$. We let $\ell_1$ and $\ell_2$ define the number of discretisation points to the right and to the left of the fixed
-point, respectively.
-\begin{align*}
-\Delta \tau_{\bar \ell,1} &= 2^{-\ell_3} t_{max},  \\
-\Delta \tau_{\bar \ell,2} &= 2^{-\ell_2} \parent{\tau_{max}-t_{max}}, \\
-\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [0,1,2,\cdots, 2^{\ell_3}] \\
-\tau_{\bar \ell, n} &= n \Delta \tau t_{\bar \ell,1}, ~~~ n \in [2^{\ell_3},2^{\ell_3}+1,\cdots, 2^{\ell_3}+2^{\ell 2}].
+&+
+\ssum{n=1}{\infty }
+\frac{c_n^2}{n}
+\sin \parent{\frac{n \pi \tau}{L}} \parent{\cos \parent{\frac{n \pi t}{L}} -  \sin \parent{\frac{n \pi t}{L}}}
+\\
+&+
+\ssum{n=1}{\infty }
+\frac{c_n^2}{n}
+\cos \parent{\frac{n \pi \tau}{L}} \parent{\cos \parent{\frac{n \pi t}{L}} -  \sin \parent{\frac{n \pi t}{L}}}
+\\
+&+
+\ssum{n=1}{\infty }
+\frac{c_n^2}{n}
+\cos \parent{\frac{2n \pi \tau}{L}}.
 \end{align*}
 
-Using this notation, we have the following rates of convergence, following the Multi-Index notation:
+The solution to the SDE can be written as
+\begin{align*}
+f \parent{t,\tau} - f_0 \parent \tau
+= \tilde f \parent{t,\tau}
++ \ssum{n=1}{N} b_n \parent t \cos \parent{\frac{n \pi \tau}{L}} + a_n \parent t \sin \parent{\frac{n \pi \tau}{L}},
+\end{align*}
+with
 \begin{align}
-\bar \beta &= [2,2,2]^T \\
-\bar \gamma &= [1,1,1]^T .
-\end{align} 
-Choosing the Simpson method for integration, we have the order of quadrature $p_Q=3$, which brings the quadrature error to 
-\begin{align}
-\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}} \leq C \parent{\Delta \tau_{\bar \ell}}^{2 \kappa p_Q} = 
-\tilde C 2^{-6 \kappa \mathrm{min} \parent{\ell_3,\ell_2}}.
-\end{align}
-On the other hand, the error of the simpson method is the sum of errors for $\tau < t_{max}$ and $\tau \leq t_{max}$.
-The error for these two domains are $\bigo{2^{-2 \ell_2}}$ and $\bigo {2^{-2 \ell_3}}$, respectively.
-Combining together, we get the weak and the strong rate
-for the Simpson quadrature:
-\begin{align}
-\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2}} \leq&  \tilde C_Q 2^{-6 \mathrm{min} \parent{\ell_3,\ell_2}}
+\tilde f \parent{t,\tau} =& c_0^2 t \parent{\tau-\frac{t}{2}},
 \\
-\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}} \leq& C_Q 2^{-2 \mathrm{min} \parent{\ell_2,\ell_3}}.
+a_n \parent{t} =& \frac{c_n^2 L^2}{\pi^2 n^2} \parent{\cos \parent{\frac{n \pi t}{L}} + \sin \parent{\frac{n \pi t}{L}}-1 } + c_n W_n \parent{t}
+\\
+b_n \parent{t} =& a_n \parent{t} -  \frac{\mathbf{1}_{\frac{n}{2} \in \mathbb Z_+}  c_n^2 L t}{\pi n}.
 \end{align}
 
+The solution above lends itself to approximate solutions to the quantity of interest.
+First, the discount factor can be approximated as:
+\begin{align*}
+&\int_0^{t_{max}} f \parent{s,s} ds
+\\
+\approx& \Delta t \ssum{n=1}{N_t} \parent{ f_0 \parent{t_n} + \tilde f \parent{t_n, t_n} +
+\ssum{k=1}{N_f}
+b_k \parent {t_n} \cos \parent{\frac{k \pi t_n}{L}} + a_k \parent {t_n} \sin \parent{\frac{k \pi t_n}{L}}
+}
+\\
+=& F_{N_t,N_f}
+,
+\end{align*}
+with $t_n = \frac{nT}{N_t}$.
+This approximation gives the follwoing approximation error
+in $N_f$:
+\begin{align*}
+&\int_0^{T} f \parent{s,s} ds -\lim_{N_t \rightarrow \infty} F_{N_t,N_f}
+\\
+=& \int_0^T \ssum{k=1}{N_f}
+\frac{c_k^2 L^2}{k^2 \pi^2}
+\parent{1+ \sin \parent{\frac{2 k \pi s}{L}}} ds
+\\
+&+
+ \int_0^T \ssum{k=1}{N_f} c_k W_k \parent s
+\parent{\sin \parent{\frac{ k \pi s}{L}} + \cos \parent{\frac{ k \pi s}{L}}}
+ds
+\\
+&-
+\int_0^T \ssum{k=1}{N_f}
+\frac{c_k^2 L}{\pi k} \cos \parent{\frac{ k \pi s}{L}},
+ds,
+\end{align*}
+
+Similarly, one may the underlying part of the 
+payoff functional
+\begin{align*}
+&\int_{\tau_1}^{\tau_2}
+f \parent{T,\tau} d\tau 
+\\
+\approx & \int_{\tau_1}^{\tau_2}
+f_0 \parent \tau + \tilde f \parent{t_n, t_n} +
+\ssum{k=1}{N_f}
+b_k \parent {T} \cos \parent{\frac{k \pi \tau}{L}} + a_k \parent {T} \sin \parent{\frac{k \pi \tau}{L}}
+d \tau.
+\\
+= &+ \int_{\tau_1}^{\tau_2}
+f_0 \parent \tau d \tau + \frac{c_0 T}{2} \parent{\tau_2^2 - \tau_2 +\tau_1 - \tau_1^2} 
+\\
+&+
+\ssum{k=1}{N_f} 
+\frac{b_k \parent T L}{k \pi} 
+\parent{\sin \parent{\frac{k \pi \tau_2}{L}}  - \sin \parent{\frac{k \pi \tau_1}{L}} }
+\\
+&-
+\ssum{k=1}{N_f} 
+\frac{a_k \parent T L}{k \pi} 
+\parent{\cos \parent{\frac{k \pi \tau_2}{L}}  - \cos \parent{\frac{k \pi \tau_1}{L}} }
+\\
+=& \Psi_{N_f} .
+\end{align*}
+The corresponding empirical rates for the approximations of the quantity of interest
+\begin{align*}
+\mathcal G \parent f =
+\expp{\expf{-\int_0^{t_{T}} f \parent{s,s} ds}\expf{-\int_{\tau_1}^{\tau_2} f \parent{T,\tau} d\tau}}
+\approx
+\expp{
+F_{N_t,N_f}
+\Psi_{N_f} }
+\end{align*}
+are presented in fig. \ref{fig:empInfDimRates}
+
+Setting $\ell = \parent{\ell_1 , \ell_2}$, we may define a Monte
+Carlo estimator using $m$ independent realisations of $\Psi_{N_f}$ and $F_{N_t,N_f}$:
 \begin{align}
-\expp{\absval{\mathcal G\parent{\bar{\bar g}} - \mathcal G \parent{g}}^{2 \kappa}}
-\end{align}
-using Theorem 3.3 of \cite{bjork2013monte}.
-
-On the other hand, theorem 3.2 and equation (3.54) states that
-\begin{align}
-\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{2 \kappa}  } ^{\frac{1}{2 \kappa}} \leq C_{5}^{CV} \parent{ \parent{\Delta t}^{\frac{1}{2}} + \Delta \tau }.
+\mathcal A_{\ell_1, \ell_2} 
+\equiv
+\ssum{m=1}{M}
+\frac{F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} }\parent m}{M}.
 \end{align}
 
-The weak rate for the discretisation error can be bounded by eq. (4.6):
-\begin{align}
-\expp{ \absval{ \mathcal G \parent{g} - \mathcal G \parent{\bar{\bar g}} }^{}  }
-= 
- \bigo{\parent{\Delta t}^2 + \parent{\Delta \tau}^2}
- = \bigo{2^{-2 \ell_1} + 2^{-2 \mathrm {min} \parent{\ell_2,\ell_3} }}
-\end{align}
+Similarly, we may extend the above to a MLMC estimator as
+\begin{align*}
+\mathcal{A}_{ML} =& \ssum{m=0}{M_0} \frac{F_{2 C_1 ,C_2 } \Psi_{C_2 }\parent m}{M_0}
+\\
+&+
+\ssum{\ell_1=1}{L}
+\ssum{m=0}{M_{\ell_1}} \frac{\parent{ F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_1}} \Psi_{C_2 2^{\ell_1} } - F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_1-1}} \Psi_{C_2 2^{\ell_1} }} \parent m }{M_{\ell_1}},
+\\
+\equiv& \ssum{\ell_1=0}{L} \ssum{m=0}{M_{\ell_1}} \frac{\Delta_{\ell_1} \parent m}{M_{\ell_1}}
+\end{align*}
+and, into a MIMC estimator through defining the appropriate two-dimensional difference operators $\Delta_{\ell_1,\ell_2}$ and a downward-closed index-set $L_K=\sset{\parent{\ell_1,\ell_2}\in \mathbb Z_+^2: I \parent{\ell_1,\ell_2}<K}$:
+\begin{align*}
+\mathcal A_{MI} =& \ssum{\ell \in L_K}{} \ssum{m=1}{M_{\ell}} \frac{\Delta_{\ell_1,\ell_2} \parent m}{M_\ell}
+\\
+\Delta_{\ell_1,\ell_2}
+\equiv&
+F_{C_1 2^{\ell_1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} } - \mathbf{1}_{\ell_1>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2-1}} \Psi_{C_2 2^{\ell_2-1} }
+\\
+& -\mathbf{1}_{\ell_2>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2}} \Psi_{C_2 2^{\ell_2} }
++ \mathbf{1}_{\ell_1,\ell_2>1} F_{C_1 2^{\ell_1-1}+1,C_2 2^{\ell_2-1}} \Psi_{C_2 2^{\ell_2-1} }.
+\end{align*}
 
-Due to the fact that the discretisation errors relating to $\tau$
-depend on the minimum of $\ell_2$ and $\ell_3$, it seems natural
-to assume $\ell_3=\ell_2$. This gives us a two-dimensional
-Multi-Index Monte Carlo formulation with the following effective parameters
-\begin{align}
-\bar \beta &= [2,2]^T ,\\
-\bar \gamma &= [1,1]^T , \\
-\bar s &= [2^{-1},1], \\
-\bar w & = [2,2].
-\end{align}
-
-On the other hand, one may set $\ell_1=\ell_2=\ell_3=\ell$
-and arrive at a one-dimensional MLMC formulation with a weak rate of $2$, strong rate of $1$, and computational cost per realisation of $W_\ell=2^{2\ell}$.
-In the sequel, we will present the results of these two variations for comparison.
-
-\section{Multi-Index Monte Carlo}
-
-\section{Multi-Level Monte Carlo}
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{nfstrong.pdf}
+        %\caption{A gull}
+        \label{fig:gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{nfweak.pdf}
+        %\caption{A tiger}
+        \label{fig:tiger}
+    \end{subfigure}
+    \\
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+    %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{ntstrong.pdf}
+        %\caption{A mouse}
+        \label{fig:mouse}
+    \end{subfigure}
+        \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{ntweak.pdf}
+        %\caption{A mouse}
+        \label{fig:mouse}
+    \end{subfigure}
+    \caption{\label{fig:empInfDimRates}Empirical strong and weak convergence rate for
+    the $N_t$ and $N_f$.}
+\end{figure}
 

### 2016-05-04 tests/HJM/HJM.py

@@ -8,6 +8,7 @@ import scipy.linalg as spal
 import time
 import matplotlib as mpl
 from mpl_toolkits.mplot3d import Axes3D
+import itertools
 
 def hash(x):
     return '%f_%f_%f_%f'%(x[0],x[-1],x[1]-x[0],x[-1]-x[-2])
@@ -88,6 +89,123 @@ def twoFactorGaussianExample(inds,t_max=1.0,tau_max=3.0,b0=0.0759,b1=-0.0439,k=0
     identifierString += 'k: %f, a2: %f, K: %f, b1: %f'%(k,a2,K,b1)
 
     return multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=t_max,tau_max=tau_max,identifierString=identifierString,verbose=verbose)    
+
+def cCovExp(L,K,n):
+    return 2*K*((1-np.exp(-abs(K*L))*((-1)**n))/(K*K+n*n*np.pi*np.pi/L/L))/L
+
+def fourierExample(inds,cfun,L=50.0,tau1=1.0,tau2=2.0,t_max=1.0,verbose=False):
+
+    ts = [time.time(),]
+
+    w = lambda k: np.pi*k/L
+    plotN = 200
+
+    # largest values of the discretisation numbers                                                                           
+    N_t = max([foo[0] for foo in inds])
+    N_f = max([foo[1] for foo in inds])
+
+    N_t = 2**(N_t)+1
+    N_f = 2**(N_f)
+
+    times = np.linspace(0,t_max,N_t)
+    dt = times[1]-times[0]
+
+    Wt = sp.randn(N_t,N_f)*np.sqrt(dt)
+    Wt[0,:] *= 0
+    Wt = np.cumsum(Wt,axis=0)
+
+    if verbose:
+        print('W mean and variance: %f, %f'%(np.mean(Wt[-1,:]),np.var(Wt[-1,:])))
+    
+    consts = np.zeros(N_t)
+    ans = np.zeros((N_t,N_f-1))
+    bns = np.zeros((N_t,N_f-1))
+
+    for row in range(0,N_t):
+        consts[row] = Wt[row,0]
+        for col in range(0,N_f-1):
+            k = col+1
+            t = times[row]
+            #ans[row,col] = ((cfun(k)/w(k))**2)*(np.cos(t*w(k))+np.sin(t*w(k))-1.0)
+            ans[row,col] += cfun(k)*Wt[row,k]
+            bns[row,col] = ans[row,col]
[…]

### 2016-05-04 tests/HJM/echo_test_cmd.py

@@ -2,22 +2,6 @@
 import numpy as np
 import argparse
 
-'''
-A script to generate a command for doing a MIMC run.
-
-Usage: ./echo_test_cmd.py
-prints the command.
-
-To run a single mimc run, type
-./echo_test_cmd.py |bash
-
-Or multiple in parallel:
-./echo_test_cmd.py -multi |parallel j x
-where x is the number of processes to run in parallel.
-
-'''
-
-
 parser = argparse.ArgumentParser(add_help=True)
 parser.register('type', 'bool',
                 lambda v: v.lower() in ("yes", "true", "t", "1"))
@@ -26,41 +10,24 @@ parser.add_argument("-multi", type="bool", action="store",
 parser.add_argument("-db_host", type=str, action="store",
                     default="localhost",
                     help="True output a single command")
-parser.add_argument("-TOL_max", type=float, action="store",
-                    default=0.001,
-                    help="True output a single command")
-parser.add_argument("-TOL_min", type=float, action="store",
-                    default=0.00001,
-                    help="True output a single command")
-parser.add_argument("-TOL_N", type=int, action="store",
-                    default=8,
-                    help="True output a single command")
-parser.add_argument("-N_iter", type=int, action="store",
-                    default=35,
-                    help="True output a single command")
-
-
 
 args = parser.parse_known_args()[0]
 
-# Note that 0.6931 is np.log(2)
 base = "run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
--qoi_seed {seed} \
--mimc_dim 2 -mimc_w 1 2 -mimc_s 1 2 -mimc_gamma 1 1 -mimc_beta 2 2 \
--mimc_bayesian {bayesian} "
+-qoi_sigma 0.1 -qoi_mu 1 -qoi_seed {seed} -mimc_moments 4 \
+-mimc_dim 2 -mimc_w 1 1 -mimc_s 1 1 -mimc_gamma 1 2 -mimc_beta 2 2 \
+-mimc_theta {theta} -mimc_bayesian {bayesian} "
 
 cmd_multi = "python " + base + "-mimc_verbose False -db True -db_tag {tag} " + " -db_host {} ".format(args.db_host)
 cmd_single = "python " + base + " -mimc_verbose True -db False "
 
 if not args.multi:
-    print(cmd_single.format(seed=0, bayesian=False, TOL=args.TOL_min))
[…]

### 2016-05-04 tests/HJM/run.py

@@ -5,28 +5,9 @@ from __future__ import print_function
 import warnings
 import os.path
 import numpy as np
-import mimclib.mimc as mimc
-import mimclib.db as mimcdb
-
-import HJM
-
-warnings.formatwarning = lambda msg, cat, filename, lineno, line: \
-                         "{}:{}: ({}) {}\n".format(os.path.basename(filename),
-                                                   lineno, cat.__name__, msg)
-#warnings.filterwarnings('error')
-
+import HJM 
 
 def addExtraArguments(parser):
-    parser.register('type', 'bool',
-                    lambda v: v.lower() in ("yes", "true", "t", "1"))
-    parser.add_argument("-db_user", type=str,
-                        action="store", help="Database User")
-    parser.add_argument("-db_host", type=str, default='localhost',
-                        action="store", help="Database Host")
-    parser.add_argument("-db_tag", type=str, default="NoTag",
-                        action="store", help="Database Tag")
-    parser.add_argument("-db", type='bool', default=False,
-                        action="store", help="Save in Database")
     parser.add_argument("-qoi_sigma", type=float, default=1.,
                         action="store", help="Volatility in GBM")
     parser.add_argument("-qoi_mu", type=float, default=1.,
@@ -35,50 +16,20 @@ def addExtraArguments(parser):
                         action="store", help="Final time in GBM")
     parser.add_argument("-qoi_S0", type=float, default=1.,
                         action="store", help="Initial condition in GBM")
-    parser.add_argument("-qoi_seed", type=int, default=-1,
-                        action="store", help="Seed for random generator")
 
+def wrapFunMIMC(inds):
+    offset = 1
+    inp = [[foo[0]+offset, foo[1]+offset, foo[1]+offset] for foo in inds]
+    return HJM.infDimTest(inp)
 
-def main():
-    import argparse
-    parser = argparse.ArgumentParser(add_help=True)
-    addExtraArguments(parser)
-    mimc.MIMCRun.addOptionsToParser(parser)
-    mimcRun = mimc.MIMCRun(**vars(parser.parse_known_args()[0]))
-    if mimcRun.params.qoi_seed >= 0:
-        np.random.seed(mimcRun.params.qoi_seed)
-
-    fnItrDone = None
-    if mimcRun.params.db:
[…]

### 2016-05-04 tests/HJM/HJM.py
sqrt convergence test

@@ -753,6 +753,15 @@ def testcude(inds,tau1=1.0,tau2=1.0):
     return rv1
     #return [np.exp(-1*foo) for foo in rv1]
 
+def aTest(ks):
+    times = np.linspace(0,1,100)
+    dt = times[1]-times[0]
+    W = sp.randn(100)
+    W[0] *=0
+    W *= np.sqrt(dt)
+    W = np.cumsum(W)*dt
+    return [np.sum(np.sin(k*times)*W) for k in ks]
+
 def testFourierConvergence():
 
     L = 10.0
@@ -803,3 +812,9 @@ def testFourierConvergence():
     plt.ylabel('$E (g-\overline{g})^2$')
     plt.savefig('ntstrong.pdf')
 
+inputs = [10**foo for foo in range(4)]
+outputs = np.array([aTest(inputs) for m in range(1000000)])
+means = np.mean(np.abs(outputs),axis=0)
+plt.loglog(inputs,means)
+plt.grid(1)
+plt.show()

### 2016-05-04 tests/HJM/HJM.py
final N_f rates

@@ -722,7 +722,7 @@ def infDimTest(inds,verbose=False,maxLev=20):
     return infDimHjmModel(inds,F,G,U,Psi,f0,kappa,maxLev=maxLev,verbose=verbose)
 
 
-def testcude(inds,tau1=1.0,tau2=1.0):
+def testcude(inds,tau1=1.0,tau2=2.0):
     ntimes = max([ind[0] for ind in inds])
     ntimes = 2**ntimes+1
     times = np.linspace(0,1,ntimes)
@@ -733,10 +733,10 @@ def testcude(inds,tau1=1.0,tau2=1.0):
     Ws *= np.sqrt(dt)
     Ws[0,:] *=0
     Ws = np.cumsum(Ws,axis=0)
-    Wfinal = Ws[-1,:]
+    Wfinal = 1*Ws[-1,:]
     for col in range(len(Ws[0])):
         Wfinal[col] /= (col+1)**2
-        Wfinal[col] *= (np.sin(col*tau2)-np.sin(col*tau2))/(col+1)
+        Wfinal[col] *= (np.sin(col*tau2)-np.sin(col*tau1))/(col+1)
         Ws[:,col] /= (col+1)**2 #(c_k bit)
         #for row in range(len(Ws)):
         #    Ws[row,col] *= np.sin(col*times[row])
@@ -748,10 +748,10 @@ def testcude(inds,tau1=1.0,tau2=1.0):
     rv2 = []
     for ind in inds:
         cutoff = 2**(ind[1])
-        rv2.append(np.sum(Wfinal[:cutoff]))
+        rv2.append(np.exp(-1*np.sum(Wfinal[:cutoff])))
     #return [np.exp(-1*rv1[foo])*np.exp(-1*rv2[foo]) for foo in range(len(inds))]
-    return rv1
-    #return [np.exp(-1*foo) for foo in rv1]
+    #return rv2
+    return [np.exp(-1*foo) for foo in rv1]
 
 def aTest(ks):
     times = np.linspace(0,1,100)
@@ -812,9 +812,10 @@ def testFourierConvergence():
     plt.ylabel('$E (g-\overline{g})^2$')
     plt.savefig('ntstrong.pdf')
 
-inputs = [10**foo for foo in range(4)]
-outputs = np.array([aTest(inputs) for m in range(1000000)])
-means = np.mean(np.abs(outputs),axis=0)
-plt.loglog(inputs,means)
+plt.figure()
[…]

### 2016-05-04 tests/HJM/HJM.py
rates for N_t 1 and 2 ...

@@ -813,9 +813,9 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 plt.figure()
-plt.loglog([2**foo for foo in range(15)], np.mean(np.abs(np.diff([testcude([[2,foo] for foo in range(16)]) for m in range(100)],axis=1))**2,axis=0))
+plt.loglog([2**foo for foo in range(15)], np.mean(np.abs(np.diff([testcude([[foo,2] for foo in range(16)]) for m in range(100)],axis=1))**2,axis=0))
 plt.grid(1)
 
-plt.loglog([2**foo for foo in range(15)], np.mean(np.abs(np.diff([testcude([[2,foo] for foo in range(16)]) for m in range(100)],axis=1)),axis=0))
+plt.loglog([2**foo for foo in range(15)], np.mean(np.abs(np.diff([testcude([[foo,2] for foo in range(16)]) for m in range(100)],axis=1)),axis=0))
 plt.grid(1)
 plt.show()

### 2016-05-06 tests/HJM/manuscript.tex
plots made for manu

@@
 \end{align}
-with the diffusion term given as the infinite-dimensional expansion of
+with the diffusion term given as the infinite-dimensional extension of
 eq. (1.1) in \cite{bjork2013monte}:
 \begin{align}
-\beta \parent{t, \tau} d \overline W \parent t = \ssum{n=0}{\infty} c_n \parent{\sin \parent{\frac{n \pi \tau}{L}} + \cos \parent{\frac{n \pi \tau}{L}}} dW_n \parent t , 
+%\beta \parent{t, \tau} d \overline W \parent t = \ssum{k=0}{\infty} c_n \parent{\sin %\parent{\ffreq{k} \tau} + \cos \parent{\ffreq{k} \tau} dW_n \parent t , 
+\beta \parent{t, \tau} d \overline W \parent t = \ssum{k=0}{\infty}
+c_n \parent{ \sin \parent{\ffreq k \tau} + \cos \parent{\ffreq k \tau} }
 \end{align}
-with $W_n$ independent Brownian motions and $L>>1$. This gives the following Fourier decomposition for the covariance
-of interests:
+with $W_n$ independent Brownian motions and
+\begin{align*}
+\ffreq k \equiv \frac{k \pi }{L}
+\end{align*}
+with $L>>1$. This gives the following Fourier decomposition for the covariance
+of increments:
 \begin{align}
@@
 \\
-= & \ssum{n=0}{\infty} c_n^2 \parent{ \cos \parent{\frac{n \pi \parent{\tau_1-\tau_2 }}{L}}}
+= & \ssum{k=0}{\infty} c_k^2 \parent{ \cos \parent{\ffreq k \parent{\tau_1-\tau_2}}}
 \end{align}
@@
 &+
-\ssum{n=1}{\infty }
-\frac{c_n^2}{n}
-\sin \parent{\frac{n \pi \tau}{L}} \parent{\cos \parent{\frac{n \pi t}{L}} -  \sin \parent{\frac{n \pi t}{L}}}
+\ssum{k=1}{\infty }
+\frac{c_k^2}{k}
+\sin \parent{\ffreq k \tau } \parent{\cos \parent{\ffreq q t} -  \sin \parent{\ffreq k t}}
+\\
+&+
+\ssum{k=1}{\infty }
+\frac{c_k^2}{k}
+\cos \parent{ \ffreq k \tau } \parent{\cos \parent{\ffreq k t} -  \sin \parent{\ffreq k t}}
 \\
@@
 \frac{c_n^2}{n}
-\cos \parent{\frac{n \pi \tau}{L}} \parent{\cos \parent{\frac{n \pi t}{L}} -  \sin \parent{\frac{n \pi t}{L}}}
-\\
-&+
-\ssum{n=1}{\infty }
-\frac{c_n^2}{n}
-\cos \parent{\frac{2n \pi \tau}{L}}.
+\cos \parent{2 \ffreq k \tau }.
 \end{align*}
@@
 \\
-a_n \parent{t} =& \frac{c_n^2 L^2}{\pi^2 n^2} \parent{\cos \parent{\frac{n \pi t}{L}} + \sin \parent{\frac{n \pi t}{L}}-1 } + c_n W_n \parent{t}
-\\
-b_n \parent{t} =& a_n \parent{t} -  \frac{\mathbf{1}_{\frac{n}{2} \in \mathbb Z_+}  c_n^2 L t}{\pi n}.
+a_k \parent{t} = b_k \parent t =& \frac{c_k^2}{\ffreq k^2}\parent{\cos \parent{\ffreq k t} + \sin \parent{\ffreq k t}-1 } - \frac{\mathbf{1}_{\frac{k}{2} \in \mathbb Z_+}  c_j^2  t}{\ffreq k}  + c_k W_k \parent{t} .
 \end{align}
@@
 \begin{align*}
-&\int_0^{t_{max}} f \parent{s,s} ds
+&\int_0^{T} f \parent{s,s} ds
 \\
@@
 \ssum{k=1}{N_f}
-b_k \parent {t_n} \cos \parent{\frac{k \pi t_n}{L}} + a_k \parent {t_n} \sin \parent{\frac{k \pi t_n}{L}}
+b_k \parent {t_n} \cos \parent{ \ffreq k t_n} + a_k \parent {t_n} \sin \parent{\ffreq k t_n}
 }
@@
 with $t_n = \frac{nT}{N_t}$.
-This approximation gives the follwoing approximation error
-in $N_f$:
+This approximation gives the following approximation error
+in $N_f$.
+
+The natural next step is, of coures, estimating the error in approximating the
+integral by $F_{N_t,N_f}$. Firstly, there is the case of the frequency cutoff:
 \begin{align*}
-&\int_0^{T} f \parent{s,s} ds -\lim_{N_t \rightarrow \infty} F_{N_t,N_f}
+&\expp{ \parent{\int_0^T f \parent{s,s} ds - \lim_{N_t \rightarrow \infty} F_{N_t,N_f}}^2}
 \\
-=& \int_0^T \ssum{k=1}{N_f}
-\frac{c_k^2 L^2}{k^2 \pi^2}
-\parent{1+ \sin \parent{\frac{2 k \pi s}{L}}} ds
+=& \expp{\int_0^T \ssum{k=N_f+1}{\infty} \parent{b_k \parent{t_n} \cos \parent{\ffreq k t_n} + a_k \parent{t_n} \sin \parent{\ffreq k t_n}} ds}
 \\
-&+
- \int_0^T \ssum{k=1}{N_f} c_k W_k \parent s
-\parent{\sin \parent{\frac{ k \pi s}{L}} + \cos \parent{\frac{ k \pi s}{L}}}
-ds
+\approx &
+\expp{\parent{\int_0^T \ssum{k=N_f+1}{\infty} c_k W_k \parent{t} \parent{\sin \parent{\ffreq k t} + \cos \parent{\ffreq k t} }}^2}
 \\
-&-
-\int_0^T \ssum{k=1}{N_f}
-\frac{c_k^2 L}{\pi k} \cos \parent{\frac{ k \pi s}{L}},
-ds,
+= &
+\expp{\parent{ \ssum{k=N_f+1}{\infty} \int_0^T c_k W_k \parent{t} \parent{\sin \parent{\ffreq k t} + \cos \parent{\ffreq k t} } ds}^2}
+\\
+= & \bigo{\ssum{k=N_f+1}{\infty} \parent{\frac{c_k}{k}}^2}.
 \end{align*}
+The bound for the weak error can be obtained using Jensen's inequality. Setting as a test
+case the temporal correlation structure as
+$
+\rho \parent{\tau_1, \tau_2} \propto \expf{ - \kappa \absval{\tau_1-\tau2}}
+$
+it follows that $c^2_k =\frac{\kappa}{L} \frac{\parent{-1}^k\parent{1-\expf{-\absval{\kappa L}}}}{\kappa^2 + \parent{\ffreq k}^2}$.
+Using this as an example, we plot the empirical strong and weak convergence rates in images \ref{img:cutoff1weak} and \ref{img:cutoff1strong} respectively.
 
-Similarly, one may the underlying part of the 
+As for the time discretisation error, we are concerned with a rectangle quadrature
+error, giving us weak and strong error rate of $1$ and $2$ respectively.
+
+Similarly, one may estimate the underlying part of the 
 payoff functional
@@
 \ssum{k=1}{N_f}
-b_k \parent {T} \cos \parent{\frac{k \pi \tau}{L}} + a_k \parent {T} \sin \parent{\frac{k \pi \tau}{L}}
+b_k \parent T \cos \parent{\ffreq k \tau} + a_k \parent {T} \sin \parent{\ffreq k \tau}
 d \tau.
 \\
-= &+ \int_{\tau_1}^{\tau_2}
-f_0 \parent \tau d \tau + \frac{c_0 T}{2} \parent{\tau_2^2 - \tau_2 +\tau_1 - \tau_1^2} 
+= & \int_{\tau_1}^{\tau_2}
+f_0 \parent \tau  + \frac{c_0 T}{2} \parent{\tau_2^2 - \tau_2 +\tau_1 - \tau_1^2} d \tau
 \\
@@
 \ssum{k=1}{N_f} 
-\frac{b_k \parent T L}{k \pi} 
-\parent{\sin \parent{\frac{k \pi \tau_2}{L}}  - \sin \parent{\frac{k \pi \tau_1}{L}} }
+\frac{b_k \parent T}{\ffreq k } 
+\parent{\sin \parent{\ffreq k \tau_2}  - \sin \parent{\ffreq k \tau_1} }
 \\
@@
 \ssum{k=1}{N_f} 
-\frac{a_k \parent T L}{k \pi} 
-\parent{\cos \parent{\frac{k \pi \tau_2}{L}}  - \cos \parent{\frac{k \pi \tau_1}{L}} }
+\frac{a_k \parent T}{\ffreq k} 
+\parent{\cos \parent{\ffreq k \tau_2}  - \cos \parent{\ffreq k \tau_1} }
 \\

### 2016-05-06 tests/HJM/HJM.py

@@ -749,7 +749,6 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
     bfinal = 0*Wfinal
     zfinal = 0
 
-    '''
     for col in range(len(ans[0])):
         k = col+1
         for row in range(len(ans)):
@@ -758,8 +757,7 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
             bns[row,col] += ((c(k)/w(k))**2 )*(ck(k,t)+sk(k,t)-1.0)
             if not k%2:
                 ans[row,col] -= c(k)*c(k)*t/w(k)
-                bns[row,col] -=c(k)*c(k)*t/w(k)
-    '''
+                bns[row,col] -= c(k)*c(k)*t/w(k)
 
     # drift part done
     for col in range(len(ans[0])):
@@ -768,6 +766,7 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
             t = times[row]
             ans[row,col] += c(k)*Ws[row,col+1]
             bns[row,col] += c(k)*Ws[row,col+1]
+            #ans[row,col] += Ws[row,col+1]*c(k)
 
     # stochastic bit done
     for col in range(len(ans[0])):
@@ -775,6 +774,7 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
             for row in range(len(ans)):
                 t =times[row]
                 ans[row,col] *= sk(k,t)
+                #ans[row,col]*np.cos(k*t)
                 bns[row,col] *= ck(k,t)
 
     # t=tau set
@@ -800,6 +800,7 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
         jump = 2**(max([foo[0] for foo in inds]) - ind[0])
         cutoff = 2**(ind[1])
         summand = ans[0:-1:jump,:cutoff]+bns[0:-1:jump,:cutoff]
+        #ummand = ans[0:-1:jump,:cutoff]
         rv1.append((dt*jump)*np.sum(summand))
         #rv1[-1] += dt*jump*np.sum(zns[:-1])
         #rv1[-1] = np.exp(-1*rv1[-1])
@@ -870,5 +871,20 @@ def testFourierConvergence():
     plt.xlabel('$log_2 N_t$')
[…]

### 2016-05-06 tests/HJM/manuscript.tex
second plots made for manu

@@
 it follows that $c^2_k =\frac{\kappa}{L} \frac{\parent{-1}^k\parent{1-\expf{-\absval{\kappa L}}}}{\kappa^2 + \parent{\ffreq k}^2}$.
-Using this as an example, we plot the empirical strong and weak convergence rates in images \ref{img:cutoff1weak} and \ref{img:cutoff1strong} respectively.
+Using this as an example, we plot the empirical strong and weak convergence rates in figures \ref{img:cutoff1weak} and \ref{img:cutoff1strong} respectively.
 
 As for the time discretisation error, we are concerned with a rectangle quadrature
-error, giving us weak and strong error rate of $1$ and $2$ respectively.
+error, giving us weak and strong error rate of $1$ and $2$ respectively. Empirical
+test of these rates is presented in figures \ref{img:cutoff2weak} and \ref{img:cutoff2strong}
+respectively.
 

### 2016-05-06 tests/HJM/HJM.py

@@ -872,19 +872,19 @@ def testFourierConvergence():
     plt.ylabel('$E (g-\overline{g})^2$')
     plt.savefig('ntstrong.pdf')
 
-sample=[testcude([[7,foo] for foo in range(11)]) for m in range(50)]
+sample=[testcude([[foo,4] for foo in range(11)]) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(10)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0))
 plt.grid(1)
-plt.xlabel('$N_f$')
+plt.xlabel('$N_t$')
 plt.ylabel('Weak Error')
-plt.savefig('weakerr1.pdf')
+plt.savefig('weakerr2.pdf')
 
 plt.figure()
 plt.loglog([2**foo for foo in range(10)],np.mean(np.abs(np.diff(sample,axis=1)**2),axis=0))
 plt.grid(1)
-plt.xlabel('$N_f$')
+plt.xlabel('$N_t$')
 plt.ylabel('Strong Error')
-plt.savefig('strongerr1.pdf')
+plt.savefig('strongerr2.pdf')
 
 

### 2016-05-07 tests/HJM/manuscript.tex
complete pics 1

--- 2016-05-06 tests/HJM/manuscript.tex
@@
 $
-\rho \parent{\tau_1, \tau_2} \propto \expf{ - \kappa \absval{\tau_1-\tau2}}
+\rho \parent{\tau_1, \tau_2} \propto \expf{ - \kappa \absval{\tau_1-\tau_2}}
 $
 it follows that $c^2_k =\frac{\kappa}{L} \frac{\parent{-1}^k\parent{1-\expf{-\absval{\kappa L}}}}{\kappa^2 + \parent{\ffreq k}^2}$.
-Using this as an example, we plot the empirical strong and weak convergence rates in images \ref{img:cutoff1weak} and \ref{img:cutoff1strong} respectively.
+Using this as an example, we plot the empirical strong and weak convergence rates in figure \ref{fig:rateFig1}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{weakerr1.pdf}
+        %\caption{A gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{strongerr1.pdf}
+        %\caption{A tiger}
+    \end{subfigure}
+    \caption{\label{fig:rateFig1} Weak (left) and strong (right) error
+    $\nnorm{\int_0^T f \parent{s,s} ds - \lim_{N_t \rightarrow \infty} F_{N_t,N_f}}{}$. $N_t= 2^9+1$.}
+\end{figure}
 
 As for the time discretisation error, we are concerned with a rectangle quadrature
-error, giving us weak and strong error rate of $1$ and $2$ respectively.
+error, giving us weak and strong error rate of $1$ and $2$ respectively. Empirical
+test of these rates is presented in figure \ref{img:rateFig2}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{weakerr2.pdf}
+        %\caption{A gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{strongerr2.pdf}
+        %\caption{A tiger}
+    \end{subfigure}
+    \caption{\label{img:rateFig2} Weak (left) and strong (right) error
+    for the temporal discretisation
+    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$. $N_f= 2^4$.}
+\end{figure}
 
@@
 \end{align*}
-The corresponding empirical rates for the approximations of the quantity of interest
+This approximation has its own approximation error with respect to $N_f$:
+\begin{align*}
+&\expp{\parent{\int_{\tau_1}^{\tau_2} f \parent{T,\tau} d \tau - \Psi_{N_f}}^2}
+\\
+=& \expp{\parent{\ssum{k=N_f+1}{\infty} \frac{b_k \parent T}{\ffreq k } 
+\parent{\sin \parent{\ffreq k \tau_2}  - \sin \parent{\ffreq k \tau_1} } - \frac{a_k \parent T}{\ffreq k} 
+\parent{\cos \parent{\ffreq k \tau_2}  - \cos \parent{\ffreq k \tau_1} }  }^2}
+\\
+= & \bigo{\ssum{k = N_f+1}{\infty} \frac{c_k^2}{\ffreq k^2}}.
+\end{align*}
+Using the exponentially decaying covariance function, we expect order $3$ strong convergence
+and, using Jensen's inequality, we may bound the weak error to order $\frac{3}{2}$. Accompanying numerical rates are presetend in figure \ref{img:rateFig3}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{weakerr3.pdf}
+        %\caption{A gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{strongerr3.pdf}
+        %\caption{A tiger}
+    \end{subfigure}
+    \caption{\label{img:rateFig3} Weak (left) and strong (right) error
+    for the temporal discretisation
+    $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$.}
+\end{figure}
+
+Overall, we may approximate a future price of a zero-coupon bond as
 \begin{align*}
@@
 \approx
-\expp{
+\underbrace{\expp{
 F_{N_t,N_f}
-\Psi_{N_f} }
+\Psi_{N_f} }}_{\overline G_{N_t,N_f}}.
 \end{align*}
-are presented in fig. \ref{fig:empInfDimRates}
+Continuing the case of exponential covariance structure where $c_k \sim \bigo{k^{-2}}$,
+we expect, based on the above computations, to observe the following rate
+\begin{align*}
+\expp{ \parent{\mathcal G - \overline G_{N_t,N_f}}^2} = \bigo{N_f^{-3} N_t^{-2}}.
+\end{align*} 
+To complement the component-wise plots in figures \ref{fig:rateFig1}, \ref{img:rateFig2} and
+\ref{img:rateFig3}, we plot the overall approximation of $\mathcal G$ with $\overline G$
+in figure \ref{img:rateFig4}.
 

### 2016-05-07 tests/HJM/HJM.py

@@ -748,7 +748,6 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
     afinal = 0*Wfinal
     bfinal = 0*Wfinal
     zfinal = 0
-
     for col in range(len(ans[0])):
         k = col+1
         for row in range(len(ans)):
@@ -774,7 +773,6 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
             for row in range(len(ans)):
                 t =times[row]
                 ans[row,col] *= sk(k,t)
-                #ans[row,col]*np.cos(k*t)
                 bns[row,col] *= ck(k,t)
 
     # t=tau set
@@ -785,15 +783,15 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
     for col in range(len(ans[0])):
         k = col + 1
         T = times[-1]
-        afinal[col] *= c(k)*Wfinal[col]
-        bfinal[col] *= c(k)*Wfinal[col]
-        afinal[col] += ((c(k)/w(k))**2)*(ck(k,T)+sk(k,T)-1.0)
-        bfinal[col] += ((c(k)/w(k))**2)*(ck(k,T)+sk(k,T)-1.0)
-        if not k%2:
-            afinal[col] -=c(k)*c(k)*T/w(k)
-            bfinal[col] -=c(k)*c(k)*T/w(k)
-        afinal *= ski(k,tau1,tau2)
-        bfinal *= cki(k,tau1,tau2)
+        afinal[col] = c(k)*Wfinal[col]
+        bfinal[col] = c(k)*Wfinal[col]
+        #afinal[col] += ((c(k)/w(k))**2)*(ck(k,T)+sk(k,T)-1.0)
+        #bfinal[col] += ((c(k)/w(k))**2)*(ck(k,T)+sk(k,T)-1.0)
+        #if not k%2:
+        #    afinal[col] -=c(k)*c(k)*T/w(k)
+        #    bfinal[col] -=c(k)*c(k)*T/w(k)
+        afinal[col] *= ski(k,tau1,tau2)
+        bfinal[col] *= cki(k,tau1,tau2)
     # integrals over the underlyings computed
     rv1 = []
     for ind in inds:
@@ -807,11 +805,11 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
     rv2 = []
[…]

### 2016-05-07 tests/HJM/manuscript.tex
complete pics 1

@@
 To complement the component-wise plots in figures \ref{fig:rateFig1}, \ref{img:rateFig2} and
-\ref{img:rateFig3}, we plot the overall approximation of $\mathcal G$ with $\overline G$
+\ref{img:rateFig3}, we plot the overall approximation of $\mathcal G$ with $\overline G_{N_t,N_f}$
 in figure \ref{img:rateFig4}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{weakerr4.pdf}
+        %\caption{A gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{strongerr4.pdf}
+        %\caption{A tiger}
+    \end{subfigure}
+    \caption{\label{img:rateFig3} Above: Weak (left) and strong (right) error
+    for the temporal discretisation
+    $\nnorm{\overline G_{65,2N_f}-\overline G_{65,N_f}}{}$,
+    along with the $N_f^{-\frac{3}{2}}$ and $N_f^{3}$ reference lines (red).}
+\end{figure}
 

### 2016-05-07 tests/HJM/HJM.py

@@ -871,10 +871,10 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 
-rates =[1.5,3]
+rates =[1,2]
 
 N=11
-sample=[testcude([[8,foo] for foo in range(N)]) for m in range(50)]
+sample=[testcude([[foo,5] for foo in range(N)]) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')

### 2016-05-07 tests/HJM/HJM.py
complete pics 1 redo

@@ -871,17 +871,17 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 
-rates =[1,2]
+rates =[1.5,3]
 
-N=11
-sample=[testcude([[foo,5] for foo in range(N)]) for m in range(50)]
+N=10
+sample=[testcude([[8,foo] for foo in range(N)]) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')
 plt.grid(1)
 plt.xlabel('$N_f$')
 plt.ylabel('Weak Error')
-plt.savefig('weakerr4.pdf')
+plt.savefig('weakerr5.pdf')
 
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)**2),axis=0))
@@ -889,6 +889,6 @@ plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[1]) for foo in range
 plt.grid(1)
 plt.xlabel('$N_f$')
 plt.ylabel('Strong Error')
-plt.savefig('strongerr4.pdf')
+plt.savefig('strongerr5.pdf')
 
 

### 2016-05-07 tests/HJM/manuscript.tex
complete pics 2 redo

@@
     \caption{\label{img:rateFig3} Weak (left) and strong (right) error
-    for the temporal discretisation
+    for frequency cutoff
     $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$.}
@@
     \end{subfigure}
+    \\
+        \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{weakerr5.pdf}
+        %\caption{A gull}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{strongerr5.pdf}
+        %\caption{A tiger}
+    \end{subfigure}
     \caption{\label{img:rateFig3} Above: Weak (left) and strong (right) error
-    for the temporal discretisation
+    for the cutoff
     $\nnorm{\overline G_{65,2N_f}-\overline G_{65,N_f}}{}$,
-    along with the $N_f^{-\frac{3}{2}}$ and $N_f^{3}$ reference lines (red).}
+    along with the $N_f^{-\frac{3}{2}}$ and $N_f^{3}$ reference lines (red).
+    Below: Weak (left) and strong (right) error for the temporal discretisation
+    $\nnorm{\overline G_{2N_t-1,16} - \overline G_{N_t,16}}{}$ with the 
+    $N_t^{-1}$ and $N_t^{-2}$ reference lines for the weak and strong error,
+    respectively (red).}
 \end{figure}

### 2016-05-07 tests/HJM/HJM.py

@@ -871,24 +871,26 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 
-rates =[1.5,3]
+rates =[1,2]
+char = 't'
+onum = 3
 
 N=10
-sample=[testcude([[8,foo] for foo in range(N)]) for m in range(50)]
+sample=[testcude([[foo,4] for foo in range(N)]) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')
 plt.grid(1)
-plt.xlabel('$N_f$')
+plt.xlabel('$N_%s$'%(char,))
 plt.ylabel('Weak Error')
-plt.savefig('weakerr5.pdf')
+plt.savefig('weakerr%d.pdf'%(onum,))
 
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)**2),axis=0))
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[1]) for foo in range(N-1)],'r-')
 plt.grid(1)
-plt.xlabel('$N_f$')
+plt.xlabel('$N_%s$'%(char,))
 plt.ylabel('Strong Error')
-plt.savefig('strongerr5.pdf')
+plt.savefig('strongerr%d.pdf'%(onum,))
 
 

### 2016-05-07 tests/HJM/manuscript.tex
new nt plot

@@
     for the temporal discretisation
-    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$. $N_f= 2^4$.}
+    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$
+    along with $N_t^{-1}$ and $N_t^{-2}$ reference lines in red. $N_f= 2^4$.}
 \end{figure}

### 2016-05-07 tests/HJM/HJM.py

@@ -722,7 +722,7 @@ def infDimTest(inds,verbose=False,maxLev=20):
     return infDimHjmModel(inds,F,G,U,Psi,f0,kappa,maxLev=maxLev,verbose=verbose)
 
 
-def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
+def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False,ret=0):
     c = lambda k: np.sqrt(0.5*cCovExp(L,10.0,k)) # fourier term for covariance
     w = lambda k: np.pi*k/L # frequency k
     sk = lambda k,x: np.sin(w(k)*x) # sine basis function
@@ -809,6 +809,10 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False):
         rv2[-1] += zi(times[-1],tau1,tau2)
     #return [np.exp(-1*rv1[foo])*np.exp(-1*rv2[foo]) for foo in range(len(inds))]
     #return rv2
+    if ret==1:
+        return rv1
+    if ret==2:
+        return rv2
     return [np.exp(-1*rv1[foo])*np.exp(-1*rv2[foo]) for foo in range(len(rv2))]
 
 def aTest(ks):

### 2016-05-07 tests/HJM/manuscript.tex
new n_f underlying plot

--- 2016-05-07 tests/HJM/manuscript.tex
@@
 error, giving us weak and strong error rate of $1$ and $2$ respectively. Empirical
-test of these rates is presented in figure \ref{img:rateFig2}.
+test of these rates is presented in figure \ref{img:rateFig3}.
 
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{weakerr2.pdf}
+        \includegraphics[width=\textwidth]{weakerr3.pdf}
         %\caption{A gull}
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{strongerr2.pdf}
+        \includegraphics[width=\textwidth]{strongerr3.pdf}
         %\caption{A tiger}
     \end{subfigure}
-    \caption{\label{img:rateFig2} Weak (left) and strong (right) error
+    \caption{\label{img:rateFig3} Weak (left) and strong (right) error
     for the temporal discretisation
-    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$. $N_f= 2^4$.}
+    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$
+    along with $N_t^{-1}$ and $N_t^{-2}$ reference lines in red. $N_f= 2^4$.}
 \end{figure}
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{weakerr3.pdf}
+        \includegraphics[width=\textwidth]{weakerr2.pdf}
         %\caption{A gull}
@@
     \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{strongerr3.pdf}
+        \includegraphics[width=\textwidth]{strongerr2.pdf}
         %\caption{A tiger}

### 2016-05-07 tests/HJM/HJM.py

@@ -875,12 +875,13 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 
-rates =[1,2]
+rates =[1.5,3]
 char = 't'
-onum = 3
+onum = 2
+ret=2
 
 N=10
-sample=[testcude([[foo,4] for foo in range(N)]) for m in range(50)]
+sample=[testcude([[2,foo] for foo in range(N)],ret=ret) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')

### 2016-05-07 tests/HJM/HJM.py
redo cutoff plot for underlying, again

@@ -876,7 +876,7 @@ def testFourierConvergence():
 
 
 rates =[1.5,3]
-char = 't'
+char = 'f'
 onum = 2
 ret=2
 

### 2016-05-07 tests/HJM/manuscript.tex
fig 3 replot

@@
     \caption{\label{fig:rateFig1} Weak (left) and strong (right) error
-    $\nnorm{\int_0^T f \parent{s,s} ds - \lim_{N_t \rightarrow \infty} F_{N_t,N_f}}{}$. $N_t= 2^9+1$.}
+    $\nnorm{F_{513,2N_f} - F_{513,N_f}}{}$ along with $N_f^{-\frac{3}{2}}$ and
+    	$N_f^{3}$ reference lines in red.}
 \end{figure}
@@
     for frequency cutoff
-    $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$.}
+    $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$ together with the $N_f^{-\frac{3}{2}}$ and
+    $N_f^{-3}$ reference lines in red.}
 \end{figure}

### 2016-05-07 tests/HJM/HJM.py

@@ -877,11 +877,11 @@ def testFourierConvergence():
 
 rates =[1.5,3]
 char = 'f'
-onum = 2
-ret=2
+onum = 1
+ret=1
 
-N=10
-sample=[testcude([[2,foo] for foo in range(N)],ret=ret) for m in range(50)]
+N=9
+sample=[testcude([[9,foo] for foo in range(N)],ret=ret) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')

### 2016-05-08 tests/HJM/manuscript.tex
fig 1 rereplot

--- 2016-05-07 tests/HJM/manuscript.tex
@@
 \end{align*}
-The bound for the weak error can be obtained using Jensen's inequality. Setting as a test
+A bound for the bias error can be obtained using Jensen's inequality. Setting as a test
 case the temporal correlation structure as
@@
     \end{subfigure}
-    \caption{\label{fig:rateFig1} Weak (left) and strong (right) error
-    $\nnorm{\int_0^T f \parent{s,s} ds - \lim_{N_t \rightarrow \infty} F_{N_t,N_f}}{}$. $N_t= 2^9+1$.}
+    \caption{\label{fig:rateFig1} Bias (left) and variance (right) error
+    $\nnorm{F_{513,2N_f} - F_{513,N_f}}{}$ along with $N_f^{-\frac{3}{2}}$ and
+    	$N_f^{3}$ reference lines in red.}
 \end{figure}
@@
 Using the exponentially decaying covariance function, we expect order $3$ strong convergence
-and, using Jensen's inequality, we may bound the weak error to order $\frac{3}{2}$. Accompanying numerical rates are presetend in figure \ref{img:rateFig3}.
+and, using Jensen's inequality, we may bound the weak error to order $\frac{3}{2}$. Accompanying numerical rates are presentend in figure \ref{img:rateFig2}.
 
@@
     \end{subfigure}
-    \caption{\label{img:rateFig3} Weak (left) and strong (right) error
+    \caption{\label{img:rateFig2} Weak (left) and strong (right) error
     for frequency cutoff
-    $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$.}
+    $\nnorm{\Psi_{2N_f}-\Psi_{N_f}}{}$ together with the $N_f^{-\frac{3}{2}}$ and
+    $N_f^{-3}$ reference lines in red.}
 \end{figure}
@@
     \end{subfigure}
-    \caption{\label{img:rateFig3} Above: Weak (left) and strong (right) error
+    \caption{\label{img:rateFig4} Above: Weak (left) and strong (right) error
     for the cutoff
@@
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{nfstrong.pdf}
-        %\caption{A gull}
-        \label{fig:gull}
-    \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{nfweak.pdf}
-        %\caption{A tiger}
-        \label{fig:tiger}
-    \end{subfigure}
-    \\
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-    %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{ntstrong.pdf}
-        %\caption{A mouse}
-        \label{fig:mouse}
-    \end{subfigure}
-        \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[width=\textwidth]{ntweak.pdf}
-        %\caption{A mouse}
-        \label{fig:mouse}
-    \end{subfigure}
-    \caption{\label{fig:empInfDimRates}Empirical strong and weak convergence rate for
-    the $N_t$ and $N_f$.}
-\end{figure}
-
 \bibliographystyle{agsm}

### 2016-05-08 tests/HJM/HJM.py

@@ -887,7 +887,7 @@ plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1))
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')
 plt.grid(1)
 plt.xlabel('$N_%s$'%(char,))
-plt.ylabel('Weak Error')
+plt.ylabel('Bias')
 plt.savefig('weakerr%d.pdf'%(onum,))
 
 plt.figure()
@@ -895,7 +895,7 @@ plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)*
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[1]) for foo in range(N-1)],'r-')
 plt.grid(1)
 plt.xlabel('$N_%s$'%(char,))
-plt.ylabel('Strong Error')
+plt.ylabel('Variance')
 plt.savefig('strongerr%d.pdf'%(onum,))
 
 

### 2016-05-08 tests/HJM/manuscript.tex
fig 3 rereplot

@@
     \end{subfigure}
-    \caption{\label{img:rateFig2} Weak (left) and strong (right) error
+    \caption{\label{img:rateFig2} Bias (left) and variance (right) error
     for frequency cutoff

### 2016-05-08 tests/HJM/HJM.py

@@ -877,8 +877,8 @@ def testFourierConvergence():
 
 rates =[1.5,3]
 char = 'f'
-onum = 1
-ret=1
+onum = 2
+ret=2
 
 N=9
 sample=[testcude([[9,foo] for foo in range(N)],ret=ret) for m in range(50)]

### 2016-05-08 tests/HJM/manuscript.tex
plenty rereplot

@@
     \end{subfigure}
-    \caption{\label{img:rateFig3} Weak (left) and strong (right) error
+    \caption{\label{img:rateFig3} Bias (left) and variance (right) error
     for the temporal discretisation
-    $\nnorm{F_{2N_t-1,N_f}-F_{N_t,N_f}}{}$
-    along with $N_t^{-1}$ and $N_t^{-2}$ reference lines in red. $N_f= 2^4$.}
+    $\nnorm{F_{2N_t-1,16}-F_{N_t,16}}{}$
+    along with $N_t^{-1}$ and $N_t^{-2}$ reference lines in red.}
 \end{figure}
@@
 \begin{align*}
-\mathcal G \parent f =
+\mathcal G \parent f =&
 \expp{\expf{-\int_0^{t_{T}} f \parent{s,s} ds}\expf{-\int_{\tau_1}^{\tau_2} f \parent{T,\tau} d\tau}}
-\approx
-\underbrace{\expp{
-F_{N_t,N_f}
-\Psi_{N_f} }}_{\overline G_{N_t,N_f}}.
+\\
+\approx&
+\underbrace{\expp{\expf{-
+F_{N_t,N_f}-
+\Psi_{N_f} }}}_{\overline G_{N_t,N_f}}.
 \end{align*}
@@
     \end{subfigure}
-    \caption{\label{img:rateFig4} Above: Weak (left) and strong (right) error
+    \caption{\label{img:rateFig4} Above: Bias (left) and variance (right) error
     for the cutoff
@@
     along with the $N_f^{-\frac{3}{2}}$ and $N_f^{3}$ reference lines (red).
-    Below: Weak (left) and strong (right) error for the temporal discretisation
+    Below: Bias (left) and variance (right) error for the temporal discretisation
     $\nnorm{\overline G_{2N_t-1,16} - \overline G_{N_t,16}}{}$ with the 
-    $N_t^{-1}$ and $N_t^{-2}$ reference lines for the weak and strong error,
+    $N_t^{-1}$ and $N_t^{-2}$ reference lines for the bias and variance error,
     respectively (red).}

### 2016-05-08 tests/HJM/HJM.py

@@ -875,13 +875,13 @@ def testFourierConvergence():
     plt.savefig('ntstrong.pdf')
 
 
-rates =[1.5,3]
-char = 'f'
-onum = 2
-ret=2
+rates =[1,2]
+char = 't'
+onum = 3
+ret=1
 
 N=9
-sample=[testcude([[9,foo] for foo in range(N)],ret=ret) for m in range(50)]
+sample=[testcude([[foo,4] for foo in range(N)],ret=ret) for m in range(50)]
 plt.figure()
 plt.loglog([2**foo for foo in range(N-1)],np.mean(np.abs(np.diff(sample,axis=1)),axis=0),'b-')
 plt.loglog([2**foo for foo in range(N-1)],[2**(-1*foo*rates[0]) for foo in range(N-1)],'r-')

### 2016-05-08 tests/HJM/HJM.py
use overline G throughout

@@ -722,7 +722,7 @@ def infDimTest(inds,verbose=False,maxLev=20):
     return infDimHjmModel(inds,F,G,U,Psi,f0,kappa,maxLev=maxLev,verbose=verbose)
 
 
-def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False,ret=0):
+def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False,ret=0,plot=False):
     c = lambda k: np.sqrt(0.5*cCovExp(L,10.0,k)) # fourier term for covariance
     w = lambda k: np.pi*k/L # frequency k
     sk = lambda k,x: np.sin(w(k)*x) # sine basis function
@@ -748,6 +748,7 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False,ret=0):
     afinal = 0*Wfinal
     bfinal = 0*Wfinal
     zfinal = 0
+
     for col in range(len(ans[0])):
         k = col+1
         for row in range(len(ans)):
@@ -758,14 +759,36 @@ def testcude(inds,L=10.0,tau1=1.0,tau2=2.0,verbose=False,ret=0):
                 ans[row,col] -= c(k)*c(k)*t/w(k)
                 bns[row,col] -= c(k)*c(k)*t/w(k)
 
-    # drift part done
+    # drift part done                                                                                                                                                                                 
     for col in range(len(ans[0])):
         k = col+1
         for row in range(len(ans)):
             t = times[row]
             ans[row,col] += c(k)*Ws[row,col+1]
             bns[row,col] += c(k)*Ws[row,col+1]
-            #ans[row,col] += Ws[row,col+1]*c(k)
+            #ans[row,col] += Ws[row,col+1]*c(k) 
+
+    if plot:
+        fig=plt.figure()
+        ax=fig.gca(projection='3d')
+        ax.set_xlabel('$t$')
+        ax.set_ylabel('$\\tau$')
+        ax.set_zlabel('$f (\\tau, t)$')
+        shorts = []
+        for row in range(len(Ws)):
+            plotx = np.linspace(0,tau2,300)
+            t =times[row]
+            plotx = plotx[plotx>=t]
[…]

### 2016-05-08 tests/HJM/HJM.py
HJM notes

@@ -8,6 +8,7 @@ import scipy.linalg as spal
 import time
 import matplotlib as mpl
 from mpl_toolkits.mplot3d import Axes3D
+import itertools
 
 def hash(x):
     return '%f_%f_%f_%f'%(x[0],x[-1],x[1]-x[0],x[-1]-x[-2])
@@ -88,6 +89,123 @@ def twoFactorGaussianExample(inds,t_max=1.0,tau_max=3.0,b0=0.0759,b1=-0.0439,k=0
     identifierString += 'k: %f, a2: %f, K: %f, b1: %f'%(k,a2,K,b1)
 
     return multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=t_max,tau_max=tau_max,identifierString=identifierString,verbose=verbose)    
+
+def cCovExp(L,K,n):
+    return 2*K*((1-np.exp(-abs(K*L))*((-1)**n))/(K*K+n*n*np.pi*np.pi/L/L))/L
+
+def fourierExample(inds,cfun,L=50.0,tau1=1.0,tau2=2.0,t_max=1.0,verbose=False):
+
+    ts = [time.time(),]
+
+    w = lambda k: np.pi*k/L
+    plotN = 200
+
+    # largest values of the discretisation numbers                                                                           
+    N_t = max([foo[0] for foo in inds])
+    N_f = max([foo[1] for foo in inds])
+
+    N_t = 2**(N_t)+1
+    N_f = 2**(N_f)
+
+    times = np.linspace(0,t_max,N_t)
+    dt = times[1]-times[0]
+
+    Wt = sp.randn(N_t,N_f)*np.sqrt(dt)
+    Wt[0,:] *= 0
+    Wt = np.cumsum(Wt,axis=0)
+
+    if verbose:
+        print('W mean and variance: %f, %f'%(np.mean(Wt[-1,:]),np.var(Wt[-1,:])))
+    
+    consts = np.zeros(N_t)
+    ans = np.zeros((N_t,N_f-1))
+    bns = np.zeros((N_t,N_f-1))
+
+    for row in range(0,N_t):
+        consts[row] = Wt[row,0]
+        for col in range(0,N_f-1):
+            k = col+1
+            t = times[row]
+            #ans[row,col] = ((cfun(k)/w(k))**2)*(np.cos(t*w(k))+np.sin(t*w(k))-1.0)
+            ans[row,col] += cfun(k)*Wt[row,k]
+            bns[row,col] = ans[row,col]
[…]

### 2016-05-22 tests/gbm/echo_test_cmd.py
Simplifications to GBM
-run.py renamed to mimc_run.py to avoid confusion
-corresponding changes to ./echo_test_cmd.py
-update .gitignore

@@ -13,7 +13,7 @@ parser.add_argument("-db_host", type=str, action="store",
 
 args = parser.parse_known_args()[0]
 
-base = "run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
+base = "mimc_run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
 -qoi_sigma 0.1 -qoi_mu 1 -qoi_seed {seed} -mimc_moments 4 \
 -mimc_dim 1 -mimc_w 1 -mimc_s 1 -mimc_gamma 1 -mimc_beta 2 \
 -mimc_theta {theta} -mimc_bayesian {bayesian} "

### 2016-05-22 tests/gbm/mimc_run.py

@@ -1,3 +1,17 @@
+
+"""
+This file contains an example of a mimclib run.
+The example solves a geometric Brownian motion SDE
+with a given constant drift and volatility, with
+given initial value and from zero to a given final value.
+
+This file does NOT contain any parallel programming, but is 
+sequential.
+
+In case of doubt, run the commands produced in echo_test_cmd.py.
+
+"""
+
 from __future__ import absolute_import
 from __future__ import division
 from __future__ import print_function

### 2016-05-22 tests/gbm/mimc_run.py
Added missing files. Added some figures to slides

@@ -1,11 +1,10 @@
-
 """
 This file contains an example of a mimclib run.
 The example solves a geometric Brownian motion SDE
 with a given constant drift and volatility, with
 given initial value and from zero to a given final value.
 
-This file does NOT contain any parallel programming, but is 
+This file does NOT contain any parallel programming, but is
 sequential.
 
 In case of doubt, run the commands produced in echo_test_cmd.py.
@@ -59,8 +58,7 @@ try:
         return output
 
 except:
-    raise
-    warnings.warn("Using Python (very slow) version for wcumsum. Consider running make")
+    warnings.warn("Using Python's (very slow) version for wcumsum. Consider running make")
     # wcumsum is like cumsum, but weighted.
     wcumsum = python_cumsum
 

### 2016-05-23 tests/plot_data.py
Presented the slides

@@ -42,7 +42,6 @@ def main():
     db_args = dict()
     if args.db_name is not None:
         db_args["db"] = args.db_name
-    print(args.db_name)
     if args.db_user is not None:
         db_args["user"] = args.db_user
     if args.db_host is not None:

### 2016-05-26 mimclib/mimc.py
Updated files and slides for second presentation

@@ -436,7 +436,7 @@ Bias={:.12e}\nStatErr={:.12e}\
         L = L or len(self.all_data.lvls)-1
         if L <= 1:
             raise Exception("Must have at least 2 levels")
-        hl = self.fnHierarchy(np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
+        hl = self.fnHierarchy(lvls=np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
         return self.Q.W * hl[-1]**self.Q.w[0]
 
     def _estimateBayesianVl(self, L=None):
@@ -446,7 +446,7 @@ Bias={:.12e}\nStatErr={:.12e}\
         L = L or oL
         if L <= 1:
             raise Exception("Must have at least 2 levels")
-        hl = self.fnHierarchy(np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
+        hl = self.fnHierarchy(lvls=np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
         M = np.concatenate((self.all_data[1:].M, np.zeros(L-oL)))
         s1 = np.concatenate((self.all_data.psums_delta[1:, 0], np.zeros(L-oL)))
         m1 = np.concatenate((self.all_data[1:].calcDeltaEl(), np.zeros(L-oL)))
@@ -471,7 +471,7 @@ Bias={:.12e}\nStatErr={:.12e}\
         L = len(self.all_data.lvls)-1
         if L <= 1:
             raise Exception("Must have at least 2 levels")
-        hl = self.fnHierarchy(np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
+        hl = self.fnHierarchy(lvls=np.arange(0, L+1).reshape((-1, 1))).reshape(1, -1)[0]
         begin = np.maximum(1, L-self.params.bayes_fit_lvls)
         M = self.all_data[begin:].M
         # m1 = self.all_data[begin:].calcDeltaEl()
@@ -500,7 +500,7 @@ estimate optimal number of levels"
             bias_est = self._estimateBayesianBias(L)
             if bias_est >= TOL and L < LsRange[-1]:
                 continue
-            Wl = self.fnWorkModel(np.arange(0, L+1).reshape((-1, 1)))
+            Wl = self.fnWorkModel(lvls=np.arange(0, L+1).reshape((-1, 1)))
             M = self._calcTheoryM(TOL,
[…]

### 2016-05-26 mimclib/plot.py

@@ -349,7 +349,7 @@ def plotTotalWorkVsLvls(ax, runs_data, *args, **kwargs):
         if hasattr(curRun.run.params, "beta"):
             rate *= np.log(curRun.run.params.beta)
         ax.add_line(FunctionLine2D(lambda x, tol=TOL, r=rate: tol**-2 * np.exp(r*x),
-                                   data=total_work[:, :2],
+                                   data=np.array([lvls, total_work[:, 1]]).transpose(),
                                    linestyle='--', c='k'))
     return plotObj
 
@@ -870,6 +870,8 @@ def genPDFBooklet(runs_data, fileName=None, exact=None, **kwargs):
     try:
         line_data, _ = plotLvlsNumVsTOL(ax, runs_data)
         if has_beta and has_w_rate and has_gamma_rate:
+            import IPython
+            IPython.embed()
             if has_beta:
                 rate = 1./np.min(np.array(params.w) * np.log(params.beta))
             else:

### 2016-05-26 mimclib/test.py

@@ -17,7 +17,7 @@ def parse_known_args(parser, return_unknown=False):
     for a in unknowns:
         if a.startswith('-'):
             warnings.warn(ArgumentWarning("Argument {} was not used!".format(a)))
-            
+
     if return_unknown:
         return knowns, unknowns
     return knowns
@@ -74,9 +74,9 @@ def RunStandardTest(fnSampleLvl=None,
         db = mimcdb.MIMCDatabase(**db_args)
         run_id = db.createRun(mimc_run=mimcRun,
                               tag=mimcRun.params.db_tag)
-        fnItrDone = lambda *a: db.writeRunData(run_id, mimcRun, *a)
+        fnItrDone = lambda **kwargs: db.writeRunData(run_id, mimcRun, **kwargs)
 
-    fnSampleLvl = lambda inds, M, fn=fnSampleLvl, *a: fn(mimcRun, inds, M)
+    fnSampleLvl = lambda inds, M, fn=fnSampleLvl: fn(mimcRun, inds, M)
 
     mimcRun.setFunctions(fnSampleLvl=fnSampleLvl, fnItrDone=fnItrDone)
 

### 2016-05-26 tests/pde/echo_test_cmd.py

@@ -17,7 +17,7 @@ base = "mimc_run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  -mimc_dim {dim} -qoi_seed
 -qoi_x0 0.3,0.2,0.6 -qoi_sigma 0.16 -qoi_scale 1 -ksp_rtol 1e-25 \
 -mimc_theta 0.2 -qoi_scale 50 \
 -ksp_type gmres -mimc_w {w} -mimc_s {s} -mimc_gamma {gamma} -mimc_beta {beta} \
--mimc_bayes_fit_lvls 5 \
+-mimc_bayes_fit_lvls 3 -mimc_moments 4 \
 -mimc_bayesian {bayesian} ".format(bayesian="{bayesian}", TOL="{TOL}",
                                    dim=dim, seed="{seed}",
                                    gamma=" ".join([str(int(3/dim))]*dim),

### 2016-05-26 mimclib/mimc.py
Fixed plot of Vl_estimate

@@ -408,16 +408,17 @@ Bias={:.12e}\nStatErr={:.12e}\
                                 self.stat_error,
                                 self.totalErrorEst())
         V = self.Vl_estimate
+        Vl = self.data.calcDeltaVl()
         E = self.data.calcDeltaEl()
         T = self.data.calcTl()
 
-        output += ("{:<8}{:^20}{:^20}{:>8}{:>15}\n".format(
-            "Level", "E", "V", "M", "Time"))
+        output += ("{:<8}{:^20}{:^20}{:^20}{:>8}{:>15}\n".format(
+            "Level", "E", "V", "sampleV", "M", "Time"))
         for i in range(0, len(self.data.lvls)):
             assert(V[i]>=0)
             #,100 * np.sqrt(V[i]) / np.abs(E[i])
-            output += ("{:<8}{:>+20.12e}{:>20.12e}{:>8}{:>15.6e}\n".format(
-                str(self.data.lvls[i]), E[i], V[i], self.data.M[i], T[i]))
+            output += ("{:<8}{:>+20.12e}{:>20.12e}{:>20.12e}{:>8}{:>15.6e}\n".format(
+                str(self.data.lvls[i]), E[i], V[i], Vl[i], self.data.M[i], T[i]))
         return output
 
     ################## Bayesian specific functions

### 2016-05-26 mimclib/plot.py

@@ -431,13 +431,13 @@ def plotVarVsLvls(ax, runs_data, *args, **kwargs):
             plotObj.append(ax.plot(np.arange(0, len(Vl)), Vl, *fine_args, **fine_kwargs))
 
     if estimate_kwargs is not None:
-        fin = np.isfinite(Vl_estimate)
-        M = np.sum(fin, 1)
-        Vl_estimate[np.logical_not(fin)] = 0
-        avg = np.sum(Vl_estimate, 1)/M
-        err = np.sqrt(((np.sum(Vl_estimate**2, 1)/M) - avg**2)/M)
+        mdat = np.ma.masked_array(Vl_estimate, np.isnan(Vl_estimate))
+        med = np.ma.median(mdat, 1).filled(np.nan)
+        min_vl = np.min(mdat, 1).filled(np.nan)
+        max_vl = np.max(mdat, 1).filled(np.nan)
+        #err = np.sqrt(((np.sum(Vl_estimate**2, 1)/M) - avg**2)/M)
         plotObj.append(ax.errorbar(np.arange(0, len(Vl)),
-                                   avg, yerr=3*err,
+                                   med, yerr=[med-min_vl, max_vl-med],
                                    **estimate_kwargs))
     return plotObj[0][0].get_xydata(), plotObj
 
@@ -628,7 +628,7 @@ def plotThetaRefVsTOL(ax, runs_data, eta, chi, *args, **kwargs):
     return summary, [line]
 
 @public
-def plotErrorsQQ(ax, runs_data, *args, **kwargs): #(runs, tol, marker='o', color="b", fig=None, label=None):
+def plotErrorsQQ(ax, runs_data, *args, **kwargs):
     """Plots Normal vs Empirical CDF of @runs_data, as
     returned by MIMCDatabase.readRunData()
     ax is in instance of matplotlib.axes
@@ -870,8 +870,6 @@ def genPDFBooklet(runs_data, fileName=None, exact=None, **kwargs):
     try:
         line_data, _ = plotLvlsNumVsTOL(ax, runs_data)
         if has_beta and has_w_rate and has_gamma_rate:
-            import IPython
-            IPython.embed()
             if has_beta:
                 rate = 1./np.min(np.array(params.w) * np.log(params.beta))
             else:

### 2016-05-26 tests/gbm/echo_test_cmd.py

@@ -13,7 +13,8 @@ args, unknowns = parser.parse_known_args()
 base = "mimc_run.py -mimc_TOL {TOL} -mimc_max_TOL 0.5  \
 -qoi_sigma 0.1 -qoi_mu 1 -qoi_seed {seed} -mimc_moments 4 \
 -mimc_dim 1 -mimc_w 1 -mimc_s 1 -mimc_gamma 1 -mimc_beta 2 \
--mimc_theta {theta} -mimc_bayesian {bayesian} -mimc_bayes_fit_lvls 4 "
+-mimc_theta {theta} -mimc_bayesian {bayesian} -mimc_bayes_fit_lvls 4 \
+-mimc_bayes_k1 0.005 "
 
 
 base += " ".join(unknowns)

### 2016-05-26 mimclib/plot.py
Change Vl_estimate and fixed a typo in gbm example

@@ -431,10 +431,13 @@ def plotVarVsLvls(ax, runs_data, *args, **kwargs):
             plotObj.append(ax.plot(np.arange(0, len(Vl)), Vl, *fine_args, **fine_kwargs))
 
     if estimate_kwargs is not None:
-        mdat = np.ma.masked_array(Vl_estimate, np.isnan(Vl_estimate))
-        med = np.ma.median(mdat, 1).filled(np.nan)
-        min_vl = np.min(mdat, 1).filled(np.nan)
-        max_vl = np.max(mdat, 1).filled(np.nan)
+        # mdat = np.ma.masked_array(Vl_estimate, np.isnan(Vl_estimate))
+        # med = np.ma.median(mdat, 1).filled(np.nan)
+        # import IPython
+        # IPython.embed()
+        min_vl = np.nanpercentile(Vl_estimate, 5, axis=1)
+        med = np.nanpercentile(Vl_estimate, 50, axis=1)
+        max_vl = np.nanpercentile(Vl_estimate, 95, axis=1)
         #err = np.sqrt(((np.sum(Vl_estimate**2, 1)/M) - avg**2)/M)
         plotObj.append(ax.errorbar(np.arange(0, len(Vl)),
                                    med, yerr=[med-min_vl, max_vl-med],

### 2016-05-26 tests/gbm/mimc_run.py

@@ -28,16 +28,6 @@ def python_wcumsum(x, w):
             output[m, i] = w[m, i]*output[m, i-1] + x[i]
     return output
 
-def addExtraArguments(parser):
-    parser.add_argument("-qoi_sigma", type=float, default=1.,
-                        action="store", help="Volatility in GBM")
-    parser.add_argument("-qoi_mu", type=float, default=1.,
-                        action="store", help="Drift in GBM")
-    parser.add_argument("-qoi_T", type=float, default=1.,
-                        action="store", help="Final time in GBM")
-    parser.add_argument("-qoi_S0", type=float, default=1.,
-                        action="store", help="Initial condition in GBM")
-
 try:
     # Try to import the DLL version of wcumsum,
     # This makes solving the SDE much faster
@@ -60,7 +50,18 @@ try:
 except:
     warnings.warn("Using Python's (very slow) version for wcumsum. Consider running make")
     # wcumsum is like cumsum, but weighted.
-    wcumsum = python_cumsum
+    wcumsum = python_wcumsum
+
+
+def addExtraArguments(parser):
+    parser.add_argument("-qoi_sigma", type=float, default=1.,
+                        action="store", help="Volatility in GBM")
+    parser.add_argument("-qoi_mu", type=float, default=1.,
+                        action="store", help="Drift in GBM")
+    parser.add_argument("-qoi_T", type=float, default=1.,
+                        action="store", help="Final time in GBM")
+    parser.add_argument("-qoi_S0", type=float, default=1.,
+                        action="store", help="Initial condition in GBM")
 
 def mySampleQoI(run, inds, M):
     meshes = (run.params.qoi_T/run.fnHierarchy(inds)).reshape(-1).astype(np.int)

### 2016-05-26 mimclib/mimc.py
See modified files

@@ -314,7 +314,7 @@ are the same as the argument ones")
                   help="Use Bayesian fitting to estimate bias, variance and optimize number \
 of levels in every iteration. This is based on CMLMC.")
         add_store('dim', type=int, help="Number of dimensions used in MIMC")
-        add_store('moments', type=int, default=2, help="Number of moments to compute")
+        add_store('moments', type=int, default=4, help="Number of moments to compute")
         add_store('reuse_samples', type='bool', default=True,
                   help="Reuse samples between iterations")
         add_store('abs_bnd', type='bool', default=False,

### 2016-05-26 mimclib/plot.py

@@ -1,3 +1,7 @@
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+
 import numpy as np
 import matplotlib.pylab as plt
 from . import mimc
@@ -198,21 +202,31 @@ def plotErrorsVsTOL(ax, runs_data, *args, **kwargs):
     run_data[i].run.data is an instance of mimc.MIMCData
     """
 
+    num_kwargs = kwargs.pop('num_kwargs', None)
     exact = kwargs.pop('exact', None)
+    relative_error = kwargs.pop('relative', True)
     if exact is None:
         # Calculate mean based on data
-        minTOL = np.min([r.finalTOL for r in runs_data])
+        # final_run_data = [d for d in run_data if d.iteration_index+1
+        #                   == d.total_iterations]
+        # minTOL = np.min([r.TOL for r in runs_data])
+        # minfinalTOL = np.min([r.TOL for r in final_runs_data])
+        # if minfinalTOL > minTOL:
+        #     warnings.warn("The runs with the least tolerance are not done")
+
+        # exact = np.mean([r.run.data.calcEg() for r in final_run_data if
+        #                  r.TOL == minfinalTOL])
+        minfinalTOL = np.min([r.finalTOL for r in runs_data])
         exact = np.mean([r.run.data.calcEg() for r in runs_data if
-                         r.finalTOL == minTOL])
+                         r.finalTOL == minfinalTOL])
+        print("Computed exact value is:", exact)
 
+    modifier = (1./exact) if relative_error else 1.
     xy = np.array([[r.finalTOL, np.abs(exact - r.run.data.calcEg()),
-                    r.run.totalErrorEst()]
-                   for r in runs_data])
-    relative_error = kwargs.pop('relative', True)
-    if relative_error:
-        xy[:, 1:] = xy[:, 1:]/exact
-    TOLs, error_est = __get_stats(xy, staton=2)
+                    r.run.totalErrorEst()] for r in runs_data])
+    xy[:, 1:] = xy[:, 1:] * modifier
 
+    TOLs, error_est = __get_stats(xy, staton=2)
     plotObj = []
 
     ax.set_xlabel('TOL')
@@ -235,10 +249,23 @@ def plotErrorsVsTOL(ax, runs_data, *args, **kwargs):
                                          error_est[:, 2]-error_est[:, 1]],
                                    **ErrEst_kwargs))
     if TOLRef_kwargs is not None:
[…]

### 2016-05-26 mimclib/plot.py
Replaced min/max by percentiles in plots

@@ -175,7 +175,9 @@ def __get_stats(data, groupby=0, staton=1):
     y = []
     for k, itr in itertools.groupby(data, key=lambda xx: xx[groupby]):
         all_y = [d[staton] for d in itr]
-        y.append([np.min(all_y), np.mean(all_y), np.max(all_y)])
+        y.append([np.nanpercentile(all_y, 5),
+                  np.nanpercentile(all_y, 50),
+                  np.nanpercentile(all_y, 95)])
         x.append(k)
     return np.array(x), np.array(y)
 

### 2016-10-09 mimc_install.sh
Remove outdated installation script

@@ -1,51 +1,0 @@
-#!/bin/bash
-
-# This script is used to install mimclib and its main dependencies
-# on a Debian based machine.
-
-# The prerequisite of this is a Debian system and root access using
-# sudo.
-
-# This script  will upgrade your existing setup prior to adding extra
-# packages. If you are not happy with this, you might want to modify this
-# script. Comments and versions of this script for other platforms
-# are welcome in the form of github issues, pull requests or
-# emails to juho.happola@iki.fi
-
-# The script might not run without supervision. Unless you have mysql server
-# installed already, it will prompt for a mysql root password.
-# Also, depending on fast the required files are being downloaded
-# you might need to enter root password a few times. The runtime
-# of this script might be around an hour, depending on your
-# network speed and which packages need to be upgraded and installed.
-
-# This has been tried out with Debian Jessy with the May 22th 2016 (AMD64)
-# and the following commit of the library:
-# d94151770d87f4f685efc3719ba3b08249d4e5ae
-
-echo -e "First, update existing packages..."
-
-sudo apt-get update
-sudo apt-get upgrade
-
-echo -e "Existing packages upgraded."
-echo -e "Installing packages..."
-
-sudo apt-get install git python-pip mysql-server mysql-client \
-     libmysqlclient-dev build-essential ipython libpng-dev libfreetype6-dev \
-     libxft-dev libpython-dev liblapack-dev libblas-dev gfortran parallel
-
-echo -e "Done installing aptitude packages."
-
-echo -e "Making and installing"
-pip install --upgrade pip
-make pip
-
-echo -e "Adding a user to mySQL:"
-echo -e "CREATE USER '$USER'@'%';\nGRANT ALL PRIVILEGES ON mimc.* TO '$USER'@'%' WITH GRANT OPTION;" | mysql -u root -p
-
-echo -e "Creating the mimc database to mySQL.\n You will need to enter mySQL root password."
-python -c 'from mimclib.db import MIMCDatabase; print MIMCDatabase().DBCreationScript();' | mysql -u $USER
-
-cd tests/gbm
-make

### 2016-10-09 mimclib/db.py
Removed ipdb imports and updated slides

@@ -5,7 +5,6 @@ from __future__ import print_function
 import numpy as np
 import cPickle
 from . import setutil
-from . import ipdb
 
 import hashlib
 __all__ = []

### 2016-10-09 mimclib/mimc.py

@@ -9,7 +9,6 @@ import numpy as np
 import itertools
 import warnings
 from . import setutil
-from . import ipdb
 from scipy.stats import norm
 
 __all__ = []

### 2016-10-09 mimclib/plot.py

@@ -5,7 +5,6 @@ from __future__ import print_function
 import numpy as np
 import matplotlib.pylab as plt
 from . import mimc
-from . import ipdb
 from matplotlib.ticker import MaxNLocator
 
 __all__ = []
@@ -361,7 +360,6 @@ def computeIterationStats(runs, work_bins, xi, filteritr, fnNorm=None,
         xy_binned[i, 5] = np.max(d[:, 7])  # max active dim
 
     xy_binned = xy_binned[xy_binned[:,0].argsort(), :]
-    #ipdb.embed()
     return xy_binned
 
 @public

### 2016-10-09 tests/gbm/mimc_run.py

@@ -135,10 +135,7 @@ def initRun(run):
 
 if __name__ == "__main__":
     import mimclib.test
-    import mimclib.ipdb
     import sys
-    mimclib.ipdb.set_excepthook()
-
     mimclib.test.RunStandardTest(fnSampleLvl=mySampleQoI,
                                  fnAddExtraArgs=addExtraArguments,
                                  fnInit=initRun)

### 2016-10-09 tests/matern/mimc_run.py

@@ -6,7 +6,6 @@ import warnings
 import os.path
 import numpy as np
 import mimclib.test
-import mimclib.ipdb as ipdb
 import mimclib.misc as misc
 from matern import SField_Matern
 from mimclib import setutil
@@ -136,8 +135,6 @@ class MyRun:
 
 
 if __name__ == "__main__":
-    ipdb.set_excepthook()
-
     SField_Matern.Init()
     run = MyRun()
     mimclib.test.RunStandardTest(fnSampleLvl=run.mySampleQoI,

### 2016-10-09 tests/particle/mimc_run.py

@@ -33,8 +33,6 @@ class ParticleField(object):
         return samples, time.time()-tStart
 
 if __name__ == "__main__":
-    from mimclib import ipdb
-    ipdb.set_excepthook()
     import mimclib.test
     field = ParticleField()
     mimclib.test.RunStandardTest(fnInit=field.Init,

### 2016-10-09 tests/pde/mimc_run.py

@@ -6,8 +6,6 @@ import numpy as np
 import warnings
 import sys
 import mimclib.test
-# import IPython.ultraTB
-# sys.excepthook = IPython.ultraTB.FormattedTB(mode='Verbose', color_scheme='Linux', call_pdb=1)
 
 warnings.filterwarnings("error")
 warnings.filterwarnings("ignore", category=mimclib.test.ArgumentWarning)
@@ -23,9 +21,6 @@ def SamplePDE(sf, run, inds, M):
     return solves, time.time() - timeStart
 
 if __name__ == "__main__":
-    from mimclib import ipdb
-    ipdb.set_excepthook()
-
     from pdelib.SField import SField
     SField.Init()
     with SField() as sf:

### 2016-10-09 tests/plot_prog.py

@@ -1,9 +1,4 @@
 #!/usr/bin/python
 if __name__ == "__main__":
     from mimclib.plot import run_program
-    from mimclib import ipdb
-    ipdb.set_excepthook()
     run_program()
-
-    # from mimclib import ipdb
-    # ipdb.embed()

### 2016-10-09 setup.py
Updated version number

@@ -18,7 +18,7 @@ setup(
     # 1st is increased if it is not compatible with previous version.
     # 2nd is increased if new features are added.
     # 3rd is increased for small fixes.
-    version="1.0.0.dev0",
+    version="1.0.1.dev0",
     author="Abdul-Lateef Haji-Ali",
     author_email="abdullateef.hajiali@kaust.edu.sa",
     description="A library implementing the MIMC and CMLMC methods.",

### 2016-10-09 mimclib/__init__.py
Fixed a bug in MLMC, added a version property for module'

@@ -3,3 +3,10 @@ from __future__ import absolute_import
 from .mimc import MIMCRun, MIMCItrData
 from . import plot
 from .db import MIMCDatabase
+
+
+try:
+    import pkg_resources  # part of setuptools
+    __version__ = pkg_resources.require("mimclib")[0].version
+except:
+    __version__ = "Not installed"

### 2016-10-09 mimclib/mimc.py

@@ -556,10 +556,10 @@ Not needed if fnHierarchy is provided.")
     def __str__(self):
         output = "Eg={}\n\
 Bias={:.12e}\nStatErr={:.12e}\
-\nTotalErrEst={:.12e}\n".format(str(self.last_itr.calcEg()),
-                                self.bias,
-                                self.stat_error,
-                                self.totalErrorEst())
+\nTotalErrEst={:.12e} | {:.12e}\n".format(str(self.last_itr.calcEg()),
+                                          self.bias, self.stat_error,
+                                          self.totalErrorEst(),
+                                          self.last_itr.TOL)
         V = self.Vl_estimate
         Vl = self.fn.Norm(self.last_itr.calcDeltaVl())
         E = self.fn.Norm(self.last_itr.calcDeltaEl())
@@ -748,7 +748,8 @@ estimate optimal number of levels"
         lvls = self.last_itr.get_lvls()
         lvls_count = self.last_itr.lvls_count
         t = np.zeros(lvls_count)
-        active = totalM >= self.last_itr.M
+        active = totalM > self.last_itr.M
+        totalM[totalM <= self.last_itr.M] = 0    # No need to do any samples
         totalM[active] -= self.last_itr.M[active]
         if np.sum(totalM) == 0:
             return False
@@ -868,11 +869,12 @@ estimate optimal number of levels"
                 if samples_added:
                     if self.fn.ItrDone is not None:
                         self.fn.ItrDone()
-                    if self.params.bayesian or self.totalErrorEst() < TOL:
-                        break
                 else:
                     # remove last iteration since it is empty
+                    assert(self.params.bayesian or self.totalErrorEst() < TOL)
                     self.iters.pop()
+                if self.params.bayesian or self.totalErrorEst() < TOL:
+                    break
 
             print_info("MIMC iteration for TOL={} took {} seconds".format(TOL, timer.toc()))
             print_info("################################################")

### 2016-10-09 setup.py

@@ -18,7 +18,7 @@ setup(
     # 1st is increased if it is not compatible with previous version.
     # 2nd is increased if new features are added.
     # 3rd is increased for small fixes.
-    version="1.0.1.dev0",
+    version="1.0.2.dev0",
     author="Abdul-Lateef Haji-Ali",
     author_email="abdullateef.hajiali@kaust.edu.sa",
     description="A library implementing the MIMC and CMLMC methods.",

### 2016-10-09 tests/particle/echo_test_cmd.py
Fixed tests

@@ -36,7 +36,7 @@ if __name__ == "__main__":
         print(cmd_single.format(seed=0, bayesian=False, TOL=0.001))
     else:
         cmd_multi = "python " + base + " -mimc_verbose 0 -db True -db_tag {tag} "
-        TOL = 0.05*np.sqrt(2.)**-21
+        TOL = 1e-4
         for i in range(0, args.tries):
             print cmd_multi.format(bayesian=False,
                                    tag="particle_{}".format(dim), TOL=TOL,

### 2017-02-07 tests/matern/miproj_run.py
Initial number of dimensions is too large

@@ -67,7 +67,7 @@ class MyRun:
             error = run.fn.Norm(run.last_itr.calcDeltaEl())
             work = run.last_itr.Wl_estimate
             prof = setutil.calc_log_prof_from_EW(error, work)
-            max_added = 100
+            max_added = 30
         else:
             prof = self.profit_calc
         max_dim = 5 + (0 if len(lvls) == 0 else np.max(lvls.get_dim()))

### 2017-02-07 tests/matern/miproj_run.sh

@@ -6,7 +6,7 @@
 COMMON="python miproj_run.py -mimc_TOL 1e-10 -qoi_seed 0 -qoi_problem 0 -qoi_sigma 0.2  \
        -qoi_x0 0.3 0.4 0.6 -ksp_rtol 1e-25 -ksp_type gmres -qoi_a0 0 -qoi_f0 1 \
        -qoi_scale 10 -qoi_df_sig 0.5 -mimc_beta 2 -mimc_gamma 1 -mimc_h0inv 3  -mimc_verbose 1 \
-       -db True -db_engine sqlite -db_name data.sql"
+       -db True -db_name mimc"
 
 function run_cmd {
     # d max_lvl nu set extra_args
@@ -33,33 +33,33 @@ function plot_cmd {
     echo ./plot_cond.py -db_name miproj_matern -db_host $HOST -db_tag $1-arcsine
 }
 
-#################### RUNNING
-run_cmd fixproj 1 10 3.5 -mimc_min_dim 0 -miproj_min_dim 20
-run_cmd fixproj 1 10 2.5 -mimc_min_dim 0 -miproj_min_dim 30
-run_cmd fixproj 1 10 1.5 -mimc_min_dim 0 -miproj_min_dim 50
-run_cmd fixproj 1 10 1.0 -mimc_min_dim 0 -miproj_min_dim 50
+if [ $1 == "plot" ]; then
+    plot_cmd fixproj-1-1.0
+    plot_cmd fixproj-1-1.5
+    plot_cmd fixproj-1-2.5
+    plot_cmd fixproj-1-3.5
 
-run_cmd miproj 1 10 3.5 -miproj_min_dim 20
-run_cmd miproj 1 10 2.5 -miproj_min_dim 30
-run_cmd miproj 1 10 1.5 -miproj_min_dim 50
-run_cmd miproj 1 10 1.0 -miproj_min_dim 50
+    plot_cmd miproj-1-1.0
+    plot_cmd miproj-1-1.5
+    plot_cmd miproj-1-2.5
+    plot_cmd miproj-1-3.5
 
-run_cmd fixproj 3 7 3.0 -mimc_min_dim 0 -miproj_min_dim 50
-run_cmd fixproj 3 7 4.5 -mimc_min_dim 0 -miproj_min_dim 50
+    # plot_cmd miproj-3-4.5
+    # plot_cmd miproj-3-3.0
+else
[…]

### 2017-02-07 mimclib/ipdb.py
Changed all time.time instances to time.clock

@@ -377,7 +377,7 @@ class MyTB(ultratb.ListTB):
         if long_header:
             # Header with the exception type, python version, and date
             pyver = 'Python ' + sys.version.split()[0] + ': ' + sys.executable
-            date = time.ctime(time.time())
+            date = time.ctime(time.clock())
 
             if etype is not None:
                 head = '%s%s%s\n%s%s%s\n%s' % (Colors.topline, '-'*75, ColorsNormal,

### 2017-02-07 mimclib/mimc.py

@@ -27,14 +27,14 @@ class Timer():
         self.tic()
 
     def tic(self):
-        self._tics.append(time.time())
+        self._tics.append(time.clock())
 
     def toc(self, pop=True):
         assert(len(self._tics) > 0)
         if pop:
-            return time.time()-self._tics.pop()
+            return time.clock()-self._tics.pop()
         else:
-            return time.time()-self._tics[-1]
+            return time.clock()-self._tics[-1]
 
     def ptoc(self, msg='Time since last tic: {:.4f} sec.', pop=False):
         assert(len(self._tics) > 0)

### 2017-02-07 mimclib/miproj.py

@@ -209,7 +209,7 @@ class MIWProjSampler(object):
         total_time = np.empty(len(lvls))
         total_work = np.empty(len(lvls))
         for alpha, ind in self.alpha_dict.iteritems():
-            tStart = time.time()
+            tStart = time.clock()
             sel_lvls = self.alpha_ind == ind
             work_per_sample = self.fnWorkModel(setutil.VarSizeList([alpha]))
             beta_indset = lvls.sublist(sel_lvls, d_start=self.d, min_dim=0)
@@ -234,13 +234,10 @@ class MIWProjSampler(object):
                     X, W = self.fnSamplePoints(c_samples - len(self.prev_samples[ind]), basis)
                     self.prev_samples[ind].add_points(fnSample, inds, X, W)
                 X, W, Y = self.prev_samples[ind].XWY
-                sampling_time = self.prev_samples[ind].total_time
             else:
-                sampling_time = time.time()
                 assert(c_samples > 0)
                 X, W = self.fnSamplePoints(c_samples, basis)
                 Y = [fnSample(inds[i], X) for i in xrange(0, len(inds))]
-                sampling_time = time.time() - sampling_time
 
             basis_values = TensorExpansion.evaluate_basis(self.fnBasis, basis, X)
             for i in xrange(0, len(inds)):
@@ -267,10 +264,10 @@ class MIWProjSampler(object):
                     psums_delta[sel_lvls, 0] += projections*mods[i]
 
             if self.reuse_samples:
-                self.prev_samples[ind].total_time += time.time() - tStart
+                self.prev_samples[ind].total_time += time.clock() - tStart
                 sampling_time = self.prev_samples[ind].total_time
             else:
-                sampling_time = time.time() - tStart
+                sampling_time = time.clock() - tStart
 
             total_time[sel_lvls] = sampling_time * samples_per_beta / np.sum(samples_per_beta)
             total_work[sel_lvls] = work_per_sample * samples_per_beta

### 2017-02-07 mimclib/misc.py

@@ -125,7 +125,7 @@ class MISCSampler(object):
         # TODO: Need to generalize to allow for array or general objects
         from . import mimc
         import time
-        t = time.time()
+        t = time.clock()
         samples = np.empty((M, len(inds)))
         for i, dind in enumerate(inds):
             alpha = dind[:self.d]
@@ -134,7 +134,7 @@ class MISCSampler(object):
             knots = self.collapsePoints(knots)
             values = self._solveAtPoints(fnSample, alpha, knots)
             samples[0, i] = np.sum(weights * values)
-        work = time.time()-t
+        work = time.clock()-t
         return samples, work
 
 @public

### 2017-02-07 mimclib/test.py

@@ -78,7 +78,7 @@ def RunStandardTest(fnSampleLvl=None,
         mimcRun.setFunctions(fnSampleAll=fnSampleAll)
 
     import time
-    tStart = time.time()
+    tStart = time.clock()
     if not hasattr(mimcRun.params, 'qoi_seed'):
         mimcRun.params.qoi_seed = np.random.randint(2**32-1)
 
@@ -118,11 +118,11 @@ def RunStandardTest(fnSampleLvl=None,
         mimcRun.doRun()
     except:
         if mimcRun.params.db:
-            db.markRunFailed(run_id, totalTime=time.time()-tStart)
+            db.markRunFailed(run_id, totalTime=time.clock()-tStart)
         raise   # If you don't want to raise, make sure the following code is not executed
 
     if mimcRun.params.db:
-        db.markRunSuccessful(run_id, totalTime=time.time()-tStart)
+        db.markRunSuccessful(run_id, totalTime=time.clock()-tStart)
     return mimcRun
 
 def run_errors_est_program(fnExactErr=None):

### 2017-02-07 tests/HJM/HJM.py

@@ -95,7 +95,7 @@ def cCovExp(L,K,n):
 
 def fourierExample(inds,cfun,L=50.0,tau1=1.0,tau2=2.0,t_max=1.0,verbose=False):
 
-    ts = [time.time(),]
+    ts = [time.clock(),]
 
     w = lambda k: np.pi*k/L
     plotN = 200
@@ -213,7 +213,7 @@ def infDimHjmModel(inds,F,G,U,Psi,f0,kappa,t_max=1.0,tau_max=2.0,identifierStrin
     Solve an infinite-dimensional HJM model, crude example
     '''
 
-    ts = [time.time(),]
+    ts = [time.clock(),]
 
     if verbose:
         print('Evaluating the Two Factor Gaussian example.')
@@ -272,7 +272,7 @@ def infDimHjmModel(inds,F,G,U,Psi,f0,kappa,t_max=1.0,tau_max=2.0,identifierStrin
     for jj in range(1,len(Ws)):
         Ws[jj,:] = Ws[jj-1,:] + np.sqrt(dt)*np.dot(covMat,sp.randn(len(taus)))
 
-    ts.append(time.time())
+    ts.append(time.clock())
 
     if verbose:
         print('Mesh generations and initialisations done in %d seconds'%(ts[-1]-ts[-2]))
@@ -349,7 +349,7 @@ def infDimHjmModel(inds,F,G,U,Psi,f0,kappa,t_max=1.0,tau_max=2.0,identifierStrin
             print('The absurd additive term equals %f'%(weirdTerm,))
         rv[-1] *= underlying
         rv[-1] += weirdTerm
-        ts.append(time.time())
+        ts.append(time.clock())
         if verbose:
             print('The quantity of interest is %f'%(rv[-1]))
             print('Time spent on the level: %d seconds'%(ts[-1]-ts[-2]))
@@ -365,7 +365,7 @@ def multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=1.0,tau_max=2.0,identi
     Template to solve HJM type problems
     '''
 
-    ts = [time.time(),]
+    ts = [time.clock(),]
 
     if verbose:
         print('Evaluating the Two Factor Gaussian example.')
@@ -414,7 +414,7 @@ def multiLevelHjmModel(inds,F,G,U,Psi,drift,vols,f0,t_max=1.0,tau_max=2.0,identi
 
     rv = []
 
-    ts.append(time.time())
+    ts.append(time.clock())
 
     if verbose:
[…]

### 2017-02-07 tests/matern/miproj_esterr.py

@@ -22,7 +22,7 @@ def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=2000):
     if N < 0:
         N = np.max([itr.parent.last_itr.lvls_max_dim()-1 for itr in itrs])
 
-    tStart = time.time()
+    tStart = time.clock()
     print("MaxL:", max_L, ", N:", N)
     # Evaluate L^2 norm using Monte Carlo
     val = [itr.calcEg() for itr in itrs]
@@ -43,7 +43,7 @@ def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=2000):
         err = 3*np.sqrt((s2/M - (s1/M)**2)/M)    # Approximate error of error estimate
         max_rel_err = np.max(np.sqrt(err / (s1/M)))
         print("Estimate with", M, "samples -> max relative error: ", max_rel_err,
-              "in", (time.time()-tStart)/60., "min")
+              "in", (time.clock()-tStart)/60., "min")
         if max_rel_err < rel_tol:
             break
         nextM *= 2

### 2017-02-07 tests/matern/misc_run.py

@@ -85,7 +85,7 @@ class MyRun:
             return
 
         import time
-        tStart = time.time()
+        tStart = time.clock()
         # estimate rates
         self.d_err_rates, \
             s_fit_rates = misc.estimate_misc_error_rates(d=run.params.min_dim,
@@ -108,9 +108,9 @@ class MyRun:
         else:
             s_err_rates = []
 
-        tEnd_rates = time.time() - tStart
+        tEnd_rates = time.clock() - tStart
         ######### Update
-        tStart = time.time()
+        tStart = time.clock()
         self.profCalc = setutil.MISCProfCalculator(self.d_err_rates +
                                                    self.d_work_rates,
                                                    s_err_rates)

### 2017-02-07 tests/particle/mimc_run.py

@@ -23,7 +23,7 @@ class ParticleField(object):
     def SampleQoI(self, run, inds, M):
         meshes = (1./run.fn.Hierarchy(inds)).astype(np.int)
         import time
-        tStart = time.time()
+        tStart = time.clock()
         Ps = meshes[:, 0]
         Ns = meshes[:, 1] if run.params.min_dim == 2 \
                           else (4*meshes[:, 0]/run.params.h0inv[0]).astype(np.int)
@@ -33,7 +33,7 @@ class ParticleField(object):
                                           sig=run.params.qoi_sig,
                                           var_sig=False,
                                           antithetic=True, dim=1)
-        return samples, time.time()-tStart
+        return samples, time.clock()-tStart
 
 if __name__ == "__main__":
     import mimclib.test

### 2017-02-07 tests/pde/mimc_run.py

@@ -12,13 +12,13 @@ warnings.filterwarnings("ignore", category=mimclib.test.ArgumentWarning)
 
 def SamplePDE(sf, run, inds, M):
     import time
-    timeStart = time.time()
+    timeStart = time.clock()
     solves = np.zeros((M, len(inds)))
     sf.BeginRuns(1./run.fn.Hierarchy(inds))
     for m in range(0, M):
         solves[m, :] = sf.Sample()
     sf.EndRuns()
-    return solves, time.time() - timeStart
+    return solves, time.clock() - timeStart
 
 if __name__ == "__main__":
     from pdelib.SField import SField

### 2017-02-07 tests/sc/mimc_run.py

@@ -90,7 +90,7 @@ class MyRun:
         return
 
         # import time
-        # tStart = time.time()
+        # tStart = time.clock()
         # # estimate rates
         # self.d_err_rates, \
         #     s_fit_rates = misc.estimate_misc_error_rates(d=0,
@@ -99,9 +99,9 @@ class MyRun:
         #                                                  d_err_rates=np.zeros([]),
         #                                                  lev2knots=lambda beta:misc.lev2knots_doubling(1+beta))
         # #################### extrapolate error rates
-        # tEnd_rates = time.time() - tStart
+        # tEnd_rates = time.clock() - tStart
         # ######### Update
-        # tStart = time.time()
+        # tStart = time.clock()
         # self.profCalc = setutil.MISCProfCalculator(np.zeros(0), s_fit_rates)
         # mimc.extend_prof_lvls(lvls, self.profCalc, run.params.min_lvls)
 

### 2017-02-07 tests/matern/miproj_esterr.py
MC estimation should work now

@@ -10,7 +10,7 @@ import time
 
 warnings.filterwarnings("always")
 
-def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=2000):
+def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=4000):
     if len(itrs) == 0:
         return np.array([])
     max_L = 1 + np.max([np.max([a[0] for a in itr.lvls_itr()]) for itr in itrs])
@@ -33,7 +33,7 @@ def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=2000):
     np.random.seed(0)
     while M < maxM:
         Y = np.random.uniform(-1, 1, size=(nextM-M, N))
-        samples = fnSample(itr.parent, [max_L], Y)
+        samples = fnSample([max_L], Y)
         errors = np.zeros((nextM-M, len(itrs)))
         for i in xrange(0, len(itrs)):
             errors[:, i] = (samples - val[i](Y))**2
@@ -50,11 +50,18 @@ def l2_error_mc(itrs, fnSample, rel_tol=0.1, maxM=2000):
     return np.sqrt(s1/M)
 
 if __name__ == "__main__":
+    from matern import SField_Matern
     from miproj_run import MyRun
     sampler = MyRun()
+    sampler.sf = None
+    SField_Matern.Init()
+    def fnSample(iters):
+        if sampler.sf is None:
+            sampler.sf = SField_Matern(iters[0].parent.params)
+        return l2_error_mc(iters, sampler.solveFor_seq)
 
     from mimclib import ipdb
     ipdb.set_excepthook()
     from mimclib import test
-    test.run_errors_est_program(lambda iters:
-                                l2_error_mc(iters, sampler.solveFor_seq))
+    test.run_errors_est_program(fnSample)
+    SField_Matern.Final()

### 2017-08-01 mimclib/miproj.py
Fixed plotting error

@@ -382,11 +382,14 @@ class MIWProjSampler(object):
             projection_time = time.clock() - tStart
             # For now, only compute sampling time
             time_taken = sam_col.sampling_time + sam_col.pt_sampling_time
-            time_taken += assembly_time_1 + assembly_time_2 + projection_time
+            #time_taken += assembly_time_1 + assembly_time_2 + projection_time
+            print(sam_col.sampling_time, sam_col.pt_sampling_time,
+                  assembly_time_1, assembly_time_2, projection_time)
 
             total_time[sel_lvls] = time_taken * totalN_per_beta / np.sum(totalN_per_beta)
             total_work[sel_lvls] = work_per_sample * totalN_per_beta + \
                                    self.proj_sample_ratio * np.cumsum(totalN_per_beta) * np.cumsum(totalBasis_per_beta)
+            total_work[sel_lvls] += assembly_time_1 + assembly_time_2 + projection_time
             self.user_data.append(Bunch(alpha=alpha,
                                         max_cond=max_cond,
                                         matrix_size=BW.shape[1],

### 2017-08-01 mimclib/plot.py

@@ -15,10 +15,18 @@ from matplotlib.ticker import MaxNLocator
 __all__ = []
 
 
+
+def unique_rows(A):
+    sorted_idx = np.lexsort(A.T)
+    sorted_Ar =  A[sorted_idx,:]
+    mask = np.append(True,np.any(np.diff(sorted_Ar,axis=0),1))
+    unq_count = np.bincount(mask.cumsum()-1)
+    return sorted_Ar[mask][np.nonzero(unq_count==1)[0]]
+
 def _scatter(ax, x, y, *args, **kwargs):
-    xy = np.unique(np.vstack(x,y), axis=0)
+    xy = unique_rows(np.vstack(x,y).transpose())
     return ax.scatter(xy[:, 0], xy[:, 1], *args, **kwargs)
-    
+
 def public(sym):
     __all__.append(sym.__name__)
     return sym
@@ -406,7 +414,7 @@ def computeIterationStats(runs, fnItrStats, arr_fnAgg, work_bins=50,
 def plotDirections(ax, runs, fnPlot,
                    fnNorm=None,
                    plot_fine=False, plot_estimate=False,
-                   directions=None, rate=None, 
+                   directions=None, rate=None,
                    dir_kwargs=[{}],
                    label_fmt='{label}', beta=None):
     if directions is None:
@@ -462,7 +470,7 @@ def plotDirections(ax, runs, fnPlot,
                                        np.log(fit_data[:, 1]))[0]
                 else:
                     cur_rate = ratefit(np.log(fit_data[:, 0]),
-                                       np.log(fit_data[:, 1]))[0]                   
+                                       np.log(fit_data[:, 1]))[0]
             add_rates[np.round(cur_rate, 2)] = fit_data
         elif rate is not None:
             ind = np.nonzero(np.array(direction) != 0)[0]
@@ -1272,11 +1280,11 @@ def add_legend(ax, handles=None, labels=None, alpha=0.5,
         # Shrink current axis by 20%
         box = ax.get_position()
         ax.set_position([box.x0, box.y0, box.width * 0.8, box.height])
-        kwargs['loc'] = 'center right'
+        kwargs['loc'] = 'center left'
         kwargs['fancybox'] = False
         kwargs['frameon'] = False
         kwargs['shadow'] = False
-        kwargs['bbox_to_anchor'] = (1, 0.5)
+        kwargs['bbox_to_anchor'] = (1., .5)
         return ax.legend(handles, labels, *args, **kwargs)
     else:
         kwargs['loc'] = loc

### 2017-08-01 tests/matern/miproj_run.py

@@ -159,7 +159,6 @@ class MyRun:
             max_dim = np.minimum(run.params.miproj_max_vars + run.params.min_dim,
                                  np.maximum(run.params.miproj_min_vars + run.params.min_dim,
                                             max_dim))
-        tStart = time.clock()
         if self.profit_calc is None:
             # Adaptive
             error = run.fn.Norm(run.last_itr.calcDeltaEl())
@@ -184,7 +183,7 @@ class MyRun:
         if len(lvls) >= 1:
             max_lvls = lvls.to_sparse_matrix().max(axis=0).todense()
             if max_lvls[0, 0] > run.params.miproj_max_lvl:
-                return  False # No more levels
+                return False  # No more levels
         return True
 
     def addExtraArguments(self, parser):

### 2017-08-01 tests/matern/miproj_run.sh

@@ -2,7 +2,7 @@
 
 # make
 # rm -f data.sql
-VERBOSE="-db True -mimc_verbose 0 "
+VERBOSE="-db False -mimc_verbose 1 "
 SEL_CMD="$1"
 EXAMPLE="$2"
 if [ -z "$EXAMPLE" ]; then
@@ -91,11 +91,11 @@ if [ "$EXAMPLE" = "sf-kink" ]; then
 
         # all_cmds "-adapt" 2 $max_lvl $N -miproj_max_vars $N \
         #          -miproj_s_proj_sample_ratio 0. -miproj_set_maxadd 1 \
-        #          -miproj_set apriori-adapt -mimc_min_dim 1 $CMN  
+        #          -miproj_set apriori-adapt -mimc_min_dim 1 $CMN
 
         all_cmds "-adapt-time" 2 $max_lvl $N -miproj_max_vars $N \
                  -miproj_s_proj_sample_ratio 0. -miproj_set_maxadd 1 \
-                 -miproj_time True -miproj_set apriori-adapt -mimc_min_dim 1 $CMN  
+                 -miproj_time True -miproj_set apriori-adapt -mimc_min_dim 1 $CMN
 
         # all_cmds "-noproj" 2 $max_lvl $N -miproj_max_vars $N \
         #          -miproj_s_alpha $ALPHA -miproj_s_proj_sample_ratio 0. \

### 2017-08-03 mimclib/mimc.py
Added a user function for bias estimation

@@ -427,7 +427,6 @@ class MIMCRun(object):
     def __init__(self, **kwargs):
         self.fn = Bunch(# Hierarchy=None, ExtendLvls=None,
                         # WorkModel=None, SampleAll=None,
-                        # ItrDone=None,
                         Norm=np.abs)
         self.params = Bunch(**kwargs)
         self.iters = []
@@ -504,6 +503,11 @@ class MIMCRun(object):
                                                              self.params.h0inv,
                                                              self.params.beta)
 
+        if not hasattr(self.fn, "EstimateBias"):
+            self.fn.EstimateBias = (self._estimateBayesianBias
+                                    if self.params.bayesian
+                                    else self._estimateBias)
+
         if self.params.dynamic_lvls and not hasattr(self.fn, "WorkModel"):
             raise NotImplementedError("Bayesian parameter fitting is only \
 supported with a given work model")
@@ -526,8 +530,9 @@ supported with a given work model")
                 raise ValueError("No default ExtendLvls for ")
 
             profCalc = setutil.TDFTProfCalculator(weights)
-            self.fn.ExtendLvls = lambda lvls: extend_prof_lvls(lvls, profCalc,
-                                                               self.params.min_lvl)
+            self.fn.ExtendLvls = lambda s=self: extend_prof_lvls(s,
+                                                                 profCalc,
+                                                                 s.params.min_lvl)
 
     def setFunctions(self, **kwargs):
         # fnSampleLvl(moments, mods, inds, M):
@@ -541,7 +546,8 @@ supported with a given work model")
             kk = k[2:] if k.startswith('fn') else k
             if kk not in ["SampleLvl", "SampleAll", "ExtendLvls",
                           "ItrStart", "ItrDone", "WorkModel",
-                          "Hierarchy", "SampleQoI", "Norm"]:
+                          "EstimateBias", "Hierarchy", "SampleQoI",
+                          "Norm"]:
                 raise KeyError("Invalid function name `{}`".format(kk))
             if kk == "SampleLvl":
                 if kwargs[k] is not None:
@@ -691,35 +697,30 @@ Bias={:.12e}\nStatErr={:.12e}\
         return self.fn.Norm(np.array([x]))[0]
 
     def _estimateBias(self):
-        if self.last_itr.lvls_count <= 1:
-            # Cannot estimate bias with only one level
-            return np.inf
[…]

### 2017-08-03 mimclib/mimc.py
Updated names of some fields to be more consistent.

@@ -22,19 +22,20 @@ def public(sym):
     return sym
 
 class Timer():
-    def __init__(self):
+    def __init__(self, clock=time.clock):
         self._tics = []
+        self._clock = clock
         self.tic()
 
     def tic(self):
-        self._tics.append(time.clock())
+        self._tics.append(self._clock())
 
     def toc(self, pop=True):
         assert(len(self._tics) > 0)
         if pop:
-            return time.clock()-self._tics.pop()
+            return self._clock()-self._tics.pop()
         else:
-            return time.clock()-self._tics[-1]
+            return self._clock()-self._tics[-1]
 
     def ptoc(self, msg='Time since last tic: {:.4f} sec.', pop=False):
         assert(len(self._tics) > 0)
@@ -164,7 +165,7 @@ class MIMCItrData(object):
         self.stat_error = np.inf     # Sampling error (based on M)
         self.exact_error = np.nan    # Sampling error (based on M)
         self.TOL = None              # Target tolerance
-        self.totalTime = None
+        self.total_time = None
         self.Q = None
         self.Vl_estimate = np.zeros(0)
         self._lvls_count = 0
@@ -185,7 +186,7 @@ class MIMCItrData(object):
         ret.M = self.M.copy()
         ret.bias = self.bias
         ret.stat_error = self.stat_error
-        ret.totalTime = self.totalTime
+        ret.total_time = self.total_time
         ret.TOL = self.TOL
         ret.Q = copy.copy(self.Q)
         ret.Vl_estimate = self.Vl_estimate.copy() if self.Vl_estimate is not None else None
@@ -480,7 +481,7 @@ class MIMCRun(object):
 
     @property
     def iter_total_times(self):
-        return np.cumsum([itr.totalTime for itr in self.iters])
+        return np.cumsum([itr.total_time for itr in self.iters])
 
     @property
     def iter_calc_total_times(self):
@@ -533,7 +534,7 @@ supported with a given work model")
         # fnSampleLvl(moments, mods, inds, M):
         #    Returns M, array: M sums of mods*inds, and total
         #    (linear) time it took to compute them
-        # fnItrDone(i, TOLs, totalTime): Called at the end of iteration
+        # fnItrDone(i, TOLs, total_time): Called at the end of iteration
         #    i out of TOLs
         # fnWorkModel(lvls): Returns work estimate of lvls
         # fnHierarchy(lvls): Returns associated hierarchy of lvls
[…]

### 2017-08-03 mimclib/miproj.py

@@ -262,8 +262,9 @@ class MIWProjSampler(object):
 
         # Add previous times and works from previous iteration
         from .mimc import Bunch
+        timer = mimc.Timer(clock=time.time)
         for alpha, ind in self.alpha_dict.iteritems():
-            tStart = time.clock()
+            timer.tic()
             sam_col = self.prev_samples[ind]
             sel_lvls = np.nonzero(self.alpha_ind == ind)[0]
             work_per_sample = self.fnWorkModel(setutil.VarSizeList([alpha]))[0]
@@ -274,9 +275,6 @@ class MIWProjSampler(object):
 
             beta_indset = lvls.sublist(sel_lvls[sam_col.beta_count:],
                                        d_start=self.d, min_dim=0)
-
-            # TODO: This assumes that beta_indset is ordered (in 1D)
-            #new_basis = setutil.VarSizeList()
             for i, beta in enumerate(beta_indset):
                 new_b = self.fnBasisFromLvl(beta)
                 if isinstance(new_b, setutil.VarSizeList):
@@ -319,22 +317,22 @@ class MIWProjSampler(object):
                 X, N_done = self.fnSamplePoints(N_todo, sam_col.basis, sam_col.min_dim)
                 N_done[:len(sam_col.N_per_basis)] += sam_col.N_per_basis
                 sam_col.N_per_basis = N_done
-                sam_col.pt_sampling_time += time.clock() - tStart
+                sam_col.pt_sampling_time += timer.toc()
 
-                tStart = time.clock()
+                timer.tic()
                 sam_col.add_points(fnSample, sub_alphas, X)
-                sam_col.sampling_time += time.clock() - tStart
+                sam_col.sampling_time += timer.toc()
             else:
-                sam_col.pt_sampling_time += time.clock() - tStart
+                sam_col.pt_sampling_time += timer.toc()
 
-            tStart = time.clock()
+            timer.tic()
             # sam_col.basis_values.re size(len(sam_col.X), len(sam_col.basis))
             # TODO: should only compute new basis_values
             basis_values = TensorExpansion.evaluate_basis(
                 self.fnBasis, sam_col.basis, sam_col.X)
             W = self.fnWeightPoints(sam_col.X, basis_values)
-            assembly_time_1 = time.clock() - tStart
-            tStart = time.clock()
+            assembly_time_1 = timer.toc()
+            timer.tic()
             from scipy.linalg import solve
[…]

### 2017-08-03 mimclib/plot.py

@@ -14,9 +14,30 @@ from matplotlib.ticker import MaxNLocator
 
 __all__ = []
 
+def unique_rows(A, return_index=False, return_inverse=False):
+    """
+    Similar to MATLAB's unique(A, 'rows'), this returns B, I, J
+    where B is the unique rows of A and I and J satisfy
+    A = B[J,:] and B = A[I,:]
 
+    Returns I if return_index is True
+    Returns J if return_inverse is True
+    """
+    A = np.require(A, requirements='C')
+    assert A.ndim == 2, "array must be 2-dim'l"
+
+    B = np.unique(A.view([('', A.dtype)]*A.shape[1]),
+                  return_index=return_index,
+                  return_inverse=return_inverse)
+
+    if return_index or return_inverse:
+        return (B[0].view(A.dtype).reshape((-1, A.shape[1]), order='C'),) \
+            + B[1:]
+    else:
+        return B.view(A.dtype).reshape((-1, A.shape[1]), order='C')
+                    
 def _scatter(ax, x, y, *args, **kwargs):
-    xy = np.unique(np.vstack(x,y), axis=0)
+    xy = unique_rows(np.vstack((x,y)).transpose())
     return ax.scatter(xy[:, 0], xy[:, 1], *args, **kwargs)
     
 def public(sym):
@@ -406,7 +427,7 @@ def computeIterationStats(runs, fnItrStats, arr_fnAgg, work_bins=50,
 def plotDirections(ax, runs, fnPlot,
                    fnNorm=None,
                    plot_fine=False, plot_estimate=False,
-                   directions=None, rate=None, 
+                   directions=None, rate=None,
                    dir_kwargs=[{}],
                    label_fmt='{label}', beta=None):
     if directions is None:
@@ -462,7 +483,7 @@ def plotDirections(ax, runs, fnPlot,
                                        np.log(fit_data[:, 1]))[0]
                 else:
                     cur_rate = ratefit(np.log(fit_data[:, 0]),
-                                       np.log(fit_data[:, 1]))[0]                   
+                                       np.log(fit_data[:, 1]))[0]
             add_rates[np.round(cur_rate, 2)] = fit_data
         elif rate is not None:
             ind = np.nonzero(np.array(direction) != 0)[0]
@@ -503,14 +524,14 @@ def plotErrorsVsTOL(ax, runs, *args, **kwargs):
     run_data[i].db_data.Creation_date
[…]

### 2017-08-03 tests/matern/miproj_run.sh
Some refining of miproj runs

@@ -9,7 +9,7 @@ if [ -z "$EXAMPLE" ]; then
     EXAMPLE='sf-kink'
 fi
 DB_CONN='-db_engine mysql -db_name mimc -db_host 129.67.187.118 '
-BASETAG="sim-$EXAMPLE-"
+BASETAG="$EXAMPLE-"
 COMMON="-qoi_seed 0 -ksp_rtol 1e-15 -ksp_type gmres  $DB_CONN "
 EST_CMD="python miproj_esterr.py $COMMON "
 RUN_CMD="OPENBLAS_NUM_THREADS=1 python miproj_run.py -qoi_example $EXAMPLE \
@@ -76,26 +76,24 @@ fi;
 
 if [ "$EXAMPLE" = "sf-kink" ]; then
     CMN='-qoi_sigma -1 -mimc_beta 1.4142135623730951 -qoi_scale 0.5 '
-    for N in 2 3 4 6
+    for N in 2
     do
-        max_lvl=10
-        # (gamma_space + w_space) / (N + kappa)
-        #ALPHA=`echo "3./$N" | bc -l`
-	ALPHA=3.
-	THETA=`echo "$N/2" | bc -l`
+        max_lvl=12
+	    ALPHA=3.
+	    THETA=`echo "$N/2" | bc -l`
 
-        # all_cmds "-tdfit" 2 $max_lvl $N -miproj_max_vars $N \
+        # all_cmds "-td-theory" 2 $max_lvl $N -miproj_max_vars $N \
         #          -miproj_s_alpha $ALPHA -miproj_s_proj_sample_ratio 0. \
         #          -miproj_s_theta $THETA -miproj_d_beta 1. -miproj_d_gamma 1. \
         #          -miproj_set apriori -mimc_min_dim 1 $CMN  -miproj_double_work True
 
-        # all_cmds "-adapt" 2 $max_lvl $N -miproj_max_vars $N \
-        #          -miproj_s_proj_sample_ratio 0. -miproj_set_maxadd 1 \
-        #          -miproj_set apriori-adapt -mimc_min_dim 1 $CMN  
-
-        all_cmds "-adapt-time" 2 $max_lvl $N -miproj_max_vars $N \
+        all_cmds "-adapt" 2 $max_lvl $N -miproj_max_vars $N \
                  -miproj_s_proj_sample_ratio 0. -miproj_set_maxadd 1 \
-                 -miproj_time True -miproj_set apriori-adapt -mimc_min_dim 1 $CMN  
[…]

### 2017-08-03 tests/matern/plot_miproj_paper.py

@@ -25,14 +25,14 @@ def plotProfits(ax, itr, *args, **kwargs):
         work = itr.calcTl()
     else:
         work = itr.calcWl()
-        
+
     lvls = list(itr.lvls_itr(min_dim=2))
     assert(np.all([len(l) == 2 for l in lvls]))
     lvls = np.array(lvls)
     prof = setutil.calc_log_prof_from_EW(error, work)
 
     max_lvl = np.max(lvls, axis=0)
-    
+
     X, Y = np.meshgrid(np.arange(0, max_lvl[0]+1), np.arange(0, max_lvl[1]+1))
     data = np.zeros((max_lvl[1]+1, max_lvl[0]+1))
     data.fill(np.nan)
@@ -42,7 +42,7 @@ def plotProfits(ax, itr, *args, **kwargs):
     ax.contourf(X, Y, data)
     ax.set_xlabel('$\\ell_1$')
     ax.set_ylabel('$\\ell_2$')
-                                
+
 def plotSeeds(ax, runs, *args, **kwargs):
     ax.set_yscale('log')
     ax.set_xscale('log')
@@ -113,8 +113,8 @@ def plotUserData(ax, runs, *args, **kwargs):
     which = kwargs.pop('which', 'cond').lower()
     def fnItrStats(run, i):
         itr = run.iters[i]
-        max_cond = np.max([d.max_cond for d in itr.db_data.user_data])
-        max_size = np.max([d.matrix_size for d in itr.db_data.user_data])
+        max_cond = np.max([d.max_cond for d in itr.userdata])
+        max_size = np.max([d.matrix_size for d in itr.userdata])
         return [i, max_size, max_cond]
 
     xy_binned = miplot.computeIterationStats(runs,
@@ -208,12 +208,12 @@ def plot_all(runs, **kwargs):
     except:
         miplot.plot_failed(ax)
 
-        
+
     print_msg("plotPorfits")
     ax = add_fig('profits')
     plotProfits(ax, runs[0].last_itr)
     ax.set_title('Err/Work')
-    
+
     ax = add_fig('profits')
     plotProfits(ax, runs[0].last_itr, work_est='time')
     ax.set_title('Err/Time')
@@ -249,7 +249,7 @@ def plot_all(runs, **kwargs):
     except:
         miplot.plot_failed(ax)
         raise
-    
+
     print_msg("plotDirections")
[…]
